### 2018-01-09 qg_ML_3.tex
YT uploaded the current draft

\documentclass[notoc]{JHEP3}
\usepackage{cite,color,psfrag,url,latexsym,graphicx,epstopdf,slashed,xspace,hyperref,enumitem,lineno}
\let\normalcolor\relax

\newcommand{\event}{ {\sc event2}}
\newcommand{\In}{ {\rm in}}
\newcommand{\Out}{ {\rm out}}
\newcommand{\nbar}{ {\bar{n}}}

\newcommand{\ds}{\displaystyle}
\newcommand{\Ord}{\mathcal{O}}
\newcommand{\sL}{\mathcal{L}}
\newcommand{\nn}{\nonumber \\}
\newcommand{\el}{\nonumber \\ &&}
\newcommand{\ezint}[3]{\int_{#1}^{#2} d #3}
\newcommand{\mint}[1]{\int {d^d #1 \over (2 \pi)^d}}
\newcommand{\mintd}[2]{\int {d^{#2} #1 \over (2 \pi)^{#2}}}
\newcommand{\mintb}[1]{\int d^d #1}
\newcommand{\M}{\mathcal{M}}
\newcommand{\T}{\theta}
\newcommand{\pt}{$p_{T}$}
\newcommand{\eqref}[1]{(\ref{#1})}
\newcommand{\text}[1]{{\rm #1}}
\newcommand{\jw}{\textsc{Jewel}~}
\newcommand{\jwpy}{\textsc{Jewel+Pythia}~}
\newcommand{\py}{\textsc{Pythia}~}
\newcommand{\bS}{\hat{S}}
\newcommand{\rS}{S_r}
\newcommand{\GNGL}{\Gamma^{\mathrm{NGL}}}
\newcommand{\gsreg}{\gamma^{\mathrm{NGL}}}

\newcommand{\thR}{R}
\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\def\PlusBreak#1{+ \nonumber \\          % splits line with plus sign
          &&  \hphantom{#1} \! \null   +}
\def\MinusBreak#1{- \nonumber \\         % splits line with minus sign
          &&  \hphantom{#1} \!  \null  -}
\def\TimesBreak#1{\times \nonumber \\    % splits lines with times sign
          &&  \hphantom{#1} \!  \null  \times}

% The following macros allow line splits allowing for line spacing adjustments
% 12 is line spacing e.g. 1pt plus 4pt and #2 is the hphantom
\def\PlusBreakAdjust#1#2{+ \nonumber \\ [#1]      % splits line with plus sign
          &&  \hphantom{#2} \! \null  +}
\def\MinusBreakAdjust#1#2{- \nonumber \\ [#1]     % splits line with minus sign
          &&  \hphantom{#2} \!  \null -}
\def\TimesBreakAdjust#1#2{\times \nonumber \\[#1] %splits lines with times sign
          &&  \hphantom{#2} \!  \null  \times}

\def\eqn#1{Eq.~(\ref{eq:#1})}
\def\Eqn#1{Eq.~(\ref{eq:#1})}
\def\eqns#1#2{Eqs.~(\ref{eq:#1}) and~(\ref{eq:#2})}
\def\Eqns#1#2{Eqs.~(\ref{eq:#1}) and~(\ref{eq:#2})}
\def\fig#1{Figure~{\ref{fig:#1}}}
%\renewcommand{\sec}[1]{Section~\ref{sec:#1}}
\newcommand{\ssec}[1]{Section~\ref{ssec:#1}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\cmb}{\begin{changemargin}}
\newcommand{\cme}{\end{changemargin}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
% \newcommand{\bea}{\begin{align}}
% \newcommand{\eea}{\end{align}}

\newcommand{\highlightA}{black}
\newcommand{\highlight}{black}

\newcommand{\D}{\delta}
\newcommand{\g}{\gamma}
\newcommand{\A}{\alpha}
\newcommand{\B}{\beta}
\newcommand{\G}{\Gamma}
\newcommand{\la}{\leftarrow}
\newcommand{\ra}{\rightarrow}
\newcommand{\mc}{\mathcal}
\newcommand{\mb}{\mathbb}
\newcommand{\gcusp}{\Gamma_{\rm cusp}}
\newcommand{\rr}{\frac{R}{1-R}}
\newcommand{\lib}{{\mathrm{Li}_2}}
\newcommand{\lic}{{\mathrm{Li}_3}}
\newcommand{\tauo}{\tau_\omega}
\newcommand{\bn}{{\bar{n}}}
\newcommand{\ecm}{E_{\rm CM}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\als}{\alpha_s}
\newcommand{\ep}{\epsilon}
\newcommand{\ws}{\widetilde{s}}
\def\Tr{\mathop{\rm Tr}\nolimits}
\def\TrFf{\mathop{\rm Tr}\nolimits F^4}
\def\si{\sigma}
%\newcommand{\SU}{\mathop{\rm SU}\nolimits}
\def\spa#1.#2{\langle#1\,#2\rangle}
\def\spb#1.#2{[#1\,#2]}
\def\sandmm#1.#2.#3{%
\left\langle\smash{#1}{\rphantom1}\right|{#2}%
\left|\smash{#3}{\rphantom1}\right]}
\def\spab#1.#2.#3{\sandmm#1.#2.#3}
\def\spba#1.#2.#3{\sandpp#1.#2.#3}
\def\spaa#1.#2.#3.#4{\sandmp#1.{#2#3}.#4}
\def\spbb#1.#2.#3.#4{\sandpm#1.{#2#3}.#4}
\def\spash#1.#2{\spa{\smash{#1}}.{\smash{#2}}}
\def\spbsh#1.#2{\spb{\smash{#1}}.{\smash{#2}}}

\def\ksl{\not{\hbox{\kern-2.3pt $k$}}}
\def\lsim{\lesssim}
\def\gsim{\gtrsim}
\def\susy{{\rm susy}}
\def\ib{{\bar\imath}}
\def\cg{c_\Gamma}
\def\e{\epsilon}
\def\musq{\mu^2}
\def\tree{{\rm tree}}
\def\oneloop{{\rm 1\! -\! loop}}
\def\twoloop{{\rm 2\! -\! loop}}
\def\Ord{{\cal O}}
\def\a{{\cal A}}
\def\cm{{\cal M}}
\def\Ft{\tilde{F}}
\def\Re{{\rm Re}}
\def\Nc{N_c}
\def\Nsym{{\cal N}=4}
\def\la{\langle}
\def\ra{\rangle}
\def\bom#1{{\mbox{\boldmath $#1$}}}
%
%
\def\lr{\leftrightarrow}
\def\li#1{{\mathop{\rm Li}\nolimits}_#1}
\def\Li{\mathop{\rm Li}\nolimits}
\def\ggtogg{{gg \to gg}}
\def\alphas{\alpha_s}
\def\draftnote#1{{\it #1}}

\def\rg{r_\Gamma}
\def\Triint{{\rm Tri}}
\def\Boxfour{{\rm Box}^{(4)}}
\def\Boxsix{{\rm Box}^{(6)}}
\def\Boxeight{{\rm Box}^{(8)}}

\def\I{{\cal I}}
\def\pol{\varepsilon}

\def\in{{\rm in}}
\def\out{{\rm out}}
\def\tom{\tau_\omega}
\newcommand{\rd}{\mathrm{d}}

\newcommand{\note}[1]{\marginpar{\footnotesize\textbf{NOTE}} (\textbf{\textcolor{blue}{#1}})}
\newcommand{\todo}[1]{\marginpar{\textbf{\textcolor{blue}{TODO}} \footnotesize{#1}} (\textbf{\textcolor{blue}{TODO}})}

\DeclareRobustCommand{\Sec}[1]{Sec.~\ref{#1}}
\DeclareRobustCommand{\Secs}[2]{Secs.~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Secss}[3]{Secs.~\ref{#1}, \ref{#2} and \ref{#3}}
\DeclareRobustCommand{\App}[1]{App.~\ref{#1}}
\DeclareRobustCommand{\Tab}[1]{Table~\ref{#1}}
\DeclareRobustCommand{\Tabs}[2]{Tables~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Fig}[1]{Fig.~\ref{#1}}
\DeclareRobustCommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Figss}[3]{Figs.~\ref{#1}, \ref{#2} and \ref{#3}}
\DeclareRobustCommand{\Eq}[1]{Eq.~(\ref{#1})}
\DeclareRobustCommand{\Eqs}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
\DeclareRobustCommand{\Eqss}[3]{Eqs.~(\ref{#1}), (\ref{#2}), and (\ref{#3})}
\DeclareRobustCommand{\Ref}[1]{Ref.~\cite{#1}}
\DeclareRobustCommand{\Refs}[2]{Refs.~\cite{#1} and \cite{#2}}

\title{Probing heavy ion collisions using quark and gluon jet substructure with machine learning}

\author{Yang-Ting Chien $^{a}$ and Raghav Kunnawalkam Elayavalli $^{b}$\\
$^{a}$ Center for Theoretical Physics\\
$~$ Massachusetts Institute of Technology, Cambridge, MA 02139, USA\\
$^{b}$ Department of Physics and Astronomy\\
$~$ Rutgers, the State University of New Jersey, New Brunswick, NJ 08901, USA\\
}

\abstract{
We study the classification of quark-initiated jets and gluon-initiated jets in proton-proton and heavy ion collisions using modern machine learning techniques. We train the deep convolutional neutral network on discretized jet images. The classification performance is compared with the multivariate analysis of several physically-constructed jet observables including the jet mass, the $p_T^D$, the multiplicity and the radial moments. We also compare with the systematic $N$ subjet expansion in telescoping deconstruction to exploit the information carried by the subjets. The quark and gluon jet samples generated from \jw are used as an example to demonstrate this general method. We find that the classification performance goes down in the \textsc{Jewel}-simulated heavy ion collisions. The information carried by the subleading subjets can be washed out by the possible subjet thermalization or randomization due to the soft event activities. Our method provides a systematically improvable framework for analyzing and comparing all jet simulations and measurements in heavy ion collisions.
}

\begin{document}

\section{Introduction}
\label{sec:intro}

The jet quenching phenomenon observed at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
and the Large Hadron Collider (LHC)\cite{Aamodt:2010jd,Abelev:2012hxa,Abelev:2013kqa,Aad:2012vca,Aad:2014bxa,Chatrchyan:2013kwa,
Chatrchyan:2012gw,Chatrchyan:2014ava,Aad:2014wha,Chatrchyan:2013exa,
Adam:2015ewa,Chatrchyan:2012gt,Chatrchyan:2011sx,Chatrchyan:2012nia,Aad:2010bu,Aad:2013sla,Adam:2015doa,Aad:2015bsa} has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions. Traditionally jet quenching refers to the dramatic suppression of hadron and jet cross sections, which is translated into the parton and jet energy loss picture. It is now established that the lost energy distributes within an angle of $O(1)$ from the jet through the missing transverse momentum measurement. By studying in details how the jet energy redistributes in heavy ion collisions, one can quantitatively extract the information about the jet-medium interaction and the underlying medium dynamics. The heavy ion jet physics has thus moved to the era of precision jet substructure and classified jet cross section studies.

The jet substructure measurements provide concrete information about the modifications of jets. From the measurements of the jet shape and jet fragmentation function, we have learned that both the transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructures depend strongly on the partonic origin of jets, the change of the fraction of quark-initiated jets and gluon-initiated jets will contribute to the substructure modifications. It has been realized that the increase of the quark jet fraction in heavy ion collisions constitutes the narrowing of the jet cores seen in various inclusive jet substructure measurements. With purer quark and gluon jet samples, one can disentangle the effect from the changing quark and gluon jet fractions and focus on the jet-by-jet modifications. Quark jets and gluon jets can then be used as independent probes.

We will study the tagging of quark and gluon jets making full use of their substructure information. In contrast to the conventional wisdom of directly comparing jets in proton-proton and heavy ion collisions, this indirect approach studies the modification of the comparison between different probes. By first comparing the probes in the same system, it may help reducing the systematic uncertainty in experimental measurements. The lowest order feature which separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons. The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets. Various jet substructure variables have been used and combined to further increase the tagging performance with a multivariate analysis, and the advances of machine learning techniques have been constantly incorporated in solving this classification problem.

In this paper we benchmark three classes of methods in quark and gluon tagging and gain insights from comparing their performances. We use modern image recognition techniques utilizing deep convolutional neural networks (DCNN). The network is trained on the discretized images of quark jets and gluon jets, and the samples are generated from the \jw heavy ion event generator. Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments. We also examine the tagging performance using five physically constructed jet substructure variables: the jet mass, the $p_T^D$, the multiplicity and two radial moments including the girth. Subsets of the variables are combined and studied in a multivariate analysis using boosted decision trees, and we look into the correlations among the variables. Lastly, we systematically expand the information within jets using the recently developed telescoping deconstruction method. The corresponding subjet expansion can be truncated at any finite order $N$ which allows us to access the information carried by the subleading subjets.

The rest of the paper is organized as follows. In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jw simulation. In Sec.~\ref{sec:tagger} we give the details about the jet image, the multivariate analysis and the telescoping deconstruction methods. In Sec.~\ref{sec:results} we show the performance of each method and the distributions of jet substructure observables. In Sec.~\ref{sec:conc} we conclude and give an outlook of future studies.

\section{Jet Samples}
\label{sec:sample}

We use the \textsc{Jewel} simulation to generate the quark-initiated and gluon-initiated jet samples through the prompt photon production channels \cite{}. The quark jets are generated by $pp\rightarrow q +\gamma$ in proton-proton collisions and ${\rm PbPb}\rightarrow q+\gamma$ in central (0-20\%) lead-lead collisions at at $\sqrt{s} = 2.76$ TeV using the standard \jw setup \cite{Zapp:2013zya}. Similarly gluon jets are generated by the $g +\gamma$ channel.

\jw is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions. The medium-induced energy loss in heavy ion collisions is based on a simple medium model consisting of thermal scattering centers undergoing the Bjorken expansion. We set the medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV from the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}. We use the \textsc{Cteq6LL} \cite{Pumplin:2002vw} and \textsc{Eps09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{Lhapdf}~\cite{Whalley:2005nh}.

%For a detailed description of the MC implementation of the jet production and its elastic/inelastic interactions with the scattering centers along with the LPM effect in \jw, please refer~\cite{}. The medium formation time and initial temperature are $\tau_\text{i}=0.6 $fm and  $T_\text{i}=485$ MeV are taken from a hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}. The PDF sets for pp and PbPb sample are \textsc{Cteq6LL}~\cite{Pumplin:2002vw} and \textsc{Eps09}~\cite{Eskola:2009uj} provided by \textsc{Lhapdf}~\cite{Whalley:2005nh}.The primary event collection was split into two halves, with one utilized for training the model(s) and the other for testing \textcolor{red}{additional to a di-jet sample, with a mixture of quark and gluon jets utilized for further evaluation of the model performance}.
	Jets are clustered with the anti-k$_{t}$ algorithm~\cite{} with the distance parameter $R = 0.4$ available in \textsc{fastjet}~\cite{} and the \jw HepMC files~\cite{} are analyzed using the \textsc{Rivet} analysis framework~\cite{}.

The question of what constitutes a "quark" or "gluon" jet is a topic of current discussion~\cite{} but for this paper, we choose the conventional definition of a quark/gluon jet with its progenitor hard scattered parton.

The kinematic cuts for jets in the analysis are $p^{\gamma}_{T} > 100, p^{jet}_{T} > 50$ GeV/c, $|\eta| < 2.0$ and $\Delta \phi_{\gamma, jet} > 2\pi/3$ to ensure that we are selecting the jet that recoils against the isolated $\gamma$.

	There are two operational modes of generating \jw events in heavy ion collisions; with and without including the recoil partons, those originating from interactions of the propagating hard scattered parton with the scattering centers. In recent studies, \jw with recoils was shown to accurately reproduce several of the important qualitative features in jet structure modifications measured by LHC experiments with the addition of background subtraction techniques introduced here~\cite{}. The jet-background in \jw consists of the scattering center's thermal momenta i.e, their momentum before the interaction. For our analysis, we compare the performance of the quark-gluon tagging for three different collections of \jw jets generated with recoils
	\begin{itemize}
		\item no grid and no background subtraction
		\item events discretized in a $\eta-\phi$ grid (0.08 x 0.08) and no background subtraction
		\item discretized events with GridSub background subtraction technique.
	\end{itemize}
	In studying these different jet collections, we extract the dependence of the tagging on key operations performed on jets in data, primarily discretization and background subtraction and discuss their impact on the physics.

\section{Taggers}
\label{sec:tagger}

	Introduce the different taggers in these collisions and talk about how one can study these.

\subsection{Multivariate Analysis}
\label{sec:mva}

	Multivariate analysis have recently been successfully employed in object classification and selection by the LHC experiments and thus are an effective physics driven baseline for our model comparisons. We choose the following five jet observables that highlight differences in the fragmentation patterns for quark and gluon jets as input to our model
	\begin{itemize}
		\item Number of jet constituents, including charged and neutral; given by the anti-k$_{t}$ algorithm
		\item Invariant Jet Mass - utilizing the 4-vector sum of the jet constituents  $M = \sqrt{E^{2} - p^{2}}$
		\item $p_{T}^{D}$ - motivated by the jet fragmentation $p^{D}_{T} = \frac{\sqrt{\sum_{i \in jet} p^{2}_{T,i}}}{p_{T, jet}}$
		\item 0.5 Radial moment - $\sum_{i \in jet} \frac{p_{T, i} (\Delta R_{jet, i})^{0.5}}{p_{T, jet}}$
		\item 1st radial moment (girth) - $\sum_{i \in jet} \frac{p_{T, i} (\Delta R_{jet, i})}{p_{T, jet}}$
	\end{itemize}

	 In Fig:~\ref{fig:jetdistributons_pp_pbpb} showcases the distributions of the aforementioned QCD-motivated jet observables in pp and PbPb collisions. The quark and gluon jets shown in Fig:~\ref{fig:jetdistributons_pp_pbpb} include the smearing effect of the grid and in the case of central PbPb, include the GridSub background subtraction method. All five jet observables showcase the ability to classify quarks vs gluons in both pp and PbPb. Recent results from LHC experiments on the effect of jet quenching on inclusive jets showcase two main features; a slight narrowing of the jet core and increased low \pt particle multiplicity in the periphery regions of the jet leading to a broadening of the jet shape. Such an effect is captured in \jw, with the introduction of the medium induced recoils and introduces a further complexity in the question of quark vs gluon classification in PbPb. In vacuum, we expect gluon jets to be broader than quark jets due to the different in the casimir color factor and thus in PbPb collisions, with both jets being quenched and broader, the ability to discriminate effectively between quark and gluon jets is expected to reduce when one only considers fragmentation based observables as we proceed to show in the upcoming sections.

	 However, it is important to note that these observables and their distributions in an experiment are highly dependent on the jet clustering algorithm, the minimum \pt cutoff for constituents and the background subtraction procedures employed. In order to facilitate a direct comparing of experiment with MC, one has to either unfold the detector resolution from these observables\footnote{The unfolding procedure for jet observables increases in dimensionality and complexity as the correlations between observables become strong. For example, the invariant jet mass has to be unfolded in 4-D of the generated and reconstructed jet \pt and mass} or run the MC through a simulation of the detector response.

	\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
	   \includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
	   \includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
	   \includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
	   \includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
	   \caption{Distributions of the jet constituents (top left), invariant jet mass (top right), $p^{D}_{T}$ (center left), the half radial moment (center right) and the jet girth or the first radial moment (bottom) for quark (darker blue) and gluon (lighter orange) induced jets (color online). The top panels of each individual distribution show the pp and the bottom panels correspond to central (0-20\%) PbPb events generated by the \jw generator.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

	The Multivariate model is defined in Keras with the aforementioned 5 input parameters followed by two dense layers with 10 nodes each. We use the rectified linear (\textit{relu}), and sigmoid activation functions for the dense nodes and the final output layer respectively. The model utilizes the Adam optimizer and it is trained for minimizing binary cross entropy the loss function. We perform cross validation on our MC sample by splitting it into two halves and keeping one for training and the other for testing.

\subsection{Jet Image}
\label{sec:image}

	The radiation pattern of a jet can be visualized as an image in the $\eta-\phi$ plane as seen by a detector with finite segmentation, much like a digital camera~\cite{}. The jet image thus represents a fixed dimensional representation of the energy deposition in the $\eta-\phi$ plane with 'color' of each pixel noted by its \pt. The size of each pixel in a jet image is especially important since. Each jet image follows the pre-processing prescription given in ~\cite{deOliveira:2015xxd} which includes fixing the highest pixel in the jet image as the origin (involving translations in $\eta$ and $\phi$) and rotating the jet image in the plane so that the principal component of the energy density is in the downward direction. This ensures a standard template for the jet image and helps in model performance. Average quark (left) and gluon jet (right) images in pp (top) and central PbPb (bottom) collisions are shown in Fig:~\ref{fig:qgjetimages} where the z axis shows the momentum scale of the jet constituents. As expected, we see a broader energy distribution around the center pixel when they are initiated by gluons in comparison with quarks in pp, whereas both categories of jets are broader in PbPb due to jet quenching effects.  It is important to note the scale of the pixel \pt in the z axis shows jet constituents with experimentally unobservable momenta due to the averaging procedure.

	\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgQuarkJet}
	   \includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgGluonJet}
	   \includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgQuarkJet}
	   \includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
	   \caption{Average quark (left) and gluon (right) pre-processed jet images shown in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \jw. The color scale in the z axis represents the average pixel \pt in units of GeV/c. }
	   \label{fig:ave_jetimage_jewel}
	\end{figure}

	A deep convolutional neural network implemented in Keras with tensorflow backend is utilized in this analysis, motivated by recent studies in classifications of boosted objects in pp collisions~\cite{}. We have three 2-D convolution layers of sequentially reducing sizes each with 8 filters and followed by a corresponding max-pooling layer. Similar to our MVA analysis, each convolution layer  along with a hyperbolic tangent activation function. Each convolution layer is followed by a max-pooling layer and the final output has a sigmoid activation function. The output layer is preceded by a sequential a dropout layer with a rejection score of 0.20, dense layer with 20 nodes and an additional dropout layer with a reduced rejection score of 0.10. The additional dropout layer helps in filtering less important features from previous layers and thus increases the speed and efficiency of the classification model. The deep convolutional neural network model is trained with the Adam optimizer, with a binary cross entropy loss function, similar to our MVA model described above.

    %layer = Convolution2D(8, 11, 11, border_mode='same')(input_layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(2,2))(layer)
    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(3,3))(layer)
    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(3,3))(layer)
    %layer = Flatten()(layer)
    %layer = Dropout(0.20)(layer)
    %layer = Dense(20)(layer)
    %layer = Dropout(0.10)(layer)
    %output_layer = Dense(1, activation='sigmoid', name='main_output')(layer)
    %model = Model(input=input_layer, output=output_layer)
    %model.compile(optimizer='adam', loss='binary_crossentropy', metrics=['accuracy'])

\subsection{Telescoping Deconstruction}
\label{sec:tjet}

In this section we apply telescoping deconstruction to quark/gluon discrimination \cite{Chien:2017decon} in both proton-proton and heavy ion collisions. It is a complete and systematic $N$-subjet expansion consisting simply of the subjet kinematic variables. The observables are combined using dense neural networks to maximize the tagging performance at each subjet order $N$. On the other hand, individual telescoping deconstruction observables are physically-meaningful which encode the all-order QCD splitting. We outline the procedure of the method and show how the subjet variables within this framework can reveal the full jet dynamics. In particular, we demonstrate the rich information of the subjet topology and subjet substructure contained at the T2 order, which is closely related to the groomed momentum sharing and the splitting function recently measured by the CMS collaboration. We also show how the subjet substructures allow us to extract the information about the partonic origins of the subjets. The quark/gluon tagging performances at each T$N$ order with increasing $N$ are shown to converge to the performance of the jet image technique using the convolutional neural network, and we compare the performances in proton-proton and heavy ion collisions. We see hints that the information contained in higher-order subleading subjets can be washed out in heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions.

Telescoping deconstruction probes the radiation around dominant energy flows in a jet with multiple angular resolutions. It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction and then reaching out to include wide-angle, soft radiations. The procedure is outlined as below,
    \begin{itemize}
        \item At order $N$, determine $N$ axes $\{\hat n_i\}=\{(\eta_i,\phi_i)\}$ along the dominant energy flows, where $y$ and $\phi_i$ are the rapidity and the azimuthal angle of the subjet axis $i$. We use the $N$subjettiness 1-pass $k_T$ axes and the Winner-Take-All (WTA) $k_T$ axes \cite{Thaler:2010tr}.
        \item Construct $N$ subjets with $M$ radii $\{R_{T,m}\}^M_{m=1}$ by assigning particles to the nearest axis according to the distance $d^2_{ij} = \Delta y_{ij}^2+\Delta \phi_{ij}^2$ between the axis $\hat n_i$ and the particle $j$ \cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}.
            \begin{equation}
                {\rm subjet}_{i,m} = \{p^\mu_j~|~d_{ij}<R_{T,m}~{\rm and}~d_{ij}<d_{kj} \forall i\neq k\}.
            \end{equation}
            The subjet radii ${R_{T,m}}$ are sampled within the range $(0,R)$ where $R$ is the jet radius. In this paper $\{R_{T,m}\}$ are chosen to be evenly spaced within the range.
        \item We form the subjet data with the subjet transverse momenta and masses $\{(p_T,m)_{i,m}\}$,
            \begin{equation}
                {p_T}_{i,m}=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)_T\;,~~~{m_{i,m}}^2=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)^2\;,
            \end{equation}
            where we sum over all the particles $j$ within the subjet $i,m$. Together with the positions of the axes these form the telescoping deconstruction observables.
    \end{itemize}

We train a neural network on the subjet variables up to T$N$ order. The details about the neural network model can be found in \cite{Chien:2017decon}. We consider $N=\{1,2,3\}$ and the performance quickly saturates at $N=3$. Beside the overall quark/gluon tagging performance, we also examine the subjet variables at the T2 order representing the information of the subjet topology and the subjet substructure,
    \begin{equation}
        z=\frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}}\;,~~~~~m_l=\min(m_1,m_2)\;,~~~~~m_h=\max(m_1,m_2)\;,
    \end{equation}
where $z$ is the transverse-momentum fraction of the softer subjet and $m_{l(h)}$ is the light (heavy) subjet mass.

\section{Results}
\label{sec:results}

	% Add the final plots comparing the different methods

	List of plots to add

	1. ROC curve of just pp showing the MVA, DCNN and TD (telescoping deconstruction)

	2. ROC curve for pbpb same as above

	3. direct comparison of roc curve pp and PbPb MVA, DCNN and TC separate - 3 plots

	4. Comparing PbPb w/ and w/o grid and subtraction. talk about the effect on this.

    \begin{figure}
	   \centering
	   \includegraphics[width=0.45\textwidth]{plots/vacuum_medium_grid}
	   \includegraphics[width=0.45\textwidth]{plots/medium_grid_gridsub}
	   \caption{}
	   \label{fig:tjet}
	\end{figure}

\section{Conclusions}
\label{sec:conc}

\section*{Acknowledgments}
The authors would like to thank Benjamin Nachman and Sevil Salur for helpful discussions. This research was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. RKE also acknowledges support from the National Science Foundation under Grant No.1067907 \& 1352081.

\appendix

Jets are collimated sprays of
particles observed at high energy colliders such as the Large Hadron Collider (LHC). Hard collisions produce energetic quarks and gluons, which shower and hadronize through the strong interaction into the final state particles captured by the detectors. This parton shower picture has been extremely successful for understanding and modeling jets. Physics at all energy scales enters %into
the parton shower and affects the final-state particle distribution. To extract specific features of the jet formation, many
%Many
physically motivated jet substructure observables, defined in terms of the observed particles, have been studied extensively \cite{Abdesselam:2010pt,Altheimer:2012mn,Altheimer:2013yza,Adams:2015hiv,Larkoski:2017jix,ATLAS-CONF-2017-064,Khachatryan:1955546}.

In this Letter, we go beyond \cite{Chien:2017xrb} and present the framework of telescoping deconstruction to systematically probe all aspects of jet formation. By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern. We demonstrate the use of the framework
in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging, and we show that the framework captures all the relevant physics of jets. Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies. Explicit examples of the $W$ isolation \cite{Chien:2017xrb} and exposing the $W$ boson in a top jet are presented. This highlights the physically meaningful nature of each telescoping deconstruction observables and demonstrates their collective power as a representation.

Recently, there has been progress in utilizing the complete information in a jet in an unbiased way %, typically making use of
using advanced machine-learning techniques. Powerful multivariate function approximators, such as neural networks, are capable of extracting useful features of the data relevant for a specific task. Examples include jet images with convolutional neural networks (CNNs)~\cite{Cogan:2014oua,deOliveira:2015xxd,Komiske:2016rsd,Kasieczka:2017nvn}, clustering histories with recurrent neural networks~\cite{Louppe:2017ipp}, and complete sets of high-level observables with dense neural networks (DNNs)~\cite{Datta:2017rhs,Datta:2017lxt,Aguilar-Saavedra:2017rzt}. See \cite{Larkoski:2017jix} for a more complete summary and discussion of recent progress. While machine-learning methods can provide large increases in performance, extracting physical messages from the trained neural network parameters is a complicated task.

Telescoping deconstruction aims to encapsulate all relevant physics information in very simple, physically-meaningful observables. The variables are close to the perturbative expansion of QCD and the parton shower picture therefore perturbative and nonperturbative physics information can be systematically extracted. Much like fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets. %axes.
At each order, %For a given order,
$N$ axes are determined by finding the dominant energy flow directions in the rapidity-azimuth plane~\cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}, which is partitioned into energy flow regions determined by the nearest axis. %Multiple
Jets at multiple angular resolutions are probed simply by the {\sl kinematics} of subjets consisting of %all
particles within different distances $R_T$ from %of
the energy flow axes.

\begin{figure*}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
\caption{\label{fig:tjet}(Top row) The telescoping deconstruction of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_T$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure*}

The telescoping deconstruction procedure is: %substructure deconstruction is:
\begin{itemize}
        \item At order $N$, determine $N$ axes $\{\hat n_i\}=\{(\eta_i,\phi_i)\}$ along the dominant energy flows, where $y$ and $\phi_i$ are the rapidity and the azimuthal angle of the subjet axis $i$. We use the $N$subjettiness 1-pass $k_T$ axes and the Winner-Take-All (WTA) $k_T$ axes \cite{Thaler:2010tr}.
        \item Construct $N$ subjets with $M$ radii $\{R_{T,m}\}^M_{m=1}$ by assigning particles to the nearest axis according to the distance $d^2_{ij} = \Delta y_{ij}^2+\Delta \phi_{ij}^2$ between the axis $\hat n_i$ and the particle $j$ \cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}.
            \begin{equation}
                {\rm subjet}_{i,m} = \{p^\mu_j~|~d_{ij}<R_{T,m}~{\rm and}~d_{ij}<d_{kj} \forall i\neq k\}.
            \end{equation}
            The subjet radii ${R_{T,m}}$ are sampled within the range $(0,R)$ where $R$ is the jet radius. In this paper $\{R_{T,m}\}$ are chosen to be evenly spaced within the range.
        \item We form the subjet data with the subjet transverse momenta and masses $\{(p_T,m)_{i,m}\}$,
            \begin{equation}
                {p_T}_{i,m}=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)_T\;,~~~{m_{i,m}}^2=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)^2\;,
            \end{equation}
            where we sum over all the particles $j$ within the subjet $i,m$. Together with the positions of the axes these form the telescoping deconstruction observables.
    \end{itemize}

The telescoping deconstruction observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses. {\sl Subjet charge}~\cite{Krohn:2012fg} information can straightforwardly be included in this framework as well. As the telescoping subjet order $N$ (T$N$ order) increases, more jet energy is covered by the subjets and the number of subjet radii sampled can be systematically decreased. In the large $N$ limit, the subjets reduce to individual particles %and
thus the basis is formally shown to be complete. The telescoping deconstruction allows one to easily exploit features both within each subjet and among all the subjets. See \Fig{fig:tjet} for the telescoping deconstruction of a top jet for $N = 1$, $2$, and $3$ where the $W$ resonance can clearly be seen.

To demonstrate the efficacy of telescoping deconstruction in capturing the full jet information,
we demonstrate the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging. Each of these problems have
signal jets with a different characteristic number of prongs: one, two, and three prongs, respectively. Events were generated from Monte Carlo simulation of proton-proton collisions at 14 TeV using \textsc{Pythia} 8.226~\cite{Sjostrand:2007gs}. Final-state non-neutrino particles are clustered into jets with \textsc{FastJet} 3.3.0~\cite{Cacciari:2011ma} using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp}. We consider the boosted regime with the jet $p_T$ between 800 GeV and 900 GeV. For quark/gluon discrimination, quark jets were generated by $pp\to q+Z(\to\nu\bar\nu)$ and gluon jets %were generated
by $pp\to g + Z(\to\nu\bar\nu)$, clustered into $R=0.4$ jets with rapidity $|y|<1.5$.

\begin{figure*}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/QG_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/W_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/Top_invROC.pdf}
\caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging. The T$N$ performance approximately saturates at T2 (quark/gluon), T1 ($W$ tagging), and T2 (top tagging) orders.}
\end{figure*}

Since $W$ and top tagging are mass resonance searches where the jet mass is the most natural and powerful discriminating variable, we disentangle mass information to probe how much additional information can be exploited for tagging. Information from the hard process about the overall jet kinematics is eliminated by translating each jet to a frame where $y$ and $\phi$ of the jet are both zero. For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets from QCD dijets. Final-state non-neutrino particles are clustered into $R=1.0$ jets with rapidity $|y|<1.5$. Signal and background jets identically populate a five-bin $1/p_T^4$ histogram in transverse momentum and a three-bin uniform histogram in mass between 75 GeV and 85 GeV for $W$ tagging and 160 GeV to 180 GeV for top tagging. Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in
quark/gluon discrimination, and between 0.05 to 0.95 in
$W$ and top tagging. Reducing the number of radius sampling is left for future work. For each problem, 200k events are generated for signal and background, with 10\% used for validation, 15\% for testing, and the remaining 75\% for training.

A neural network consisting of three dense layers with 100 nodes each is trained on the telescoping deconstruction observables up to T$N$ order. The training can be performed on a typical laptop CPU in fewer than five minutes. One could also use boosted decision trees to combine variables \cite{Chien:2017xrb}. We compare the performance of our method to jet images using a CNN, which requires a graphics processing unit (GPU) to train. The jet images are size $33\times 33$ and span a $2R\times 2R$ patch of the rapidity-azimuth plane with the intensity of each pixel corresponding to the total $p_T$ of particles in the pixel. The jet images are standardized according to the procedure in \Ref{Komiske:2016rsd}. The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer. A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2. The dropout rate was taken to be 0.1 for all layers. All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}. Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used. The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.

The performance of the trained models can be captured in a Receiver Operating Characteristic (ROC) curve which plots the inverse of the background mistag rate at different signal efficiencies, where a higher curve indicates better classification performance. \Fig{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images. The T$N$ performance converges quickly and is comparable to the performance of the jet images approach. % Note that since
The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance. %This is highlighted in the top case, where the telescoping deconstruction DNN outperforms the CNN at T2 order and above.
\begin{figure*}
\centering
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in %the case of
top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure*}

For top tagging, there is a significant increase in performance from T1 to T2 order in \Fig{ROCs}, and the performance saturates beyond this order due to the sensitivity of T2 to the $W$ in top jets. For $W$ tagging, T1 order is sufficient to achieve most of the classification performance, which unambiguously confirms the $W$ isolation feature in the boosted regime compared to the QCD background~\cite{Chien:2017xrb}. Clearly, the T1 order probes the depletion of the radiation at large angles within $W$ jets, whereas the QCD background jets continue to accrue mass from radiation at large angles. For quark/gluon discrimination, there is a significant increase in performance from T1 to T2 and a smaller increase from T2 to T3 where the performance saturates, suggesting the usefulness of T2 subjet substructure and its sensitivity to subjet flavors. This confirms that the T$N$ expansion converges efficiently and telescoping deconstruction faithfully represents the jet information.

In addition to being a complete jet representation, the telescoping deconstruction allows physical information to be easily extracted. \Fig{masses} shows the heavy subjet mass distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background. As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass. In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they %continue to
accrue mass at larger radii. The top jets at T2 order clearly and automatically show the $W$ peak which is completely absent in the background distributions. At T3 order, both top and background distributions appear QCD-like, with the wider background distributions %slightly wider
due to the prevalence of gluon subjets. This highlights again subjet substructure in telescoping deconstruction.

To conclude, we present the telescoping deconstruction framework as a complete, physically-motivated expansion of the jet.
%Neural networks were trained on the telescoping deconstruction observables up to different T$N$ orders and applied to quark/gluon discrimination, boosted $W$ and top tagging.
Telescoping deconstruction observables up to different T$N$ orders are combined using multivariate methods and applied to quark/gluon discrimination, boosted $W$ and top tagging.
%The performance of the models indicates that the
The T$N$ performance converges efficiently and compares favorably to a CNN trained on jet images. Further, the individual telescoping deconstruction observables were shown to cleanly encode relevant features of jets such as the isolation of $W$ jets, the presence of the $W$ in top jets, and the quark- or gluon- like nature of the heaviest prongs in top and QCD jets. This complete and general framework is suitable for systematic studies of full jet dynamics, with rich future applications in parton-shower tuning, heavy ion hard probe studies \cite{Chien:2017hij}, and identification of hadronic electroweak-boson emissions in ultra-high energy jets. The extension to hard-process tagging and resonance search at the event level is also promising.

\bibliographystyle{JHEP3}
\bibliography{qg_ML_ref}

\end{document}

### 2018-01-28 qg_ML_4.tex
YT updated the draft

\documentclass[notoc]{JHEP3}
\usepackage{cite,color,psfrag,url,latexsym,graphicx,epstopdf,slashed,xspace,hyperref,enumitem,lineno}
\let\normalcolor\relax

\newcommand{\nn}{\nonumber \\}
\newcommand{\pt}{$p_{T}$}
\newcommand{\eqref}[1]{(\ref{#1})}
\newcommand{\text}[1]{{\rm #1}}
\newcommand{\jw}{\textsc{Jewel}}
\newcommand{\jwpy}{\textsc{Jewel+Pythia}}
\newcommand{\py}{\textsc{Pythia}~}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\def\PlusBreak#1{+ \nonumber \\          % splits line with plus sign
          &&  \hphantom{#1} \! \null   +}
\def\MinusBreak#1{- \nonumber \\         % splits line with minus sign
          &&  \hphantom{#1} \!  \null  -}
\def\TimesBreak#1{\times \nonumber \\    % splits lines with times sign
          &&  \hphantom{#1} \!  \null  \times}

% The following macros allow line splits allowing for line spacing adjustments
% 12 is line spacing e.g. 1pt plus 4pt and #2 is the hphantom
\def\PlusBreakAdjust#1#2{+ \nonumber \\ [#1]      % splits line with plus sign
          &&  \hphantom{#2} \! \null  +}
\def\MinusBreakAdjust#1#2{- \nonumber \\ [#1]     % splits line with minus sign
          &&  \hphantom{#2} \!  \null -}
\def\TimesBreakAdjust#1#2{\times \nonumber \\[#1] %splits lines with times sign
          &&  \hphantom{#2} \!  \null  \times}

\def\eqn#1{Eq.~(\ref{eq:#1})}
\def\Eqn#1{Eq.~(\ref{eq:#1})}
\def\eqns#1#2{Eqs.~(\ref{eq:#1}) and~(\ref{eq:#2})}
\def\Eqns#1#2{Eqs.~(\ref{eq:#1}) and~(\ref{eq:#2})}
\def\fig#1{Figure~{\ref{fig:#1}}}
%\renewcommand{\sec}[1]{Section~\ref{sec:#1}}
\newcommand{\ssec}[1]{Section~\ref{ssec:#1}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\cmb}{\begin{changemargin}}
\newcommand{\cme}{\end{changemargin}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
% \newcommand{\bea}{\begin{align}}
% \newcommand{\eea}{\end{align}}

\DeclareRobustCommand{\Sec}[1]{Sec.~\ref{#1}}
\DeclareRobustCommand{\Secs}[2]{Secs.~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Secss}[3]{Secs.~\ref{#1}, \ref{#2} and \ref{#3}}
\DeclareRobustCommand{\App}[1]{App.~\ref{#1}}
\DeclareRobustCommand{\Tab}[1]{Table~\ref{#1}}
\DeclareRobustCommand{\Tabs}[2]{Tables~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Fig}[1]{Fig.~\ref{#1}}
\DeclareRobustCommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\DeclareRobustCommand{\Figss}[3]{Figs.~\ref{#1}, \ref{#2} and \ref{#3}}
\DeclareRobustCommand{\Eq}[1]{Eq.~(\ref{#1})}
\DeclareRobustCommand{\Eqs}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
\DeclareRobustCommand{\Eqss}[3]{Eqs.~(\ref{#1}), (\ref{#2}), and (\ref{#3})}
\DeclareRobustCommand{\Ref}[1]{Ref.~\cite{#1}}
\DeclareRobustCommand{\Refs}[2]{Refs.~\cite{#1} and \cite{#2}}

\title{Probing heavy ion collisions using quark and gluon jet substructure}

\author{Yang-Ting Chien $^{a}$ and Raghav Kunnawalkam Elayavalli $^{b}$\\
$^{a}$ Center for Theoretical Physics\\
$~$ Massachusetts Institute of Technology, Cambridge, MA 02139\\
$^{b}$ Department of Physics and Astronomy\\
$~$ Rutgers, the State University of New Jersey, New Brunswick, NJ 08901
}

\abstract{
We introduce a novel jet quenching study framework utilizing both quark-initiated jets and gluon-initiated jets as independent probes of heavy ion collisions. We perform quark gluon discrimination in proton-proton and heavy ion collisions using jet substructure techniques to exploit all the jet features. Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are collectively examined in a multivariate analysis. In comparison, we perform the classification using machine-learning methods by training a deep convolutional neutral network on discretized jet images. Furthermore, we develop the telescoping deconstruction framework with a systematic subjet expansion to exploit the information carried by the subjets at multiple angular scales. We draw connections to the soft-drop subjet distribution and illuminate modifications of quark jets and gluon jets using Lund diagrams. The quark and gluon jet samples generated from \jwpy ~are used as an example to demonstrate our general method. We find that the classification performance goes down in the \jwpy-simulated heavy ion collisions. The information carried by the subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation. Our method provides a systematically improvable framework for analyzing and comparing jet simulations and measurements in heavy ion collisions.
}

\begin{document}

\section{Introduction}
\label{sec:intro}

The jet quenching phenomenon observed at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
and the Large Hadron Collider (LHC)\cite{Aamodt:2010jd,Abelev:2012hxa,Abelev:2013kqa,Aad:2012vca,Aad:2014bxa,Chatrchyan:2013kwa,
Chatrchyan:2012gw,Chatrchyan:2014ava,Aad:2014wha,Chatrchyan:2013exa,
Adam:2015ewa,Chatrchyan:2012gt,Chatrchyan:2011sx,Chatrchyan:2012nia,Aad:2010bu,Aad:2013sla,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha} has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions. The dramatic suppression of hadron and jet cross sections was understood using the energy loss picture. By studying in details how the jet energy redistributes in heavy ion collisions, one can quantitatively extract the information about the jet-medium interaction and the underlying medium dynamics. The heavy ion jet physics has thus moved to the era of precision jet substructure and jet cross section studies.

Jet substructure measurements provide concrete information about the modifications of jets. From the measurements of the jet shape \cite{} and jet fragmentation function \cite{}, we have learned that both the transverse and longitudinal momentum distributions inside jets are significantly modified. Since jet substructures depend strongly on the partonic origin of jets, the change of the fraction of quark-initiated jets and gluon-initiated jets will contribute to the substructure modifications. It has been realized that the increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of the jet cores seen in various inclusive jet substructure measurements. With purer quark and gluon jet samples, one can disentangle the effect from the changing quark and gluon jet fractions and focus on the jet-by-jet modifications. Quark jets and gluon jets can then be used as independent probes.

We will study the tagging of quark and gluon jets making full use of their substructure information. In contrast to the conventional wisdom of directly comparing jets in proton-proton and heavy ion collisions, this indirect approach studies the modification of the comparison between different probes. By first comparing the probes in the same system, it may help reducing the systematic uncertainty in experimental measurements. The lowest order feature which separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons. The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets. Various jet substructure variables have been used and combined to further increase the tagging performance with a multivariate analysis, and the advances of machine learning techniques have been constantly incorporated in solving this classification problem.

In this paper we benchmark three classes of methods in quark and gluon tagging and gain insights from comparing their performances. We use modern image recognition techniques utilizing deep convolutional neural networks (DCNN). The network is trained on the discretized images of quark jets and gluon jets, and the samples are generated from the \jw heavy ion event generator. Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments. We also examine the tagging performance using five physically constructed jet substructure variables: the jet mass, the $p_T^D$, the multiplicity and two radial moments including the girth. Subsets of the variables are combined and studied in a multivariate analysis using boosted decision trees, and we look into the correlations among the variables. Lastly, we systematically expand the information within jets using the recently developed telescoping deconstruction method. The corresponding subjet expansion can be truncated at any finite order $N$ which allows us to access the information carried by the subleading subjets.

The rest of the paper is organized as follows. In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy ~simulation. In Sec.~\ref{sec:qg} we give the details about the multivariate analysis, the jet image, and the telescoping deconstruction methods. In Sec.~\ref{sec:results} we show the performance of each method and the distributions of jet substructure observables. In Sec.~\ref{sec:conc} we conclude and give an outlook of future studies.

\section{Quark and Gluon Jet Samples}
\label{sec:sample}

To use quark jets and gluon jets as independent probes of heavy ion collisions, we first need to identify the probes we are using. In this section we describe the quark and gluon jet samples we use to demonstrate our general jet quenching study method. As an example, the jet samples are generated from the \jwpy ~simulations of proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=13$ TeV using the standard setup \cite{Zapp:2013zya}. We use the prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ for constructing the quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions. It is possible to construct the jet samples using a completely physical way by performing hard process and kinematical selections to enhance either the quark or the gluon jet fraction. This can then be directly applied to analyze other jet simulations and experimental data in the future.

In each collision event, jets are reconstructed using the anti-$k_{t}$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$. We impose the following kinematic cuts,
\begin{equation}
    p^{\gamma}_T > 100~{\rm GeV}, ~~~~ p^{\rm jet}_T > 50~{\rm GeV}, ~~~~ |\eta^{\rm jet}|<2, ~~~~\Delta \phi_{\gamma, {\rm jet}} > 2\pi/3\;,
\end{equation}
where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon. \jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions. The medium-induced energy loss in heavy ion collisions is calculated based on a medium model consisting of thermal scattering centers undergoing the Bjorken expansion. The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}. We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHApdf}~\cite{Whalley:2005nh}. We include recoil partons which are scattering centers interacting with the propagating hard scattered partons in the \jw~simulations. It was shown that recoil contributions are important to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. On the other hand, a large fraction of the jet energy comes from the huge underlying event background in heavy ion collisions which needs to be subtracted before one could study the intrinsic modification of jets. We adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo}. The \jw ~output HepMC files are analyzed using the \textsc{Rivet} analysis framework~\cite{}.

%For a detailed description of the MC implementation of the jet production and its elastic/inelastic interactions with the scattering centers along with the LPM effect in \jw, please refer~\cite{}. The primary event collection was split into two halves, with one utilized for training the model(s) and the other for testing \textcolor{red}{additional to a di-jet sample, with a mixture of quark and gluon jets utilized for further evaluation of the model performance}. The question of what constitutes a "quark" or "gluon" jet is a topic of current discussion~\cite{} but for this paper, we choose the conventional definition of a quark/gluon jet with its progenitor hard scattered parton. There are two operational modes of generating \jw events in heavy ion collisions; with and without including the recoil partons, those originating from interactions of the propagating hard scattered parton with the scattering centers. In recent studies, \jw with recoils was shown to accurately reproduce several of the important qualitative features in jet structure modifications measured by LHC experiments with the addition of background subtraction techniques introduced here~\cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}. The jet-background in \jw consists of the scattering center's thermal momenta i.e, their momentum before the interaction. In studying these different jet collections, we extract the dependence of the tagging on key operations performed on jets in data, primarily discretization and background subtraction and discuss their impact on the physics.

With finite detector angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC expeirements. To study how such discretization as well as background subtraction affect the analysis, we use the following three different sets of jets in heavy ion collisions, all with the medium recoil contributions,
\begin{itemize}
    \item neither discretization nor background subtraction
	\item with discretization but without background subtraction
	\item with discretization and the GridSub background subtraction.
\end{itemize}
For jets in proton-proton collisions, we use the sets of jet samples with or without discretization. In reality, the underlying event background subtraction for jet substructure observables is highly challenging and nontrivial, which can introduce significant smearing effects. To disentangle such an issue, one common practice adopted by the CMS collaboration is to smear jets in pp collisions using minimum bias heavy ion events before comparing with jets in heavy ion collisions. In this work we do not address this issue with the \jwpy-simulated jet samples and assume that the built-in background subtraction procedure allow a fair comparison between jets in pp and PbPb collisions.

\section{Quark and Gluon Jet Substructure}
\label{sec:qg}
\begin{figure}
	   \centering
	   \includegraphics[width=0.6\textwidth]{plots/qg_HI}
	   \caption{Illustration of the interplay between jet quenching (vertical) and quark gluon discrimination (horizontal). }
	   \label{fig:quenching_discrimination}
\end{figure}

We will use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns. In doing so one attempts to identify jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms. This can be quantified by studying the modification of specific jet substructure observables. Such direct approach is closely related to the outstanding problem of discriminating quark and gluon jets, where the goal is to exploit jet features which show differences between quark jets and gluon jets. One may compare quark jets and gluon jets in the same collision system and use their differences as a way to quantify the jet modification. This can help disentangle jet quenching effects on common features between quark and gluon jets and cancel systematic uncertainties in experiments. A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination}.

Quark gluon discrimination has been studied extensively in the past\cite{}. With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful ones \cite{Gallicchio:2011xq,Gallicchio:2012ez}. On the other hand, using modern machine learning image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance can be significantly increased. This suggests that additional jet features have been identified by the deep neural network to help the tagging task. We use both methods to gain insights on how jet quenching is manifested in these two jet representations. Furthermore, we develop the telescoping deconstruction framework which represents the whole jet information using subjets with different angular resolution. It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet features. The details about telescoping deconstruction are provided in Appendix.

\subsection{Physics-motivated Multivariate Analysis}
\label{sec:mva}

\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
	   \caption{Distributions of the jet constituents (top left), invariant jet mass (top right), $p^{D}_{T}$ (center left), the half radial moment (center right) and the jet girth or the first radial moment (bottom) for quark (darker blue) and gluon (lighter orange) induced jets (color online). The top panels of each individual distribution show the pp and the bottom panels correspond to central (0-20\%) PbPb events generated by the \jw generator.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

Multivariate analysis have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus an effective, physics-motivated baseline in our study. Given a jet reconstructed using the anti-$k_{t}$ algorithm \cite{Cacciari:2008gp}, we choose the following set of five jet substructure observables \cite{} to highlight the differences in quark jets and gluon jets,
	\begin{itemize}
		\item Jet mass: $m = \sqrt{(\sum_{i\in {\rm jet}} p_i)^2}$. 
        \item First radial moment (girth): $\sum_{i \in {\rm jet}} p^i_{T} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p^i_{T} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
        \item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^{i}_{T}}^2}/p^{\rm jet}_{T}$.
        \item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
	\end{itemize}

Fig:~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the above jet observables in pp and PbPb collisions. The quark and gluon jets used in plotting Fig:~\ref{fig:jetdistributons_pp_pbpb} include the grid dicretization. Furthermore, jets in PbPb collisions are subtracted using the GridSub background subtraction method. Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions. In vacuum, they reflect the fact that gluon jets have larger jet mass and radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$. In the \jwpy-simulated PbPb collisions, for both quark and gluon jets the distributions are consistently modified towards larger values. On the other hand, $p_T^D$ and the pixel multiplicity are IRC unsafe. We see that there is significant increase of the pixel multiplicity and decrease of the values of $p_T^D$. This is due to the presence of the increased number of soft particles inside jets from the \jw ~medium recoil contributions \cite{}.

It is important to note that these observables are sensitive to the jet clustering algorithm, the minimum \pt ~cutoff of jet constituents and the background subtraction procedures employed. In order to facilitate a direct comparison between experimental data and Monte Carlo simulations, one has to either unfold the detector resolution from these observables or perform Monte Carlo simulations with detector response. The unfolding procedure for jet observables increases in dimensionality and complexity as the correlations among observables become strong. For example, the jet mass will have to be unfolded in the 4-dimensional distribution of the generated and reconstructed jet \pt ~and mass.

We combine the five jet substructure observables using a multivariate model implemented in Keras, with two dense layers of 10 nodes each. We use the rectified linear unit (\textit{relu}) and sigmoid activation functions for the dense layers and the output layer, respectively. The model training utilizes the Adam optimizer and the binary cross entropy loss function. We perform cross validation using Monte Carlo samples which are split into two halves for training and testing.

%All five jet observables showcase the ability to classify quarks vs gluons in both pp and PbPb. Recent results from LHC experiments on the effect of jet quenching on inclusive jets showcase two main features; a slight narrowing of the jet core and increased low \pt particle multiplicity in the periphery regions of the jet leading to a broadening of the jet shape. Such an effect is captured in \jw, with the introduction of the medium induced recoils and introduces a further complexity in the question of quark vs gluon classification in PbPb. In vacuum, we expect gluon jets to be broader than quark jets due to the different in the casimir color factor and thus in PbPb collisions, with both jets being quenched and broader, the ability to discriminate effectively between quark and gluon jets is expected to reduce when one only considers fragmentation based observables as we proceed to show in the upcoming sections.

\subsection{Jet Image}
\label{sec:image}

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgGluonJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgQuarkJet}
\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
\caption{Average quark (left) and gluon (right) pre-processed jet images shown in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \jw. The color scale in the z axis represents the average pixel \pt in units of GeV/c. }
\label{fig:qgjetimages}
\end{figure}

The multivariate analysis discussed previously is a constructive approach to collect and examine the usefulness of jet features which may help distinguish quark jets from gluon jets. An alternative approach is to use image recognition techniques which optimize comprehensive multivariate models and identify all possible, useful features. The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane of transverse momentum deposit in detector calorimeter cells, much like a digital camera. The jet image thus represents a fixed-dimensional representation of jets. Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in $\eta$ and $\phi$ plane so that the highest pixel centers at the origin. Also, the image is rotated so that the principal component of the energy density is along the $\eta=0, \phi<0$ direction.

Fig:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions. The color of each image pixel represents the transverse momentum deposited in the pixel. We see the broader energy distribution around the center pixel for gluon jets compared to quark jets in pp collisions, and the significantly broader distributions for jets in PbPb collisions. Note that in experiment it may be very challenging to directly analyze jet images due to detector and measurement considerations. Here we mainly use the jet image approach as a useful comparison of the classification performance.

A deep convolutional neural network implemented in Keras with the TensorFlow backend is utilized in this analysis. The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer. The convolution layers use hyperbolic tangent activation function, and the output layer uses sigmoid activation function. The output layer is preceded by three layers, a dropout layer with a rejection score of 0.20, a dense layer with 20 nodes and an additional dropout layer with a reduced rejection score of 0.10. The additional dropout layer helps filtering less important features from previous layers and thus increases the speed and efficiency of the model training. The model is trained using the Adam optimizer with binary cross entropy loss function.

\subsection{Telescoping Deconstruction}
\label{sec:tjet}

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/QCDjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/QCDjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/QCDjetT3_telescoped2.pdf}
\caption{Illustration of telescoping deconstruction at T1 (left panel), T2 (middle panel) and T3 (right panel) orders for a QCD jet. The crosses are the Winner-Take-All $k_T$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. See Appendix for more details about the procedures.}
\label{fig:tjetQCD}
\end{figure}

The multivariate analysis and jet image recognition discussed previously are two characteristically different methods. The multivariate analysis uses physics-motivated observables and is a ``bottom-up" approach to collect useful jet features. However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve the method. On the other hand, jet images represent low-level, comprehensive jet information. The training of a deep neural network is a ``top-down" approach to identify all the useful features through the model optimization. However, extracting physical messages from the trained neural network parameters can be a challenging task.

We develop and apply the telescoping deconstruction (TD) framework in our study of quark and gluon jet substructures in pp and PbPb collisions. Telescoping deconstruction aims to embrace both advantages of multivariate analysis and jet image recognition. It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized basis of jet substructures. It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables. The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically-meaningful which encode the hard splitting of jets and non-perturbative physics. The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions. It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction and then reaching out to include wide-angle, soft radiation. Fig:~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders. The details about the telescoping deconstruction procedures are provided in Appendix, including the applications on boosted $W$ and top tagging in high energy physics.

For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with steps of 0.08. The subjet observables up to the $N$-th order are combined using dense neural networks to maximize the classification performance. We consider $N=\{1,2,3\}$ and the performance quickly saturates at $N=3$. {\bf add some ML details}

Beside the overall quark gluon classification performance using telescoping deconstruction, subjet variables within this framework can reveal many aspects of jet dynamics. In particular, there is rich information contained in subjet topology and subjet substructure. Recently, the CMS \cite{} and STAR \cite{} collaborations measured the groomed momentum sharing observable $z_g$. One reclusters jets using the Cambridge/Aachen algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and sequentially removes wide angle, soft radiation until the following condition is satisfied by a branching,
\begin{equation}
    z_{\rm cut} < \frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}} \equiv z_g\;.
    \label{SD}
\end{equation}
Here $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two branches, and $z_{\rm cut}$ is the soft-drop parameter which is set to $0.1$ in our analysis. Note that the Cambridge/Aachen clustering tree is angular-ordered, therefore the above procedure identifies the most wide angle, soft subjet which carries a significant fraction of the jet transverse momentum. The size of the soft subjet is dynamically determined and related to the angle $r_g$ between the two soft-drop branches, defined as the groomed jet radius. The observables $z_g$ and $r_g$ thus encode the kinematic information of the two subjets which is in the category of subjet topology. In vacuum, the observable $z_g$ is closely related to the Altarelli-Parisi splitting functions~\cite{Altarelli:1977zs}. The physical meaning of $z_g$ in heavy ion collisions have also been investigated \cite{Chien:2016led}.

We use telescoping deconstruction to analyze the two soft-drop branches defining the $z_g$ observable. At the T2 order, we choose the subjet radius to be half the angle between the two deconstruction axes for subjet reconstruction. The distribution of the momentum fraction $z_{\rm T2}$ of the softer subjet is significantly different from the $z_g$ distribution, especially for gluon jets. Note that the axes used in the telescoping deconstruction procedure are defined with the Winner-Take-All scheme which favors energetic particles. We find that the axes at the T2 order may both align with the hard branch and are not able to efficiently tag the soft branch when it consists of soft particles. The situation is significantly improved at the T3 order where we use three axes to capture the soft branch in soft-drop. We combine two closest subjets, reducing the number of subjets to two, and the distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark and gluon jets are very similar. In comparison, we reconstruct subjets around the axes of the soft-drop branches to probe their energy distributions. We find that the distribution of the momentum fraction $z_{{\rm T2},g}$ of the softer subjet is very similar to that of $z_{\rm T3}$ and different from that of $z_g$. This suggests that the active areas of the two soft-drop branches are different. On the other hand, the subjet substructure may allow us to extract the information about the partonic origin of subjets. We look at the scaleless ratio $m_i/p_{T_i}$ where $m_i$ is the light (heavy) subjet mass, or the mass of the hard (soft) subjet.

\section{Quark and Gluon Jet Modification and Discrimination}
\label{sec:results}

We discussed jet representations using multiple physics-motivated variables, jet images and telescoping deconstruction. In this section we will use these information in the study of jet modification and quark gluon discrimination. For quark gluon discrimination, the performances are shown using Receiver Operating Characteristic (ROC) curves which plot the gluon jet efficiency as a function of the quark jet efficiency. We also perform discriminations between jets in pp and AA collisions for both quark jets and gluon jets, and the performance is evaluated by the ROC curves showing AA jet efficiency as a function of the pp jet efficiency. A lower ROC curve represents a better performance.

\begin{figure}
	   \centering
	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_pp_2p76TeV_genLevel_QvsG.pdf}
	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
	   \caption{}
\label{fig:ROC_all}
\end{figure}

Fig:~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and AA (right panel) collisions, using particle momenta when evaluating the observables without $\eta=\phi$ discretization and background subtraction. We see that overall the methods give similar performances, implying that the three jet representations capture sufficient jet information. Note that there is monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$ to $\rm T1+T2+T3$ with increasing input information of higher-order subleading subjets, and the performance saturates quickly at the T2 order. In the AA case, combining physics variables gives better performance due to the high particle multiplicity in the \jwpy-simulated AA events. Telescoping deconstruction with radius scan in steps of 0.08 is not able to capture the full multiplicity information at the T3 order.

\begin{figure}
	   \centering
       \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
       \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
	   \caption{}
\label{fig:ROC_TD_input}
\end{figure}

By comparing the performances in pp and AA collisions, we clearly see that the performances drops in AA collisions. To have a closer look at the change of the classification performance from pp to AA collisions and the convergence of the TD performance, we plot the ROC curves of telescoping deconstruction in both pp and AA collisions in the left panel of Fig:~\ref{fig:ROC_TD_input}. We see the shrink of the performance improvement from T1 to $\rm T1+T2$ in AA collisions. This suggests that the information carried in higher-order subleading subjets can be washed out in heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions in \jwpy. In the right panel of Fig:~\ref{fig:ROC_TD_input} we plot the TD ROC curves with different inputs with or without $\eta-\phi$ discretization and background subtraction. Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance. On the other hand, the performance drops with the background subtraction performed. This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction.

\begin{figure}
	   \centering
	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
	   \caption{}
\label{fig:ROC_qq_gg}
\end{figure}
Fig:~\ref{fig:ROC_all} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and AA collisions using physics-motivated variables, jet images and telescoping deconstruction. We see huge difference between jets in pp and AA collisions therefore we plot the ROC curve using log scale in the y-axis. The significantly better performance in the gluon jet discrimination suggests that the modification of gluon jets is larger. We observe that the increase of the pixel multiplicity is a key feature in jet modification.

\section{Quark and Gluon Jets in Lund Diagrams}
\label{sec:lund}

\begin{figure}
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
	   \caption{}
\label{fig:Lund_full}
\end{figure}
Fig:~\ref{fig:Lund_full} shows the Lund diagrams for quark jets (left) and gluon jets (right) in pp (top) and AA (bottom) collisions. Note that only the splitting in the soft-drop hard branch enter into the Lund diagrams. We see the significant contributions from soft, wide-angle radiation in AA collisions.

\begin{figure}
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
	   \caption{}
\label{fig:Lund_bkg}
\end{figure}
Fig:~\ref{fig:Lund_bkg} shows the Lund diagrams for C/A trees before the one which passes the soft-drop condition. We see that the soft, wide-angle radiation is isolated.

\begin{figure}
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
	   \caption{}
\label{fig:Lund_hard}
\end{figure}
Fig:~\ref{fig:Lund_hard} shows the Lund diagrams for C/A trees after the hard branch in soft-drop. We see that the vacuum structure starts to emerge in the soft-drop AA jets. However, soft-drop can not remove all the soft radiation within the hard branch in AA collisions.

\begin{figure}
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_zgrg.pdf}
	   \caption{}
\label{fig:Lund_zgrg}
\end{figure}

\section{Conclusions}
\label{sec:conc}

\section*{Acknowledgments}
The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributing to Appendix. The authors also thank Benjamin Nachman and Sevil Salur for helpful discussions, and . Y.-T. Chien was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008. R.K. Elayavalli acknowledges support from the National Science Foundation under Grant No.1067907 \& 1352081 (+JETSCAPE?).

\newpage
\appendix
\section*{Appendix}
\section{Probing subjet energy flows with telescoping deconstruction}

Jets are collimated sprays of particles observed at high energy colliders such as the Large Hadron Collider (LHC). Hard collisions produce energetic quarks and gluons, which shower and hadronize through the strong interaction into the final state particles captured by the detectors. This parton shower picture has been extremely successful for understanding and modeling jets. Physics at all energy scales enters the parton shower and affects the final-state particle distribution. To extract specific features of the jet formation, many physically motivated jet substructure observables, defined in terms of the observed particles, have been studied extensively \cite{Abdesselam:2010pt,Altheimer:2012mn,Altheimer:2013yza,Adams:2015hiv,Larkoski:2017jix,ATLAS-CONF-2017-064,Khachatryan:1955546}.

In this appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation. By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern. We demonstrate the use of the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging, and we show that the framework captures relevant physics of jets. Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies. Explicit examples of the $W$ isolation \cite{Chien:2017xrb} and exposing the $W$ boson in a top jet are presented. This highlights the physically meaningful nature of each telescoping deconstruction observables and demonstrates their collective power as a representation.

Recently, there has been progress in utilizing the complete information in a jet in an unbiased way using advanced machine-learning techniques. Powerful multivariate function approximators, such as neural networks, are capable of extracting useful features of the data relevant for a specific task. Examples include jet images with convolutional neural networks (CNNs)~\cite{Cogan:2014oua,deOliveira:2015xxd,Komiske:2016rsd,Kasieczka:2017nvn}, clustering histories with recurrent neural networks~\cite{Louppe:2017ipp}, complete sets of high-level observables with dense neural networks (DNNs)~\cite{Datta:2017rhs,Datta:2017lxt,Aguilar-Saavedra:2017rzt} and linear basis of energy flow polynomials \cite{Komiske:2017aww}. See \cite{Larkoski:2017jix} for a more complete summary and discussion of recent progress. %While machine-learning methods can provide large increases in performance, extracting physical messages from the trained neural network parameters is a complicated task.

Telescoping deconstruction aims to encapsulate relevant physics information in simple, physical observables. The variables are close to the perturbative expansion of QCD and the parton shower picture therefore perturbative and nonperturbative physics information can be systematically extracted. Similar to fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets. At each order, $N$ axes are determined by finding the dominant energy flow directions in the rapidity-azimuth plane~\cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}, which is partitioned into energy flow regions determined by the nearest axis. Jets at multiple angular resolutions are probed simply by the {\sl kinematics} of subjets consisting of particles within different distances $R_T$ from the energy flow axes.

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
\caption{\label{fig:tjet}(Top row) The telescoping deconstruction of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_T$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure}

The telescoping deconstruction procedure is:
\begin{itemize}
        \item At order $N$, determine $N$ axes $\{\hat n_i\}=\{(\eta_i,\phi_i)\}$ along the dominant energy flows, where $y$ and $\phi_i$ are the rapidity and the azimuthal angle of the subjet axis $i$. We use the Winner-Take-All (WTA) $k_T$ axes \cite{Thaler:2010tr}.
        \item Construct $N$ subjets with $M$ radii $\{R_{T,m}\}^M_{m=1}$ by assigning particles to the nearest axis according to the distance $d^2_{ij} = \Delta y_{ij}^2+\Delta \phi_{ij}^2$ between the axis $\hat n_i$ and the particle $j$ \cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}.
            \begin{equation}
                {\rm subjet}_{i,m} = \{p^\mu_j~|~d_{ij}<R_{T,m}~{\rm and}~d_{ij}<d_{kj} \forall i\neq k\}.
            \end{equation}
            The subjet radii ${R_{T,m}}$ are sampled within the range $(0,R)$ where $R$ is the jet radius. In this paper $\{R_{T,m}\}$ are chosen to be evenly spaced within the range.
        \item We form the subjet data with the subjet transverse momenta and masses $\{(p_T,m)_{i,m}\}$,
            \begin{equation}
                {p_T}_{i,m}=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)_T\;,~~~{m_{i,m}}^2=\Big(\sum_{j\in~{\rm subjet}_{i,m}}p^\mu_j\Big)^2\;,
            \end{equation}
            where we sum over all the particles $j$ within the subjet $i,m$. Together with the positions of the axes these form the telescoping deconstruction observables.
    \end{itemize}

The telescoping deconstruction observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses. {\sl Subjet charge}~\cite{Krohn:2012fg} information can be included in this framework as well. As the telescoping subjet order $N$ (T$N$ order) increases, more jet energy is covered by the subjets and the number of subjet radii sampled can be systematically decreased. In the large $N$ limit, the subjets reduce to individual particles with the full jet information. The telescoping deconstruction allows one to exploit features both within each subjet and among all the subjets. See \Fig{fig:tjet} for the telescoping deconstruction of a top jet for $N = 1$, $2$, and $3$ where the $W$ resonance can clearly be seen.

To demonstrate the efficacy of telescoping deconstruction in capturing the full jet information,
we apply the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging. Each of these problems have
signal jets with a different characteristic number of prongs: one, two, and three prongs, respectively. Events were generated from Monte Carlo simulation of proton-proton collisions at 14 TeV using \textsc{Pythia} 8.226~\cite{Sjostrand:2007gs}. Final-state non-neutrino particles are clustered into jets with \textsc{FastJet} 3.3.0~\cite{Cacciari:2011ma} using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp}. We consider the boosted regime with the jet $p_T$ between 800 GeV and 900 GeV. For quark/gluon discrimination, quark jets were generated by $pp\to q+Z(\to\nu\bar\nu)$ and gluon jets %were generated
by $pp\to g + Z(\to\nu\bar\nu)$, clustered into $R=0.4$ jets with rapidity $|y|<1.5$.

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/QG_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/W_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/Top_invROC.pdf}
\caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging. The T$N$ performance approximately saturates at T2 (quark/gluon), T1 ($W$ tagging), and T2 (top tagging) orders.}
\end{figure}

Since $W$ and top tagging are mass resonance searches where the jet mass is the most natural and powerful discriminating variable, we disentangle mass information to probe how much additional information can be exploited for tagging. Information from the hard process about the overall jet kinematics is eliminated by translating each jet to a frame where $y$ and $\phi$ of the jet are both zero. For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets from QCD dijets. Final-state non-neutrino particles are clustered into $R=1.0$ jets with rapidity $|y|<1.5$. Signal and background jets identically populate a five-bin $1/p_T^4$ histogram in transverse momentum and a three-bin uniform histogram in mass between 75 GeV and 85 GeV for $W$ tagging and 160 GeV to 180 GeV for top tagging. Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in
quark/gluon discrimination, and between 0.05 to 0.95 in
$W$ and top tagging. Reducing the number of radius sampling is left for future work. For each problem, 200k events are generated for signal and background, with 10\% used for validation, 15\% for testing, and the remaining 75\% for training.

A neural network consisting of three dense layers with 100 nodes each is trained on the telescoping deconstruction observables up to T$N$ order. The training can be performed on a typical laptop CPU in fewer than five minutes. One could also use boosted decision trees to combine variables \cite{Chien:2017xrb}. We compare the performance of our method to jet images using a CNN, which requires a graphics processing unit (GPU) to train. The jet images are size $33\times 33$ and span a $2R\times 2R$ patch of the rapidity-azimuth plane with the intensity of each pixel corresponding to the total $p_T$ of particles in the pixel. The jet images are standardized according to the procedure in \Ref{Komiske:2016rsd}. The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer. A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2. The dropout rate was taken to be 0.1 for all layers. All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}. Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used. The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.

The performance of the trained models can be captured in a Receiver Operating Characteristic (ROC) curve which plots the inverse of the background mistag rate at different signal efficiencies, where a higher curve indicates better classification performance. \Fig{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images. The T$N$ performance converges quickly and is comparable to the performance of the jet images approach. % Note that since
The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance. %This is highlighted in the top case, where the telescoping deconstruction DNN outperforms the CNN at T2 order and above.
\begin{figure}
\centering
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in %the case of
top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure}

For top tagging, there is a significant increase in performance from T1 to T2 order in \Fig{ROCs}, and the performance saturates beyond this order due to the sensitivity of T2 to the $W$ in top jets. For $W$ tagging, T1 order is sufficient to achieve most of the classification performance, which unambiguously confirms the $W$ isolation feature in the boosted regime compared to the QCD background~\cite{Chien:2017xrb}. Clearly, the T1 order probes the depletion of the radiation at large angles within $W$ jets, whereas the QCD background jets continue to accrue mass from radiation at large angles. For quark/gluon discrimination, there is a significant increase in performance from T1 to T2 and a smaller increase from T2 to T3 where the performance saturates, suggesting the usefulness of T2 subjet substructure and its sensitivity to subjet flavors. This confirms that the T$N$ expansion converges efficiently and telescoping deconstruction faithfully represents the jet information.

In addition to being a useful jet representation, the telescoping deconstruction allows physical information to be easily extracted. \Fig{masses} shows the heavy subjet mass distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background. As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass. In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they %continue to
accrue mass at larger radii. The top jets at T2 order clearly and automatically show the $W$ peak which is completely absent in the background distributions. At T3 order, both top and background distributions appear QCD-like, with the wider background distributions %slightly wider
due to the prevalence of gluon subjets. This highlights again subjet substructure in telescoping deconstruction.

\bibliographystyle{JHEP3}
\bibliography{qg_ML_ref}

\end{document}

%We present the telescoping deconstruction framework as a complete, physically-motivated expansion of the jet. Neural networks were trained on the telescoping deconstruction observables up to different T$N$ orders and applied to quark/gluon discrimination, boosted $W$ and top tagging. Telescoping deconstruction observables up to different T$N$ orders are combined using multivariate methods and applied to quark/gluon discrimination, boosted $W$ and top tagging. The T$N$ performance converges efficiently and compares favorably to a CNN trained on jet images. Further, the individual telescoping deconstruction observables were shown to cleanly encode relevant features of jets such as the isolation of $W$ jets, the presence of the $W$ in top jets, and the quark- or gluon- like nature of the heaviest prongs in top and QCD jets. This complete and general framework is suitable for systematic studies of full jet dynamics, with rich future applications in parton-shower tuning, heavy ion hard probe studies, and identification of hadronic electroweak-boson emissions in ultra-high energy jets. The extension to hard-process tagging and resonance search at the event level is also promising.

    %layer = Convolution2D(8, 11, 11, border_mode='same')(input_layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(2,2))(layer)
    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(3,3))(layer)
    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
    %layer = Activation('tanh')(layer)
    %layer = MaxPooling2D(pool_size=(3,3))(layer)
    %layer = Flatten()(layer)
    %layer = Dropout(0.20)(layer)
    %layer = Dense(20)(layer)
    %layer = Dropout(0.10)(layer)
    %output_layer = Dense(1, activation='sigmoid', name='main_output')(layer)
    %model = Model(input=input_layer, output=output_layer)
    %model.compile(optimizer='adam', loss='binary_crossentropy', metrics=['accuracy'])

### 2018-02-05 qg_ML_4.tex
update from raghav, started going through the text

@@
 \newcommand{\text}[1]{{\rm #1}}
-\newcommand{\jw}{\textsc{Jewel}}
-\newcommand{\jwpy}{\textsc{Jewel+Pythia}}
+\newcommand{\jw}{\textsc{Jewel}~}
+\newcommand{\jwpy}{\textsc{Jewel+Pythia}~}
 \newcommand{\py}{\textsc{Pythia}~}
@@
 
-\author{Yang-Ting Chien $^{a}$ and Raghav Kunnawalkam Elayavalli $^{b}$\\
+\author{Yang-Ting Chien $^{a}$ and Raghav Kunnawalkam Elayavalli $^{b,c}$\\
 $^{a}$ Center for Theoretical Physics\\
@@
 $^{b}$ Department of Physics and Astronomy\\
+$~$ Wayne State University, Detroit, MI 48201\\
+$^{c}$ Department of Physics and Astronomy\\
 $~$ Rutgers, the State University of New Jersey, New Brunswick, NJ 08901
@@
 \abstract{
-We introduce a novel jet quenching study framework utilizing both quark-initiated jets and gluon-initiated jets as independent probes of heavy ion collisions.
-We perform quark gluon discrimination in proton-proton and heavy ion collisions using jet substructure techniques to exploit all the jet features.
+We introduce a novel framework to study the phenomenon of jet quenching utilizing both quark-initiated jets and gluon-initiated jets as independent probes of heavy ion collisions.
+We perform quark-gluon discrimination in proton-proton and heavy ion collisions using jet substructure techniques to exploit all the jet features.
 Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are collectively examined in a multivariate analysis.
 In comparison, we perform the classification using machine-learning methods by training a deep convolutional neutral network on discretized jet images.
-Furthermore, we develop the telescoping deconstruction framework with a systematic subjet expansion to exploit the information carried by the subjets at multiple angular scales.
+Furthermore, we introduce the telescoping deconstruction framework with a systematic subjet expansion to exploit the information carried by the subjets at multiple angular scales.
 We draw connections to the soft-drop subjet distribution and illuminate modifications of quark jets and gluon jets using Lund diagrams.
 The quark and gluon jet samples generated from \jwpy ~are used as an example to demonstrate our general method.
-We find that the classification performance goes down in the \jwpy-simulated heavy ion collisions.
+We find that the classification performance worsens in the \jwpy-simulated heavy ion collisions.
 The information carried by the subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation.

### 2018-02-09 qg_ML_4.tex
update - still need a lot of work on language on abstract, lund diagrams and conclusions

@@
 
-The jet quenching phenomenon observed at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
-and the Large Hadron Collider (LHC)\cite{Aamodt:2010jd,Abelev:2012hxa,Abelev:2013kqa,Aad:2012vca,Aad:2014bxa,Chatrchyan:2013kwa,
+The jet quenching phenomenon observed in experiments at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
+and the Large Hadron Collider (LHC)~\cite{Aamodt:2010jd,Abelev:2012hxa,Abelev:2013kqa,Aad:2012vca,Aad:2014bxa,Chatrchyan:2013kwa,
 Chatrchyan:2012gw,Chatrchyan:2014ava,Aad:2014wha,Chatrchyan:2013exa,
 Adam:2015ewa,Chatrchyan:2012gt,Chatrchyan:2011sx,Chatrchyan:2012nia,Aad:2010bu,Aad:2013sla,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha} has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions.
-The dramatic suppression of hadron and jet cross sections was understood using the energy loss picture.
-By studying in details how the jet energy redistributes in heavy ion collisions, one can quantitatively extract the information about the jet-medium interaction and the underlying medium dynamics.
-The heavy ion jet physics has thus moved to the era of precision jet substructure and jet cross section studies.
+The dramatic suppression of hadron and jet cross sections is understood using the medium induced energy loss picture~\cite{Connors:2017ptx}.
+By studying the (re)distribution of the lost jet energy in heavy ion collisions, one can quantitatively extract key features of jet-medium interaction and the underlying medium dynamics.
+With the advent of ever more accurate data to model comparisons of jet structure measurements, field of heavy ion jet physics has thus moved to the era of precision jet substructure and jet cross section studies.
 
-Jet substructure measurements provide concrete information about the modifications of jets.
+Jet substructure measurements provide concrete and consistent physics insights on the modification of  jets in a heavy ion environment.
 From the measurements of the jet shape \cite{} and jet fragmentation function \cite{}, we have learned that both the transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructures depend strongly on the partonic origin of jets, the change of the fraction of quark-initiated jets and gluon-initiated jets will contribute to the substructure modifications.
-It has been realized that the increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of the jet cores seen in various inclusive jet substructure measurements.
-With purer quark and gluon jet samples, one can disentangle the effect from the changing quark and gluon jet fractions and focus on the jet-by-jet modifications.
-Quark jets and gluon jets can then be used as independent probes.
+Since jet substructure depend strongly on the partonic origin of jets, changing the fraction of quark-initiated jets and gluon-initiated jets will contribute to any substructure modifications observed in experiment.
+It has recently been realized that the increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of the jet cores seen in various inclusive jet substructure measurements~\cite{}.
+With purer identified quark and gluon jet samples, one can disentangle such an effect due to the changing quark and gluon jet fractions and focus on the jet-by-jet modifications.
+Quark jets and gluon jets can thus be used as independent jet quenching probes.
 
-We will study the tagging of quark and gluon jets making full use of their substructure information.
-In contrast to the conventional wisdom of directly comparing jets in proton-proton and heavy ion collisions, this indirect approach studies the modification of the comparison between different probes.
-By first comparing the probes in the same system, it may help reducing the systematic uncertainty in experimental measurements.
-The lowest order feature which separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
+We systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions by making full use of their substructure information.
+In contrast to the conventional wisdom of directly comparing inclusive jets in proton-proton and heavy ion collisions, this indirect approach studies the modification of the comparison between different probes.
+By first comparing the probes in the same system, it may help in reducing the systematic uncertainty in relevant experimental measurements.
+The lowest order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
 The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets.
-Various jet substructure variables have been used and combined to further increase the tagging performance with a multivariate analysis, and the advances of machine learning techniques have been constantly incorporated in solving this classification problem.
+Various jet substructure variables and their combinations have been used thus far within a multivariate technique to tag such jet favors and advances in machine learning techniques have recently been incorporated in achieving higher tagging efficiency for this classification problem.
 
-In this paper we benchmark three classes of methods in quark and gluon tagging and gain insights from comparing their performances.
-We use modern image recognition techniques utilizing deep convolutional neural networks (DCNN).
-The network is trained on the discretized images of quark jets and gluon jets, and the samples are generated from the \jw heavy ion event generator.
-Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments.
+In this paper we benchmark three classes of methods in quark and gluon tagging and gain insights from their comparisons.
+We employ a modern 2D image recognition technique involving a deep convolutional neural networks (DCNN).
+The network is trained on the discretized images of quark jets and gluon jets.
+A jet image is a representation of the jet constituents's energy distribution on a 2D plane orthogonal to the jet axis with each jet image undergoing standard pre-processed described here!\cite{}.
 We also examine the tagging performance using five physically constructed jet substructure variables: the jet mass, the $p_T^D$, the multiplicity and two radial moments including the girth.
-Subsets of the variables are combined and studied in a multivariate analysis using boosted decision trees, and we look into the correlations among the variables.
-Lastly, we systematically expand the information within jets using the recently developed telescoping deconstruction method.
-The corresponding subjet expansion can be truncated at any finite order $N$ which allows us to access the information carried by the subleading subjets.
+Subsets of the variables are combined and studied in a multilayer perceptron (MLP) with two hidden layers in order to include correlations amongst the variables.
+Lastly, we systematically expand the information within jets using the recently developed telescoping deconstruction (TD) method.
+The TD at a given order $N$, deconstructs a jet into its fragmentation basis with regard to $N$ axis which are defined using the winner take all (WTA) method.
+The corresponding subjet expansion can be truncated at any finite order which allows us to access the information carried by the subleading subjets.
 
@@
 In Sec.~\ref{sec:results} we show the performance of each method and the distributions of jet substructure observables.
-In Sec.~\ref{sec:conc} we conclude and give an outlook of future studies.
+We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.
+
+% and the samples are generated from the \jw heavy ion event generator.
+Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments.
 
@@
 
-To use quark jets and gluon jets as independent probes of heavy ion collisions, we first need to identify the probes we are using.
-In this section we describe the quark and gluon jet samples we use to demonstrate our general jet quenching study method.
-As an example, the jet samples are generated from the \jwpy ~simulations of proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=13$ TeV using the standard setup \cite{Zapp:2013zya}.
-We use the prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ for constructing the quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
-It is possible to construct the jet samples using a completely physical way by performing hard process and kinematical selections to enhance either the quark or the gluon jet fraction.
+To successfully use quark and gluon jets are independent probes of heavy ion collisions, we need to first define and identify the respective probes.
+In this paper, we define jet flavor based on the progenitor parton in the parton shower evolution.
+We use the \jwpy MC simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
+\jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
+The prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$, provide the quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
+It is possible to construct the jet samples by performing hard process and kinematical selections (such as dijet vs z+jet) to enhance either the gluon or the quark jet fraction.
 This can then be directly applied to analyze other jet simulations and experimental data in the future.
 
-In each collision event, jets are reconstructed using the anti-$k_{t}$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$.
+In each event, jets are reconstructed using the anti-$k_{t}$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$.
 We impose the following kinematic cuts,
@@
 where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon.
-\jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
-The medium-induced energy loss in heavy ion collisions is calculated based on a medium model consisting of thermal scattering centers undergoing the Bjorken expansion.
+The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing the Bjorken expansion.
 The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
 We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHApdf}~\cite{Whalley:2005nh}.
-We include recoil partons which are scattering centers interacting with the propagating hard scattered partons in the \jw~simulations.
+We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with the thermal scattering centers due to interactinos with the propagating hard scattered partons.
 It was shown that recoil contributions are important to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-On the other hand, a large fraction of the jet energy comes from the huge underlying event background in heavy ion collisions which needs to be subtracted before one could study the intrinsic modification of jets.
-We adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo}.
-The \jw ~output HepMC files are analyzed using the \textsc{Rivet} analysis framework~\cite{}.
+In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo}.
+The \jw simulation produces HepMC~\cite{} files which are intern read using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.
 
@@
 
-With finite detector angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC expeirements.
+To emulate finite detector effects on the jet's angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC.
 To study how such discretization as well as background subtraction affect the analysis, we use the following three different sets of jets in heavy ion collisions, all with the medium recoil contributions,
@@
 For jets in proton-proton collisions, we use the sets of jet samples with or without discretization.
-In reality, the underlying event background subtraction for jet substructure observables is highly challenging and nontrivial, which can introduce significant smearing effects.
-To disentangle such an issue, one common practice adopted by the CMS collaboration is to smear jets in pp collisions using minimum bias heavy ion events before comparing with jets in heavy ion collisions.
-In this work we do not address this issue with the \jwpy-simulated jet samples and assume that the built-in background subtraction procedure allow a fair comparison between jets in pp and PbPb collisions.
+In data, the fluctuating underlying is subtracted via experiment specific background subtraction techniques.
+Since \jw does not reproduce the full underlying event in our quenched samples, we do not need to apply those overall background removal methods.
+Upon the application of the grid subtraction technique, these collections of jets represent an ideally subtracted jet in experiment enabling us to focus our attention and highlight the contribution of jet quenching as opposed to additional nontrivial smearing effects.
 
@@
 \label{sec:qg}
-\begin{figure}
+\begin{figure}[h]
 	   \centering
@@
 
-We will use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns.
+We use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns.
 In doing so one attempts to identify jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms.
-This can be quantified by studying the modification of specific jet substructure observables.
-Such direct approach is closely related to the outstanding problem of discriminating quark and gluon jets, where the goal is to exploit jet features which show differences between quark jets and gluon jets.
+Such direct approaches are closely related to the outstanding problem of discriminating quark and gluon jets, where the goal is to exploit jet features which show differences between quark jets and gluon jets.
 One may compare quark jets and gluon jets in the same collision system and use their differences as a way to quantify the jet modification.
 This can help disentangle jet quenching effects on common features between quark and gluon jets and cancel systematic uncertainties in experiments.
-A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination}.
+A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the color factor difference (horizontal arrows) are highlighted.
 
-Quark gluon discrimination has been studied extensively in the past\cite{}.
+Quark gluon discrimination has been studied extensively in the recent past, particularly at higher kinematic selections at the LHC inorder to highlight beyond the standard model signatures\cite{}.
 With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful ones \cite{Gallicchio:2011xq,Gallicchio:2012ez}.
-On the other hand, using modern machine learning image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance can be significantly increased.
+On the other hand, using modern machine learning image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to be better at the high momenta selections at the LHC.
 This suggests that additional jet features have been identified by the deep neural network to help the tagging task.
 We use both methods to gain insights on how jet quenching is manifested in these two jet representations.
-Furthermore, we develop the telescoping deconstruction framework which represents the whole jet information using subjets with different angular resolution.
-It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet features.
-The details about telescoping deconstruction are provided in Appendix.
+Furthermore, we develop the telescoping deconstruction framework which represents the whole jet information using subjets with varying angular resolution.
+It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features.
+The more technical details about the telescoping deconstruction algorithm and an example utilization of comparing QCD and W jets are provided in the Appendix.
 
@@
 
-Multivariate analysis have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus an effective, physics-motivated baseline in our study.
+Multivariate analysis have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study.
 Given a jet reconstructed using the anti-$k_{t}$ algorithm \cite{Cacciari:2008gp}, we choose the following set of five jet substructure observables \cite{} to highlight the differences in quark jets and gluon jets,
 	\begin{itemize}
-		\item Jet mass: $m = \sqrt{(\sum_{i\in {\rm jet}} p_i)^2}$. 
+		\item Jet mass: $m = \sqrt{E^{2} - (\sum_{i\in {\rm jet}} p_i)^2}$. 
         		\item First radial moment (girth): $\sum_{i \in {\rm jet}} p^i_{T} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
@@
 
-Fig:~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the above jet observables in pp and PbPb collisions.
-The quark and gluon jets used in plotting Fig:~\ref{fig:jetdistributons_pp_pbpb} include the grid dicretization.
-Furthermore, jets in PbPb collisions are subtracted using the GridSub background subtraction method.
+Figure :~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
+Both pp and PbPb jets include the grid discretization and furthermore the jets in PbPb collisions are subtracted using the GridSub background subtraction method.
 Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions.
 In vacuum, they reflect the fact that gluon jets have larger jet mass and radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$.
-In the \jwpy-simulated PbPb collisions, for both quark and gluon jets the distributions are consistently modified towards larger values.
+In the \jwpy-simulated PbPb collisions, for both quark and gluon jets the distributions are modified towards larger values.
 On the other hand, $p_T^D$ and the pixel multiplicity are IRC unsafe.
@@
 It is important to note that these observables are sensitive to the jet clustering algorithm, the minimum \pt ~cutoff of jet constituents and the background subtraction procedures employed.
-In order to facilitate a direct comparison between experimental data and Monte Carlo simulations, one has to either unfold the detector resolution from these observables or perform Monte Carlo simulations with detector response.
+In order to facilitate a direct comparison between experimental data and Monte Carlo simulations, one has to either unfold the detector resolution from these observables or perform Monte Carlo simulations with the appropriate detector response.
 The unfolding procedure for jet observables increases in dimensionality and complexity as the correlations among observables become strong.
@@
 
-We combine the five jet substructure observables using a multivariate model implemented in Keras, with two dense layers of 10 nodes each.
+We combine the five jet substructure observables using a multivariate model implemented in Keras, with two hidden dense layers of 10 nodes each.
 We use the rectified linear unit (\textit{relu}) and sigmoid activation functions for the dense layers and the output layer, respectively.
 The model training utilizes the Adam optimizer and the binary cross entropy loss function.
-We perform cross validation using Monte Carlo samples which are split into two halves for training and testing.
+We perform cross validation using Monte Carlo samples which are split into two random halves for training and testing.
 
@@
 The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane of transverse momentum deposit in detector calorimeter cells, much like a digital camera.
-The jet image thus represents a fixed-dimensional representation of jets.
+The jet image thus represents a fixed-dimensional representation of jet fragmentataion.
 Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in $\eta$ and $\phi$ plane so that the highest pixel centers at the origin.
@@
 
-Fig:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
+Figure:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
 The color of each image pixel represents the transverse momentum deposited in the pixel.
@@
 Note that in experiment it may be very challenging to directly analyze jet images due to detector and measurement considerations.
-Here we mainly use the jet image approach as a useful comparison of the classification performance.
+Here we mainly use the jet image approach as a useful and state of the art comparison of the classification performance.
 
-A deep convolutional neural network implemented in Keras with the TensorFlow backend is utilized in this analysis.
+A deep convolutional neural network is implemented in Keras with the TensorFlow backend.
 The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer.
@@
 Telescoping deconstruction aims to embrace both advantages of multivariate analysis and jet image recognition.
-It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized basis of jet substructures.
+It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized fragmentation basis of jet substructures.
 It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables.
-The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically-meaningful which encode the hard splitting of jets and non-perturbative physics.
+The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically motivated to encode the hard splitting of jets and its relevant non-perturbative physics within each subjet variable.
 The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions.
@@
 
-For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with steps of 0.08.
+For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with 5 steps of 0.08.
 The subjet observables up to the $N$-th order are combined using dense neural networks to maximize the classification performance.
-We consider $N=\{1,2,3\}$ and the performance quickly saturates at $N=3$.
-{\bf add some ML details}
+We consider $N=\{1,2,3\}$ in this study and find the performance quickly saturates at $N=3$.
+Each order in telescoping provides $N_{steps} \times 2$ variables in addition to the subjet axis information such as $\eta, \phi$, i.e for each step in $\Delta r$, we calculate the subjet transverse momenta and the mass.
+Thus, third order TD includes by definition, all variables from the second and first order TD expansion of a jet.
+These telescoping variables are fed into a multi layer perceptron, much like the one we used in the physics variables discussion, with two hidden dense layers and a final output layer and the activation and optimization functions.
 
 Beside the overall quark gluon classification performance using telescoping deconstruction, subjet variables within this framework can reveal many aspects of jet dynamics.
-In particular, there is rich information contained in subjet topology and subjet substructure.
-Recently, the CMS \cite{} and STAR \cite{} collaborations measured the groomed momentum sharing observable $z_g$.
+In particular, there is rich information contained in subjet topology and substructure.
+Recently, the CMS~\cite{} and STAR~\cite{} collaborations measured the groomed momentum sharing observable $z_g$.
 One reclusters jets using the Cambridge/Aachen algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and sequentially removes wide angle, soft radiation until the following condition is satisfied by a branching,
@@
 The distribution of the momentum fraction $z_{\rm T2}$ of the softer subjet is significantly different from the $z_g$ distribution, especially for gluon jets.
-Note that the axes used in the telescoping deconstruction procedure are defined with the Winner-Take-All scheme which favors energetic particles.
+We remind the reader that the axes used in the telescoping deconstruction procedure are defined with the Winner-Take-All scheme which favors energetic particles.
 We find that the axes at the T2 order may both align with the hard branch and are not able to efficiently tag the soft branch when it consists of soft particles.
 The situation is significantly improved at the T3 order where we use three axes to capture the soft branch in soft-drop.
-We combine two closest subjets, reducing the number of subjets to two, and the distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark and gluon jets are very similar.
-In comparison, we reconstruct subjets around the axes of the soft-drop branches to probe their energy distributions.
-We find that the distribution of the momentum fraction $z_{{\rm T2},g}$ of the softer subjet is very similar to that of $z_{\rm T3}$ and different from that of $z_g$.
-This suggests that the active areas of the two soft-drop branches are different.
-On the other hand, the subjet substructure may allow us to extract the information about the partonic origin of subjets.
-We look at the scaleless ratio $m_i/p_{T_i}$ where $m_i$ is the light (heavy) subjet mass, or the mass of the hard (soft) subjet.
+We combine two closest subjets, whilst applying the grooming criteria on the soft subjet and the distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark (left) and gluon (right) jets are very similar as shown in Fig::~\ref{fig:comp_z}.
+The solid and dotted lines represent the soft dropped $z_{g}$ in pp and PbPb where we can see an enhancement of asymmetrical jets in PbPb.
+The subjet fraction calculated from the TD third order (T3) for both quark and gluon jets are less modified due to quenching (comparing the dashed with the dot-dashed in Fig:~\ref{fig:comp_z}) as compared with the softdropped zg.
+
+\begin{figure}[h]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_z.pdf}
+	   \caption{Transverse momentum fraction of the sub-leading subjet for Quark (left) and Gluon (right) jets in pp (solid) with soft drop, pp with TD T3 (dashed), quenched PbPb soft drop (dotted) and quenched PbPb TD T3 (dot dashed) lines.
+Subjet fraction is enclosed between 0.1 and 0.5 due to the $z_{cut}$ applied while grooming.}
+\label{fig:comp_z}
+\end{figure}
+
+In addition to the momentum fraction of the subjets, there is also recent interest in measuring the angular separation between the two subjets.
+We show the calculation of the this angular distance between the leading and subleading subjets in Fig:~\ref{fig:comp_rg} for quark (left) and gluon (right) jets.
+For the $R{g}$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
+While this particular observable has not yet been experimentally measured at the LHC or the RHIC, the phenomenon of broadening or widening of the jet observed in these MC calculations are consistent with measurements of the jet shape~\cite{}.
+Both the softdrop $R_{g}$ and the T3 $\theta$ show deviations due to quenching, but the quenched distributions do not match up with each other.
+Softdrop shows a narrow peak for quenched jets at 0.3, where as the T3 distribution is broader since the sub-jet axis definition are not the same (Softdrop uses C/A whereas TD uses WTA).
+Thus these are similar observables that in comparison, we extract the physics due to quenching but these do not provide an apples to apples comparison.
+
+\begin{figure}[h]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg.pdf}
+	   \caption{Angular distance between the leading and subleading subjets for Quark (left) and Gluon (right) jets in pp (solid) with soft drop, pp with TD T3 (dashed), quenched PbPb soft drop (dotted) and quenched PbPb TD T3 (dot dashed) lines.
+The quenched $R_{g}$ distribution is significantly modified compared the vacuum case.}
+\label{fig:comp_rg}
+\end{figure}
+
+We also study the subjet mass for both the subjets that represent the two major prongs of our jet.
+The subjet mass distribution for hard (solid - pp, dotted - PbPb) and soft (dashed - pp, dot-dashed - PbPb) is shown for quark (left) and gluon (right) jets in Fig:~\ref{fig:comp_subjet_m}.
+In this particular case, hard or soft refers to the subjet's momentum with harder meaning the one with higher transverse momentum.
+The first point of difference is seen due to the effect of quenching, i.e.
+the pp hard and soft subjet mass are comparable with the softer subjet having a larger tail.
+The quenched softer subjet as expected has a much larger tail towards larger masses indication its constituent distribution is spread out over a larger area as opposed to the quenched harder subjet.
+
+\begin{figure}[h]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_subjet_m.pdf}
+	   \caption{Leading subjet mass for Quark (left) and Gluon (right) jets in pp (solid) and  quenched PbPb (dotted) in comparison with the subleading subjet mass in pp (dashed) and quenched PbPb (dot dashed) lines.
+}
+\label{fig:comp_subjet_m}
+\end{figure}
+
+WRITE THE CONCLUSION HERE FOR THIS PARTICULAR SECTION - WHAT DO WE LEARN
 
@@
 For quark gluon discrimination, the performances are shown using Receiver Operating Characteristic (ROC) curves which plot the gluon jet efficiency as a function of the quark jet efficiency.
-We also perform discriminations between jets in pp and AA collisions for both quark jets and gluon jets, and the performance is evaluated by the ROC curves showing AA jet efficiency as a function of the pp jet efficiency.
-A lower ROC curve represents a better performance.
+Each ROC curve is labeled in the following figures with the corresponding area under the curve (auc) where the best performance is noted by large values of quark efficiency with small values of gluon efficiency leading to a small auc.
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
-	   \caption{}
+	\caption{ROC curves for quark vs gluon efficiency for particle jets in pp (left) and PbPb (right).
+The different colored curves show the variety of classification models studied in this paper and the quantity in the brackets marked by auc represents the area under the curve.
+The green solid and dotted lines are for the MLP with physics variables w/ and w/o the pixel multiplicity.
+The blue solid line is with the DCNN on jet images and the brown, orange and teal colors are for the TD with T1, T2 and T3 orders respectively.
+(color online)}
 \label{fig:ROC_all}
@@
 
-Fig:~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and AA (right panel) collisions, using particle momenta when evaluating the observables without $\eta=\phi$ discretization and background subtraction.
-We see that overall the methods give similar performances, implying that the three jet representations capture sufficient jet information.
-Note that there is monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$ to $\rm T1+T2+T3$ with increasing input information of higher-order subleading subjets, and the performance saturates quickly at the T2 order.
-In the AA case, combining physics variables gives better performance due to the high particle multiplicity in the \jwpy-simulated AA events.
-Telescoping deconstruction with radius scan in steps of 0.08 is not able to capture the full multiplicity information at the T3 order.
+Figure:~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and PbPb (right panel) collisions for particle jets i.e jets reconstructed with particle momenta without $\eta=\phi$ discretization and background subtraction.
+We see an overall compatibility amongst the methods used leading to similar classification performances, implying that the three jet representations capture sufficient jet information.
+Note that there is monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$ to $\rm T1+T2+T3$ with increasing input information of higher-order subleading subjets.
+Is it important to note that the TD classification performance saturates quickly at the T2 order with additional higher order offering very little improvement in the quark efficiency given a gluon efficiency.
+In the PbPb case (right of Fig:~\ref{fig:ROC_all}), the MLP with physics variables gives better performance due to the high particle multiplicity in the \jwpy-simulated PbPb events.
+Measuring the jet multiplicity in heavy ion collisions is very tricky due to the underlying and its correlations with the jet.
+In nature of this difficulty, we also show the ROC curve for the MLP with physics variables without jet multiplicity denoted by the dotted green line in Fig:~\ref{fig:ROC_all} which reduces the overall auc by $0.02$ and becomes consistent with our DCNN.
+Even though the TD with radius scan in steps of 0.08 is unable to accurately capture the full multiplicity information at the T3 order it still within a few percent difference of the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
+In order to directly compare with data, we apply the grid discretization for both pp and PbPb with its specific GridSub background subtraction technique and compare the various orders of TD on the left of Fig:~\ref{fig:ROC_TD_grid}.
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
         \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
-       \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
-	   \caption{}
-\label{fig:ROC_TD_input}
+        \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
+	\caption{(Left) ROC curves for quark vs gluon efficiency with the TD method for grid jets in pp (dotted lines) and PbPb (solid lines) (w/ GridSub).
+The different colors are for different orders (T1 - blue, T2 - red, T3 - green; color online).
+(Right) ROC curves for quark vs gluon efficiency with the TD model at the third order for pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid no bkg-sub (solid red) and PbPb grid sub (solid blue) jets.
+The corresponding area under the curves are shown in the brackets.}
+	\label{fig:ROC_TD_grid}
 \end{figure}
 
-By comparing the performances in pp and AA collisions, we clearly see that the performances drops in AA collisions.
-To have a closer look at the change of the classification performance from pp to AA collisions and the convergence of the TD performance, we plot the ROC curves of telescoping deconstruction in both pp and AA collisions in the left panel of Fig:~\ref{fig:ROC_TD_input}.
-We see the shrink of the performance improvement from T1 to $\rm T1+T2$ in AA collisions.
+We clearly see that the performances drops in PbPb collisions along with a reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
 This suggests that the information carried in higher-order subleading subjets can be washed out in heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions in \jwpy.
-In the right panel of Fig:~\ref{fig:ROC_TD_input} we plot the TD ROC curves with different inputs with or without $\eta-\phi$ discretization and background subtraction.
+On the right of Fig:~\ref{fig:ROC_TD_grid} we plot the TD ROC curves for jet with inputs with or without $\eta-\phi$ discretization and background subtraction.
 Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance.
-On the other hand, the performance drops with the background subtraction performed.
-This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction.
+On the other hand, the performance drops with the background subtraction performed (slight increase in the auc).
+This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction quantitatively bringing the two flavors of jets together in their fragmentation basis.
 
-\begin{figure}
+\begin{figure}[h]
 	   \centering
@@
 	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
-	   \caption{}
+	   \caption{ROC curves for a given jet to be pp-like (in vacuum) or PbPb-like (quenched) for quark (left) and gluon jets on the grid (right).
+As before, the MLP with physics variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at first, second and third order are in solid maroon, orange and cyan lines.}
 \label{fig:ROC_qq_gg}
 \end{figure}
-Fig:~\ref{fig:ROC_all} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and AA collisions using physics-motivated variables, jet images and telescoping deconstruction.
-We see huge difference between jets in pp and AA collisions therefore we plot the ROC curve using log scale in the y-axis.
-The significantly better performance in the gluon jet discrimination suggests that the modification of gluon jets is larger.
-We observe that the increase of the pixel multiplicity is a key feature in jet modification.
+
+For a given jet flavor, i.e.
+quark or gluon jet, one can also study with the various aforementioned methods, how the medium modifies it by comparing a vacuum jet with a quenched jet.
+Figure:~\ref{fig:ROC_qq_gg} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
+We see huge difference between jets and all methods perform significantly better at the task of identifying if a jet is quenched or not as opposed to if its a quark or gluon jet and therefore we plot the ROC curve using log scale in the y-axis.
+We observe that the increase of the pixel multiplicity is a key feature in aiding the MLP in identifying quenched jets with the MLP only slightly worse than the DCNN.
+The better performance in the gluon jet discrimination (right panel of Fig:~\ref{ROC_qq_gg}) suggests that gluon jets are significantly more modified than quark jets due to medium induced quenching.
+
+From our studies of the ROC curve in comparing quarks vs gluons and in comparing vacuum vs quenched jets, we can begin to qualitatively understand the effect of quenching as implemented in the \jw MC.
+Quenched jets end up with significantly more jet constituents which are distributed away from the principal axis of the jet as seen in the saturation of TD classification performance from first to third order.
+This modification appears to affect gluon jets significantly more than quark jets since a gluon jet with its increased multiplicity and wider parton shower due to the color factor, can all undergo quenching, or in the case of \jw, interact with the scattering centers leading to a broader more quenched jet.
+For a systematic study of such differences in quenched quark vs gluon, we employ the use of Lund Diagram as we shall discuss in the next section.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	   \caption{}
+	\caption{Lund diagrams plotted with the relative $z$ of the soft subjet in the harder branch for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
+(See text for definition of the axis)}
 \label{fig:Lund_full}
 \end{figure}
-Fig:~\ref{fig:Lund_full} shows the Lund diagrams for quark jets (left) and gluon jets (right) in pp (top) and AA (bottom) collisions.
-Note that only the splitting in the soft-drop hard branch enter into the Lund diagrams.
-We see the significant contributions from soft, wide-angle radiation in AA collisions.
 
-\begin{figure}
+Any jet observable on the market is designed as some function of the momentum and the distance scale that is used in the jet reconstruction.
+These two scales can be represented as $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$, where $z$ is the transverse momentum fraction of the softer subjet relative to sum of the two subjets, $\Delta R$ is the angular scale between the two subjets in a given branch and $R$ is the jet radius.
+Such a representation of a jet is known as the Lund Diagram~\cite{Salam:2016yht} and the prescription to create such a diagram for a given jet is as follows:
+\begin{itemize}
+	\item Cluster your jet with a jet R say $0.4$ and a particular jet algorithm, in our case anti-k$_{t}$ 
+	\item Take the jet constituents and re-cluster with the C/A algorithm which combines the nearest particles with each other
+	\item Walk backwards along the clustering tree and for each branch calculate the $z=p^{2}_{T}/(p^{1}_{T}+p^{2}_{T})$ and $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$ 
+	\item Fill the Lund diagram at each split 
+	\item Continue in walking backwards along the hard branch till you reach only one particle 
+\end{itemize}
+
+\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	   \caption{}
+	\caption{Lund diagrams specifically showing subjets that are removed and tagged as background ($z<0.1$) for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
+(See text for definition of the axis)}
 \label{fig:Lund_bkg}
 \end{figure}
-Fig:~\ref{fig:Lund_bkg} shows the Lund diagrams for C/A trees before the one which passes the soft-drop condition. We see that the soft, wide-angle radiation is isolated.
 
-\begin{figure}
+The beauty of a Lund diagram is its convenient separation of wide soft angle radiation close to the y-axis, collinear hard radiation along the diagonal and with contribution from hadronization and underlying event in a line parallel to the x-axis.
+These Lund diagrams have also been recently used in distinguishing between QCD and boosted heavy objects due to their unique feature of featuring jets of a constant mass along the main diagonal (opposite to the collinear hard radiation).
+We show the Lund diagrams for quark (left) and gluon (right)jets in pp (top) and PbPb (bottom) in Fig:~\ref{fig:Lund_full} where the z is calculated relative to each split.
+We see an enhanced region in PbPb collisions close to the y-axis in comparison to pp which corresponds to increased wide soft angle in quenched jets.
+Comparing left to right (as we discussed in Fig:~\ref{fig:quenching_discrimination}) one also sees gluon jets include a larger fraction of the soft radiation and these are susceptible to hadronization and underlying event effects.
+The hard collinear radiation region of the Lund diagram, that is dominated in pp is shown to reduce in its significance for quenched jets.
+The core of the jet remains similar between pp and PbPb, but the secondary sub-jet ultimately gets pushed to wider angles and to softer momenta.
+Such a picture is exactly what we discussed in Fig:~\ref{fig:comp-zg}-\ref{fig:comp_subjet_m} but this time, visible via the Lund diagram.
+
+\begin{figure}[h]
 	   \centering
 	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
-	   \caption{}
+	   \caption{Lund diagrams created by following the first hard branch selected by $z>0.1$ for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
+(See text for definition of the axis)}
 \label{fig:Lund_hard}
 \end{figure}
-Fig:~\ref{fig:Lund_hard} shows the Lund diagrams for C/A trees after the hard branch in soft-drop.
-We see that the vacuum structure starts to emerge in the soft-drop AA jets.
-However, soft-drop can not remove all the soft radiation within the hard branch in AA collisions.
 
-\begin{figure}
-	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_zgrg.pdf}
-	   \caption{}
-\label{fig:Lund_zgrg}
-\end{figure}
+One can now study the effect of grooming with the help of the Lund diagrams by rejecting the soft radiation ($z < 0.1$) that is typically groomed a way the softdrop method.
+Fig:~\ref{fig:Lund_bkg} shows the Lund diagrams for our jets in pp and PbPb  before the one which passes the soft-drop condition.
+We see that the soft, wide-angle radiation is isolated and the hard collinear core is essentially removed.
+The bottom right panel of Fig:~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft wide angle radiation.
+These soft jet constituents in the jet's periphery are precisely the ones that show up as a modification of the jet shape.
+Within the framework of this diagrammatic representation, one can now build up these observables and follow the harder branch leading from the first split.
+Fig:~\ref{fig:Lund_hard} shows the Lund diagrams after the soft branch (background) has been removed via the soft-drop procedure.
+We see that the vacuum structure starts to emerge in the soft-drop PbPb jets.
+However, soft-drop can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation.
+We understand this behavior as the correlated medium response that belongs to the jet but is often removed due to a statistical treatment of background in experiments.
 
@@
 
+We have shown a systematic study of modification to the jet structure due to the effects of quenching in heavy ion collisions as implemented within the \jw MC generator.
+The medium modification is studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets.
+We compare a MLP with physics motivated jet variables and a state of the art DCNN applied on jet images to a our method of Telescoping Deconstruction and find the classification performance worsens for quenched jets when compared with vacuum pp jets.
+All of the models studied perform consistently with each other giving us the confidence that the TD to third order is extracting fundamental fragmentation patters of a given jet.
+
+DISCUSSION - THIS IS A VERY IMPORTANT PART
+
 \section*{Acknowledgments}
-The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributing to Appendix.
-The authors also thank Benjamin Nachman and Sevil Salur for helpful discussions, and .
-Y.-T.
-Chien was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
-R.K.
-Elayavalli acknowledges support from the National Science Foundation under Grant No.1067907 \& 1352081 (+JETSCAPE?).
+The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix.
+The authors also thank Benjamin Nachman, Joern Putschke, Konrad Tywoniuk and Sevil Salur for helpful discussions.
+We thank Gavin Salam and the organizers of CERN jet workshop (2017) for bringing to our attention the concept of Lund Diagrams and their usage in studying jet modifications.
+YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
+RKE acknowledges partial support from the NSF under Grant No.1067907 \& 1352081 and DOE (need to ask Joern).
 
@@
 %This is highlighted in the top case, where the telescoping deconstruction DNN outperforms the CNN at T2 order and above.
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \end{document}
-
-%We present the telescoping deconstruction framework as a complete, physically-motivated expansion of the jet.
-Neural networks were trained on the telescoping deconstruction observables up to different T$N$ orders and applied to quark/gluon discrimination, boosted $W$ and top tagging.
-Telescoping deconstruction observables up to different T$N$ orders are combined using multivariate methods and applied to quark/gluon discrimination, boosted $W$ and top tagging.
-The T$N$ performance converges efficiently and compares favorably to a CNN trained on jet images.
-Further, the individual telescoping deconstruction observables were shown to cleanly encode relevant features of jets such as the isolation of $W$ jets, the presence of the $W$ in top jets, and the quark- or gluon- like nature of the heaviest prongs in top and QCD jets.
-This complete and general framework is suitable for systematic studies of full jet dynamics, with rich future applications in parton-shower tuning, heavy ion hard probe studies, and identification of hadronic electroweak-boson emissions in ultra-high energy jets.
-The extension to hard-process tagging and resonance search at the event level is also promising.
-
-    %layer = Convolution2D(8, 11, 11, border_mode='same')(input_layer)
-    %layer = Activation('tanh')(layer)
-    %layer = MaxPooling2D(pool_size=(2,2))(layer)
-    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
-    %layer = Activation('tanh')(layer)
-    %layer = MaxPooling2D(pool_size=(3,3))(layer)
-    %layer = Convolution2D(8, 3, 3, border_mode='same')(layer)
-    %layer = Activation('tanh')(layer)
-    %layer = MaxPooling2D(pool_size=(3,3))(layer)
-    %layer = Flatten()(layer)
-    %layer = Dropout(0.20)(layer)
-    %layer = Dense(20)(layer)
-    %layer = Dropout(0.10)(layer)
-    %output_layer = Dense(1, activation='sigmoid', name='main_output')(layer)
-    %model = Model(input=input_layer, output=output_layer)
-    %model.compile(optimizer='adam', loss='binary_crossentropy', metrics=['accuracy'])

### 2018-02-17 qg_ML.tex
YT updated the draft up to the TD section.

--- 2018-02-09 qg_ML_4.tex
@@
 \newcommand{\py}{\textsc{Pythia}~}
+
+\DeclareRobustCommand{\Sec}[1]{Sec.~\ref{#1}}
+\DeclareRobustCommand{\Secs}[2]{Secs.~\ref{#1} and \ref{#2}}
+\DeclareRobustCommand{\Secss}[3]{Secs.~\ref{#1}, \ref{#2} and \ref{#3}}
+\DeclareRobustCommand{\App}[1]{App.~\ref{#1}}
+\DeclareRobustCommand{\Tab}[1]{Table~\ref{#1}}
+\DeclareRobustCommand{\Tabs}[2]{Tables~\ref{#1} and \ref{#2}}
+\DeclareRobustCommand{\Fig}[1]{Fig.~\ref{#1}}
+\DeclareRobustCommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
+\DeclareRobustCommand{\Figss}[3]{Figs.~\ref{#1}, \ref{#2} and \ref{#3}}
+\DeclareRobustCommand{\Eq}[1]{Eq.~(\ref{#1})}
+\DeclareRobustCommand{\Eqs}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
+\DeclareRobustCommand{\Eqss}[3]{Eqs.~(\ref{#1}), (\ref{#2}), and (\ref{#3})}
+\DeclareRobustCommand{\Ref}[1]{Ref.~\cite{#1}}
+\DeclareRobustCommand{\Refs}[2]{Refs.~\cite{#1} and \cite{#2}}
 
@@
 
-\DeclareRobustCommand{\Sec}[1]{Sec.~\ref{#1}}
-\DeclareRobustCommand{\Secs}[2]{Secs.~\ref{#1} and \ref{#2}}
-\DeclareRobustCommand{\Secss}[3]{Secs.~\ref{#1}, \ref{#2} and \ref{#3}}
-\DeclareRobustCommand{\App}[1]{App.~\ref{#1}}
-\DeclareRobustCommand{\Tab}[1]{Table~\ref{#1}}
-\DeclareRobustCommand{\Tabs}[2]{Tables~\ref{#1} and \ref{#2}}
-\DeclareRobustCommand{\Fig}[1]{Fig.~\ref{#1}}
-\DeclareRobustCommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\DeclareRobustCommand{\Figss}[3]{Figs.~\ref{#1}, \ref{#2} and \ref{#3}}
-\DeclareRobustCommand{\Eq}[1]{Eq.~(\ref{#1})}
-\DeclareRobustCommand{\Eqs}[2]{Eqs.~(\ref{#1}) and (\ref{#2})}
-\DeclareRobustCommand{\Eqss}[3]{Eqs.~(\ref{#1}), (\ref{#2}), and (\ref{#3})}
-\DeclareRobustCommand{\Ref}[1]{Ref.~\cite{#1}}
-\DeclareRobustCommand{\Refs}[2]{Refs.~\cite{#1} and \cite{#2}}
-
 \title{Probing heavy ion collisions using quark and gluon jet substructure}
@@
 \abstract{
-We introduce a novel framework to study the phenomenon of jet quenching utilizing both quark-initiated jets and gluon-initiated jets as independent probes of heavy ion collisions.
-We perform quark-gluon discrimination in proton-proton and heavy ion collisions using jet substructure techniques to exploit all the jet features.
-Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are collectively examined in a multivariate analysis.
-In comparison, we perform the classification using machine-learning methods by training a deep convolutional neutral network on discretized jet images.
-Furthermore, we introduce the telescoping deconstruction framework with a systematic subjet expansion to exploit the information carried by the subjets at multiple angular scales.
-We draw connections to the soft-drop subjet distribution and illuminate modifications of quark jets and gluon jets using Lund diagrams.
-The quark and gluon jet samples generated from \jwpy ~are used as an example to demonstrate our general method.
-We find that the classification performance worsens in the \jwpy-simulated heavy ion collisions.
-The information carried by the subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation.
-Our method provides a systematically improvable framework for analyzing and comparing jet simulations and measurements in heavy ion collisions.
+We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
+We exploit jet features to highlight the differences between quark jets and gluon jets and how differently they are modified in the medium.
+We perform quark-gluon and proton-heavy ion discriminations for jets in proton-proton and heavy ion collisions.
+Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are examined in a multivariate analysis.
+In comparison, we perform the classifications using image-recognition methods by training a deep convolutional neutral network on jet images.
+To systematically extract jet substructure information, we develop the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
+We draw connections to the soft-drop subjet distribution and illuminate jet modifications using Lund diagrams.
+The quark and gluon jet samples generated from \jwpy ~are used as an example to establish our general method.
+We find that the quark gluon discrimination performance worsens in the \jwpy simulated heavy ion collisions.
+The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation.
+Our work suggests a systematically improvable framework for testing and comparing jet simulations and measurements in heavy ion collisions.
 }
@@
 The jet quenching phenomenon observed in experiments at the Relativistic Heavy Ion Collider (RHIC) \cite{Adcox:2001jp,Adler:2002xw,Adcox:2004mh,Arsene:2004fa,Back:2004je,Adams:2005dq}
-and the Large Hadron Collider (LHC)~\cite{Aamodt:2010jd,Abelev:2012hxa,Abelev:2013kqa,Aad:2012vca,Aad:2014bxa,Chatrchyan:2013kwa,
-Chatrchyan:2012gw,Chatrchyan:2014ava,Aad:2014wha,Chatrchyan:2013exa,
-Adam:2015ewa,Chatrchyan:2012gt,Chatrchyan:2011sx,Chatrchyan:2012nia,Aad:2010bu,Aad:2013sla,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha} has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions.
-The dramatic suppression of hadron and jet cross sections is understood using the medium induced energy loss picture~\cite{Connors:2017ptx}.
-By studying the (re)distribution of the lost jet energy in heavy ion collisions, one can quantitatively extract key features of jet-medium interaction and the underlying medium dynamics.
-With the advent of ever more accurate data to model comparisons of jet structure measurements, field of heavy ion jet physics has thus moved to the era of precision jet substructure and jet cross section studies.
+and the Large Hadron Collider (LHC)~\cite{Aamodt:2010jd,Aad:2010bu,Chatrchyan:2011sx,Aamodt:2011vg,CMS:2012aa,Chatrchyan:2012nia,Abelev:2012hxa,Chatrchyan:2012gt,Aad:2012vca,
+Abelev:2013kqa,Aad:2013sla,Chatrchyan:2013exa,Aad:2014bxa,Khachatryan:2014bva,
+Adam:2015ewa,Adam:2015doa,Aad:2015bsa,Khachatryan:2015lha,Aad:2015wga,
+Khachatryan:2016odn,Khachatryan:2016xxp,Sirunyan:2016znt,Khachatryan:2016ypw,
+Sirunyan:2017isk,Sirunyan:2017oug,Sirunyan:2017lzi,Sirunyan:2017jic,Sirunyan:2017xss,Sirunyan:2017qhf}
+has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions (see ~\cite{Connors:2017ptx} for a review on heavy ion jet measurements).
+The dramatic suppression of hadron and jet cross sections is understood using the medium-induced energy loss picture \cite{Gyulassy:1993hr,Wang:1994fx,Zakharov:1996fv,Zakharov:1997uu,Baier:1996kr,Baier:1998kq,
+Gyulassy:2000er,Gyulassy:2000fs,Wiedemann:2000za,Wang:2001ifa,
+Arnold:2001ba,Arnold:2001ms,Arnold:2002ja,Casalderrey-Solana:2014bpa}.
+In order to quantitatively extract key features of the jet-medium interaction and the underlying medium dynamics, a precise understanding of the redistribution of jet energy in heavy ion collisions is becoming essential.
+With the proliferation of accurate comparisons between theoretical calculations and measurements \cite{Chatrchyan:2013kwa,Khachatryan:2016tfj,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec,Khachatryan:2016erx,
+Acharya:2017goa,Sirunyan:2017bsd,CMS:2017xdn}, the field of heavy ion jet physics has entered the era of precision jet substructure and jet cross section studies.
 
-Jet substructure measurements provide concrete and consistent physics insights on the modification of  jets in a heavy ion environment.
-From the measurements of the jet shape \cite{} and jet fragmentation function \cite{}, we have learned that both the transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructure depend strongly on the partonic origin of jets, changing the fraction of quark-initiated jets and gluon-initiated jets will contribute to any substructure modifications observed in experiment.
-It has recently been realized that the increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of the jet cores seen in various inclusive jet substructure measurements~\cite{}.
-With purer identified quark and gluon jet samples, one can disentangle such an effect due to the changing quark and gluon jet fractions and focus on the jet-by-jet modifications.
-Quark jets and gluon jets can thus be used as independent jet quenching probes.
+Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment.
+From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, we have established that both the transverse and longitudinal momentum distributions inside jets are significantly modified.
+Since jet substructure depends strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments.
+It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda}.
+With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the change of quark and gluon jet fractions caused by jet-medium interactions.
+This then allows one to focus on studying how differently quark and gluon jets are modified and use them as independent jet quenching probes.
+Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease.
+On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize the contributions from medium responses \cite{Tachibana:2017syd}.
+Quark and gluon jets will provide the opportunity for studying how differently the medium responds to different probes.
 
-We systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions by making full use of their substructure information.
-In contrast to the conventional wisdom of directly comparing inclusive jets in proton-proton and heavy ion collisions, this indirect approach studies the modification of the comparison between different probes.
-By first comparing the probes in the same system, it may help in reducing the systematic uncertainty in relevant experimental measurements.
-The lowest order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
+The use of quark jets and gluon jets as different hard probes of heavy ion collisions is closely related to the tagging of these two probes, which is an outstanding problem \cite{Gras:2017jty,Frye:2017yrw}.
+The lowest-order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
 The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets.
-Various jet substructure variables and their combinations have been used thus far within a multivariate technique to tag such jet favors and advances in machine learning techniques have recently been incorporated in achieving higher tagging efficiency for this classification problem.
+Extra distinctive features can be included to improve the tagging by combining various jet substructure variables in a multivariate analysis, and advances in machine learning techniques have recently been incorporated which helps achieving higher tagging efficiency \cite{Komiske:2016rsd}.
 
-In this paper we benchmark three classes of methods in quark and gluon tagging and gain insights from their comparisons.
-We employ a modern 2D image recognition technique involving a deep convolutional neural networks (DCNN).
-The network is trained on the discretized images of quark jets and gluon jets.
-A jet image is a representation of the jet constituents's energy distribution on a 2D plane orthogonal to the jet axis with each jet image undergoing standard pre-processed described here!\cite{}.
-We also examine the tagging performance using five physically constructed jet substructure variables: the jet mass, the $p_T^D$, the multiplicity and two radial moments including the girth.
-Subsets of the variables are combined and studied in a multilayer perceptron (MLP) with two hidden layers in order to include correlations amongst the variables.
-Lastly, we systematically expand the information within jets using the recently developed telescoping deconstruction (TD) method.
-The TD at a given order $N$, deconstructs a jet into its fragmentation basis with regard to $N$ axis which are defined using the winner take all (WTA) method.
-The corresponding subjet expansion can be truncated at any finite order which allows us to access the information carried by the subleading subjets.
+Besides the conventional wisdom in jet quenching studies by directly comparing jets in proton-proton and heavy ion collisions, we will systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
+This indirect, complimentary approach studies the modification of the differences between quark jets and gluon jets which disentangles modifications of their common features.
+By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements.
+We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}.
+However, the method is easily applicable to other heavy ion jet simulations \cite{} and experimental data as well.
+
+We will use three classes of methods and gain insights from comparing their performances.
+We combine in a multivariate analysis five physics-motivated and representative jet substructure variables which are effective in quark gluon discrimination \cite{Gallicchio:2012ez}.
+They include the jet mass, two radial moments including the girth, the $p_T^D$ and the pixel multiplicity.
+Subsets of the variables are combined and studied in a multilayer perceptron (MLP) with two hidden layers to extract correlations among the variables.
+We also employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN).
+The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a representation of the jet energy distribution.
+Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}.
+Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method using subjets with multiple angular resolutions.
+At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axis which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
+The corresponding subjet expansion can be truncated at any finite order which allows us to order-by-order examine the information carried by subleading subjets.
+These three methods are used in discriminating quark and gluon jets, as well as discriminating jets in proton-proton and heavy ion collisions.
+We will see that new insights can be gained from comprehensively comparing different probes across different collision systems.
+
+Because of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}.
+In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$.
+Within the catchment area of jets the significant background radiation needs to be subtracted before meaningful jet observable measurements can be made.
+Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Berta:2014eza,Komiske:2017ubm} is preferred.
+Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables.
+We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the telescoping deconstruction framework.
+We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent jets and illustrate the regions of phase space in the context of soft drop grooming procedure.
 
@@
 In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy ~simulation.
-In Sec.~\ref{sec:qg} we give the details about the multivariate analysis, the jet image, and the telescoping deconstruction methods.
-In Sec.~\ref{sec:results} we show the performance of each method and the distributions of jet substructure observables.
+In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as distributions of jet substructure observables.
+In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations.
+In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space separated by the soft-drop procedure.
 We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.
 
-% and the samples are generated from the \jw heavy ion event generator.
-Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments.
+% Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments.
 
@@
 
-To successfully use quark and gluon jets are independent probes of heavy ion collisions, we need to first define and identify the respective probes.
-In this paper, we define jet flavor based on the progenitor parton in the parton shower evolution.
-We use the \jwpy MC simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
+To use quark and gluon jets as independent probes of heavy ion collisions, we need to first define and identify the respective probes.
+In this paper, we define jet flavor based on the parton from hard scattering which initiates the parton shower evolution.
+We use the \jwpy Monte Carlo simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
 \jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
-The prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$, provide the quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
-It is possible to construct the jet samples by performing hard process and kinematical selections (such as dijet vs z+jet) to enhance either the gluon or the quark jet fraction.
+We use prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
+It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
+$Z$+jet) to enhance either the gluon or the quark jet fraction.
 This can then be directly applied to analyze other jet simulations and experimental data in the future.
@@
 where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon.
-The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing the Bjorken expansion.
+The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion.
 The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
 We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHApdf}~\cite{Whalley:2005nh}.
-We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with the thermal scattering centers due to interactinos with the propagating hard scattered partons.
-It was shown that recoil contributions are important to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo}.
-The \jw simulation produces HepMC~\cite{} files which are intern read using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.
+We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons.
+It was shown that recoil contributions are important in \jwpy to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+The \jw simulation produces HepMC~\cite{} files which are processed using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.
 
@@
 
-To emulate finite detector effects on the jet's angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC.
+To emulate finite detector effects on the jet angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC.
 To study how such discretization as well as background subtraction affect the analysis, we use the following three different sets of jets in heavy ion collisions, all with the medium recoil contributions,
@@
 	\item with discretization but without background subtraction
-	\item with discretization and the GridSub background subtraction.
+	\item with discretization and the GridSub \cite{KunnawalkamElayavalli:2017hxo} background subtraction.
 \end{itemize}
 For jets in proton-proton collisions, we use the sets of jet samples with or without discretization.
-In data, the fluctuating underlying is subtracted via experiment specific background subtraction techniques.
-Since \jw does not reproduce the full underlying event in our quenched samples, we do not need to apply those overall background removal methods.
-Upon the application of the grid subtraction technique, these collections of jets represent an ideally subtracted jet in experiment enabling us to focus our attention and highlight the contribution of jet quenching as opposed to additional nontrivial smearing effects.
+In experiments, the fluctuating, underlying event background is subtracted via specific background subtraction techniques.
+Since \jw does not produce full underlying events in quenched samples, we can not directly apply those background subtraction methods.
+With the application of the grid subtraction (GridSub) technique, the collection of jets represent ideally subtracted jets in experiment.
+This enables us to focus on and highlight the effects of jet quenching, as opposed to additional, nontrivial smearing effects.
 
@@
 \label{sec:qg}
-\begin{figure}[h]
+\begin{figure}[t]
 	   \centering
@@
 We use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns.
-In doing so one attempts to identify jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms.
-Such direct approaches are closely related to the outstanding problem of discriminating quark and gluon jets, where the goal is to exploit jet features which show differences between quark jets and gluon jets.
-One may compare quark jets and gluon jets in the same collision system and use their differences as a way to quantify the jet modification.
-This can help disentangle jet quenching effects on common features between quark and gluon jets and cancel systematic uncertainties in experiments.
-A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the color factor difference (horizontal arrows) are highlighted.
+In doing so one attempts to identify qualitatively different jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms.
+Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which show differences between them.
+The idea is that one can compare quark and gluon jet substructures in the same collision system and use their differences as a way to quantify the jet modification.
+New insights can be gained by putting jet substructures in multiple contexts.
+This can also help disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments.
+A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
 
-Quark gluon discrimination has been studied extensively in the recent past, particularly at higher kinematic selections at the LHC inorder to highlight beyond the standard model signatures\cite{}.
-With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful ones \cite{Gallicchio:2011xq,Gallicchio:2012ez}.
-On the other hand, using modern machine learning image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to be better at the high momenta selections at the LHC.
+Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures \cite{}.
+With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful selections \cite{Gallicchio:2011xq,Gallicchio:2012ez}.
+On the other hand, using modern machine-learning, image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to significantly improve at high transverse momentum.
 This suggests that additional jet features have been identified by the deep neural network to help the tagging task.
 We use both methods to gain insights on how jet quenching is manifested in these two jet representations.
-Furthermore, we develop the telescoping deconstruction framework which represents the whole jet information using subjets with varying angular resolution.
+Furthermore, we develop the telescoping deconstruction framework which represents the jet information using subjets with varying angular resolution.
 It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features.
-The more technical details about the telescoping deconstruction algorithm and an example utilization of comparing QCD and W jets are provided in the Appendix.
+More details about the telescoping deconstruction algorithm and examples comparing QCD, W and top jets are provided in the Appendix.
 
@@
 	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
-	   \caption{Distributions of the jet constituents (top left), invariant jet mass (top right), $p^{D}_{T}$ (center left), the half radial moment (center right) and the jet girth or the first radial moment (bottom) for quark (darker blue) and gluon (lighter orange) induced jets (color online).
-The top panels of each individual distribution show the pp and the bottom panels correspond to central (0-20\%) PbPb events generated by the \jw generator.}
+	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets.
+The top panels of each individual distribution show the pp ones and the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
 	   \label{fig:jetdistributons_pp_pbpb}
@@
 
-Multivariate analysis have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study.
-Given a jet reconstructed using the anti-$k_{t}$ algorithm \cite{Cacciari:2008gp}, we choose the following set of five jet substructure observables \cite{} to highlight the differences in quark jets and gluon jets,
+Multivariate analyses have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study.
+Given a jet reconstructed using the anti-$k_{t}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
 	\begin{itemize}
-		\item Jet mass: $m = \sqrt{E^{2} - (\sum_{i\in {\rm jet}} p_i)^2}$. 
+		\item Jet mass: $m = \sqrt{E^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
         		\item First radial moment (girth): $\sum_{i \in {\rm jet}} p^i_{T} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
@@
         		\item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^{i}_{T}}^2}/p^{\rm jet}_{T}$.
-        		\item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
+        		\item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
 	\end{itemize}
 
-Figure :~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
-Both pp and PbPb jets include the grid discretization and furthermore the jets in PbPb collisions are subtracted using the GridSub background subtraction method.
+Figure ~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
+Both pp and PbPb jets include the grid discretization.
+Furthermore, jets in PbPb collisions are subtracted using the GridSub background subtraction method.
 Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions.
@@
 We see that there is significant increase of the pixel multiplicity and decrease of the values of $p_T^D$.
-This is due to the presence of the increased number of soft particles inside jets from the \jw ~medium recoil contributions \cite{}.
+This is due to the presence of the increased number of soft particles inside jets from the \jw medium recoil contributions \cite{KunnawalkamElayavalli:2017hxo}.
 
@@
 
-We combine the five jet substructure observables using a multivariate model implemented in Keras, with two hidden dense layers of 10 nodes each.
-We use the rectified linear unit (\textit{relu}) and sigmoid activation functions for the dense layers and the output layer, respectively.
-The model training utilizes the Adam optimizer and the binary cross entropy loss function.
+We combine the five jet substructure observables using a multivariate model implemented in Keras \cite{keras}, with two hidden dense layers of 10 nodes each.
+We use the rectified linear unit (ReLU) \cite{nair2010rectified} and sigmoid activation functions for the dense layers and the output layer, respectively.
+The model training utilizes the Adam optimizer \cite{adam} and the binary cross entropy loss function.
 We perform cross validation using Monte Carlo samples which are split into two random halves for training and testing.
 
-%All five jet observables showcase the ability to classify quarks vs gluons in both pp and PbPb.
-Recent results from LHC experiments on the effect of jet quenching on inclusive jets showcase two main features; a slight narrowing of the jet core and increased low \pt particle multiplicity in the periphery regions of the jet leading to a broadening of the jet shape.
+%increased low \pt particle multiplicity in the periphery regions of the jet leading to a broadening of the jet shape.
 Such an effect is captured in \jw, with the introduction of the medium induced recoils and introduces a further complexity in the question of quark vs gluon classification in PbPb.
-In vacuum, we expect gluon jets to be broader than quark jets due to the different in the casimir color factor and thus in PbPb collisions, with both jets being quenched and broader, the ability to discriminate effectively between quark and gluon jets is expected to reduce when one only considers fragmentation based observables as we proceed to show in the upcoming sections.
 
@@
 
-The multivariate analysis discussed previously is a constructive approach to collect and examine the usefulness of jet features which may help distinguish quark jets from gluon jets.
-An alternative approach is to use image recognition techniques which optimize comprehensive multivariate models and identify all possible, useful features.
+The multivariate analysis discussed previously is a constructive approach to collect and examine the usefulness of jet features which may help distinguish quark jets from gluon jets, or jets in different collision systems.
+An alternative approach is to use image-recognition techniques which optimize comprehensive multivariate models and identify all possible, useful features.
 The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane of transverse momentum deposit in detector calorimeter cells, much like a digital camera.
-The jet image thus represents a fixed-dimensional representation of jet fragmentataion.
-Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in $\eta$ and $\phi$ plane so that the highest pixel centers at the origin.
+The jet image thus provides a fixed-dimensional representation of jet fragmentation.
+Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the highest pixel centers at the origin.
 Also, the image is rotated so that the principal component of the energy density is along the $\eta=0, \phi<0$ direction.
 
-Figure:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
+Figure~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
 The color of each image pixel represents the transverse momentum deposited in the pixel.
@@
 Note that in experiment it may be very challenging to directly analyze jet images due to detector and measurement considerations.
-Here we mainly use the jet image approach as a useful and state of the art comparison of the classification performance.
+Here we mainly use the jet image approach as a useful and state-of-the-art comparison of the classification performance.
 
-A deep convolutional neural network is implemented in Keras with the TensorFlow backend.
+A deep convolutional neural network is implemented in Keras with the TensorFlow \cite{} backend.
 The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer.
@@
 On the other hand, jet images represent low-level, comprehensive jet information.
-The training of a deep neural network is a ``top-down" approach to identify all the useful features through the model optimization.
+The training of a deep neural network is a ``top-down" approach to identify all useful features through model optimization.
 However, extracting physical messages from the trained neural network parameters can be a challenging task.
@@
 It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables.
-The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically motivated to encode the hard splitting of jets and its relevant non-perturbative physics within each subjet variable.
+The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
 The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions.
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction and then reaching out to include wide-angle, soft radiation.
-Fig:~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders.
-The details about the telescoping deconstruction procedures are provided in Appendix, including the applications on boosted $W$ and top tagging in high energy physics.
+Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders.
+The details about the telescoping deconstruction procedures are provided in the Appendix, including the applications on boosted $W$ and top tagging in high energy physics.
 
-For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with 5 steps of 0.08.
+For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with $N_{\rm steps} = 5$ steps of 0.08.
 The subjet observables up to the $N$-th order are combined using dense neural networks to maximize the classification performance.
 We consider $N=\{1,2,3\}$ in this study and find the performance quickly saturates at $N=3$.
-Each order in telescoping provides $N_{steps} \times 2$ variables in addition to the subjet axis information such as $\eta, \phi$, i.e for each step in $\Delta r$, we calculate the subjet transverse momenta and the mass.
-Thus, third order TD includes by definition, all variables from the second and first order TD expansion of a jet.
-These telescoping variables are fed into a multi layer perceptron, much like the one we used in the physics variables discussion, with two hidden dense layers and a final output layer and the activation and optimization functions.
+Each telescoping deconstruction order has $N_{\rm steps} \times 2$ variables of subjet transverse momenta and subjet masses, in addition to the subjet axis information $\eta, \phi$.
+%Thus, third order TD includes by definition, all variables from the second and first order TD expansion of a jet.
+These telescoping variables are cumulatively combined and input into a multi layer perceptron, much like the one we used in the previous multivariate analysis, with two hidden dense layers and a final output layer of activation and optimization functions.
 
-Beside the overall quark gluon classification performance using telescoping deconstruction, subjet variables within this framework can reveal many aspects of jet dynamics.
-In particular, there is rich information contained in subjet topology and substructure.
-Recently, the CMS~\cite{} and STAR~\cite{} collaborations measured the groomed momentum sharing observable $z_g$.
-One reclusters jets using the Cambridge/Aachen algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and sequentially removes wide angle, soft radiation until the following condition is satisfied by a branching,
+Beside the overall classification performance using telescoping deconstruction, subjet variables within this framework can reveal many aspects of jet dynamics.
+In particular, there is rich information contained in subjet topology and subjet substructure.
+Recently, the CMS~\cite{Sirunyan:2017bsd} and STAR~\cite{Kauder:2017mhg} collaborations measured the groomed momentum sharing observable $z_g$.
+One reclusters jets using the Cambridge/Aachen (C/A) algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and sequentially removes wide angle, soft radiation until the following condition is satisfied by a branching,
 \begin{equation}
@@
 In vacuum, the observable $z_g$ is closely related to the Altarelli-Parisi splitting functions~\cite{Altarelli:1977zs}.
-The physical meaning of $z_g$ in heavy ion collisions have also been investigated \cite{Chien:2016led}.
+The physical meaning of $z_g$ in heavy ion collisions have also been actively investigated \cite{Chien:2016led,Mehtar-Tani:2016aco,Milhano:2017nzm,Chang:2017gkt,Li:2017wwc}.
 
-We use telescoping deconstruction to analyze the two soft-drop branches defining the $z_g$ observable.
-At the T2 order, we choose the subjet radius to be half the angle between the two deconstruction axes for subjet reconstruction.
-The distribution of the momentum fraction $z_{\rm T2}$ of the softer subjet is significantly different from the $z_g$ distribution, especially for gluon jets.
-We remind the reader that the axes used in the telescoping deconstruction procedure are defined with the Winner-Take-All scheme which favors energetic particles.
-We find that the axes at the T2 order may both align with the hard branch and are not able to efficiently tag the soft branch when it consists of soft particles.
-The situation is significantly improved at the T3 order where we use three axes to capture the soft branch in soft-drop.
-We combine two closest subjets, whilst applying the grooming criteria on the soft subjet and the distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark (left) and gluon (right) jets are very similar as shown in Fig::~\ref{fig:comp_z}.
-The solid and dotted lines represent the soft dropped $z_{g}$ in pp and PbPb where we can see an enhancement of asymmetrical jets in PbPb.
-The subjet fraction calculated from the TD third order (T3) for both quark and gluon jets are less modified due to quenching (comparing the dashed with the dot-dashed in Fig:~\ref{fig:comp_z}) as compared with the softdropped zg.
-
-\begin{figure}[h]
+\begin{figure}[t]
 	   \centering
@@
 
-In addition to the momentum fraction of the subjets, there is also recent interest in measuring the angular separation between the two subjets.
-We show the calculation of the this angular distance between the leading and subleading subjets in Fig:~\ref{fig:comp_rg} for quark (left) and gluon (right) jets.
-For the $R{g}$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
-While this particular observable has not yet been experimentally measured at the LHC or the RHIC, the phenomenon of broadening or widening of the jet observed in these MC calculations are consistent with measurements of the jet shape~\cite{}.
-Both the softdrop $R_{g}$ and the T3 $\theta$ show deviations due to quenching, but the quenched distributions do not match up with each other.
+We use telescoping deconstruction to analyze the two soft-drop branches defining the $z_g$ observable.
+As a first attempt, at the T2 order we choose the subjet radius proportional to the angle between the two deconstruction axes for subjet reconstruction.
+The proportionality constant $C$ is chosen to be smaller than $0.5$ in order for the two subjets be separated in angle.
+We impose a similar condition for the momentum fraction $z_{\rm T2}$ of the softer subjet,
+\begin{equation}
+    z_{\rm cut} < \frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}} \equiv z_{\rm T2}\;,
+\label{T2_z}
+\end{equation}
+where $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two exclusively constructed subjets.
+The resulting $z_{\rm T2}$ distribution is significantly different from the $z_g$ distribution, especially for gluon jets.
+As a reminder, the axes used in the telescoping deconstruction procedure are defined with the winner-take-all scheme which favors energetic particles.
+We find that the axes determined at the T2 order may both align with the hard soft-drop branch and are not able to efficiently tag the soft branch when it consists of soft particles.
+
+The situation is significantly improved at the T3 order where three axes are used to capture the soft branch in soft-drop.
+Similarly, the subjet radii are chosen to be proportional to the angles among three axes.
+We combine two closest subjets and construct the momentum fraction of the resulting two subjets with a similar kinematic cut,
+\begin{equation}
+    z_{\rm cut} < \frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}} \equiv z_{\rm T3}\;.
+\label{T3_z}
+\end{equation}
+This essentially reclusters the three-subjet system using the C/A algorithm and selects the hard branching.
+The distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark (left) and gluon (right) jets are very similar as shown in Figure~\ref{fig:comp_z}.
+The solid and dotted lines represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of asymmetric jets in PbPb collisions.
+In comparison, the soft subjet momentum fraction at the T3 order for both quark and gluon jets is less modified by quenching (see the comparison between the dashed and the dot-dashed lines in Figure~\ref{fig:comp_z}).
+
+In addition to the subjet momentum fraction, there is also recent interest in measuring the angular separation between the two subjets.
+Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ and the angle $\theta_{\rm T3}$ between the two subjets for quark (left) and gluon (right) jets.
+
+For the $r_g$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
+While this particular observable has not yet been experimentally measured at the LHC or the RHIC, the phenomenon of broadening of the jet observed in these MC calculations are consistent with measurements of the jet shape~\cite{}.
+Both the soft-drop $r_g$ and the T3 $\theta$ show deviations due to quenching, but the quenched distributions do not match up with each other.
 Softdrop shows a narrow peak for quenched jets at 0.3, where as the T3 distribution is broader since the sub-jet axis definition are not the same (Softdrop uses C/A whereas TD uses WTA).
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	   \centering
@@
 \label{fig:comp_rg}
+\end{figure}
+
+\begin{figure}[t]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_subjet_m.pdf}
+	   \caption{Leading subjet mass for Quark (left) and Gluon (right) jets in pp (solid) and  quenched PbPb (dotted) in comparison with the subleading subjet mass in pp (dashed) and quenched PbPb (dot dashed) lines.
+}
+\label{fig:comp_subjet_m}
 \end{figure}
@@
 The quenched softer subjet as expected has a much larger tail towards larger masses indication its constituent distribution is spread out over a larger area as opposed to the quenched harder subjet.
-
-\begin{figure}[h]
-	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_subjet_m.pdf}
-	   \caption{Leading subjet mass for Quark (left) and Gluon (right) jets in pp (solid) and  quenched PbPb (dotted) in comparison with the subleading subjet mass in pp (dashed) and quenched PbPb (dot dashed) lines.
-}
-\label{fig:comp_subjet_m}
-\end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\centering
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\centering
         \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
-        \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
+        \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
 	\caption{(Left) ROC curves for quark vs gluon efficiency with the TD method for grid jets in pp (dotted lines) and PbPb (solid lines) (w/ GridSub).
@@
         \item At order $N$, determine $N$ axes $\{\hat n_i\}=\{(\eta_i,\phi_i)\}$ along the dominant energy flows, where $y$ and $\phi_i$ are the rapidity and the azimuthal angle of the subjet axis $i$.
-We use the Winner-Take-All (WTA) $k_T$ axes \cite{Thaler:2010tr}.
+We use the Winner-Take-All (WTA) $k_T$ axes \cite{Bertolini:2013iqa}.
         \item Construct $N$ subjets with $M$ radii $\{R_{T,m}\}^M_{m=1}$ by assigning particles to the nearest axis according to the distance $d^2_{ij} = \Delta y_{ij}^2+\Delta \phi_{ij}^2$ between the axis $\hat n_i$ and the particle $j$ \cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}.

### 2018-02-20 qg_ML.tex
YT updated draft before Lund diagram section

@@
 We exploit jet features to highlight the differences between quark jets and gluon jets and how differently they are modified in the medium.
-We perform quark-gluon and proton-heavy ion discriminations for jets in proton-proton and heavy ion collisions.
+We perform quark-gluon and proton-heavy ion jet discriminations for jets in proton-proton and heavy ion collisions.
 Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are examined in a multivariate analysis.
@@
 Since jet substructure depends strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments.
-It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda}.
+It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
 With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the change of quark and gluon jet fractions caused by jet-medium interactions.
@@
 They include the jet mass, two radial moments including the girth, the $p_T^D$ and the pixel multiplicity.
-Subsets of the variables are combined and studied in a multilayer perceptron (MLP) with two hidden layers to extract correlations among the variables.
+Subsets of the variables are combined and studied in a multi-layer perceptron (MLP) with two hidden layers to extract correlations among the variables.
 We also employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN).
@@
 Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders.
-The details about the telescoping deconstruction procedures are provided in the Appendix, including the applications on boosted $W$ and top tagging in high energy physics.
+Details about the telescoping deconstruction procedures are provided in the Appendix, including the applications to boosted $W$ and top tagging in high energy physics.
 
@@
 %Thus, third order TD includes by definition, all variables from the second and first order TD expansion of a jet.
-These telescoping variables are cumulatively combined and input into a multi layer perceptron, much like the one we used in the previous multivariate analysis, with two hidden dense layers and a final output layer of activation and optimization functions.
+These telescoping variables are cumulatively combined and input into a multi-layer perceptron, much like the one we used in the previous multivariate analysis, with two hidden dense layers and a final output layer of activation and optimization functions.
 
-Beside the overall classification performance using telescoping deconstruction, subjet variables within this framework can reveal many aspects of jet dynamics.
+Beside the overall classification performance using telescoping deconstruction which we will present in the next section, subjet variables within this framework can reveal many aspects of jet dynamics.
 In particular, there is rich information contained in subjet topology and subjet substructure.
@@
 Here $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two branches, and $z_{\rm cut}$ is the soft-drop parameter which is set to $0.1$ in our analysis.
-Note that the Cambridge/Aachen clustering tree is angular-ordered, therefore the above procedure identifies the most wide angle, soft subjet which carries a significant fraction of the jet transverse momentum.
+Note that the Cambridge/Aachen clustering tree is angular-ordered, therefore the above procedure identifies the most wide angle, soft subjet which carries a significant fraction of the jet transverse momentum larger than $z_{\rm cut}$.
 The size of the soft subjet is dynamically determined and related to the angle $r_g$ between the two soft-drop branches, defined as the groomed jet radius.
@@
 The proportionality constant $C$ is chosen to be smaller than $0.5$ in order for the two subjets be separated in angle.
-We impose a similar condition for the momentum fraction $z_{\rm T2}$ of the softer subjet,
+We impose a similar condition for the momentum fraction $z_{\rm T2}$ of the softer subjet to be larger than $z_{\rm cut}$,
 \begin{equation}
@@
 \end{equation}
-where $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two exclusively constructed subjets.
-The resulting $z_{\rm T2}$ distribution is significantly different from the $z_g$ distribution, especially for gluon jets.
-As a reminder, the axes used in the telescoping deconstruction procedure are defined with the winner-take-all scheme which favors energetic particles.
-We find that the axes determined at the T2 order may both align with the hard soft-drop branch and are not able to efficiently tag the soft branch when it consists of soft particles.
+where $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two exclusively constructed subjets at the T2 order.
+The resulting $z_{\rm T2}$ distribution is quite different from the $z_g$ distribution, especially for gluon jets.
+Note that the axes used in the telescoping deconstruction procedure are defined with the winner-take-all scheme which favors energetic particles.
+We find that the axes determined at the T2 order may both align with the hard soft-drop branch and are not able to efficiently tag the soft branch when it consists mainly of soft particles.
 
-The situation is significantly improved at the T3 order where three axes are used to capture the soft branch in soft-drop.
-Similarly, the subjet radii are chosen to be proportional to the angles among three axes.
-We combine two closest subjets and construct the momentum fraction of the resulting two subjets with a similar kinematic cut,
+The situation is significantly improved at the T3 order where three axes $\{\hat n_i\}$ are used to capture the soft branch in soft-drop.
+Similarly, the subjet radii $R_i$ for the $i$-th subjet are chosen to be proportional to the angles among the three axes: $R_i = C \max_{j} \{\theta_{ij}\}$ where $\theta_{ij}=\cos^{-1}(\hat n_i\cdot\hat n_j)$.
+We combine the two closest subjets and construct the momentum fraction of the softer of the resulting two subjets with a similar kinematic cut,
 \begin{equation}
-    z_{\rm cut} < \frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}} \equiv z_{\rm T3}\;.
+    z_{\rm cut} < \frac{\min(p_{T_1},p_{T_2})}{p_{T_1}+p_{T_2}} \equiv z_{\rm T3}\;,
 \label{T3_z}
 \end{equation}
-This essentially reclusters the three-subjet system using the C/A algorithm and selects the hard branching.
-The distributions of the momentum fraction $z_{\rm T3}$ of the softer subjet for quark (left) and gluon (right) jets are very similar as shown in Figure~\ref{fig:comp_z}.
-The solid and dotted lines represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of asymmetric jets in PbPb collisions.
-In comparison, the soft subjet momentum fraction at the T3 order for both quark and gluon jets is less modified by quenching (see the comparison between the dashed and the dot-dashed lines in Figure~\ref{fig:comp_z}).
+as well as the angle between the two subjets $\theta_{\rm T3} > \Delta$ where $\Delta$ is chosen to be 0.1 \cite{Sirunyan:2017bsd}.
+This identifies the kinematics of the branching at the root of the C/A clustering of the three subjets.
+The distributions of the momentum fraction $z_{\rm T3}$ for quark (left) and gluon (right) jets are very similar as shown in Figure~\ref{fig:comp_z}.
+The solid and dotted lines represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy-simulated PbPb collisions \cite{Milhano:2017nzm}.
+On the other hand, the momentum fraction $z_{\rm T3}$ for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
+See the comparison between the dashed and the dot-dashed lines in Figure~\ref{fig:comp_z}.
 
-In addition to the subjet momentum fraction, there is also recent interest in measuring the angular separation between the two subjets.
-Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ and the angle $\theta_{\rm T3}$ between the two subjets for quark (left) and gluon (right) jets.
+In addition to the subjet momentum fraction which provides information about the longitudinal momentum distribution, the angular separation between the two subjets encodes information about the momentum distribution transverse to the jet direction.
+Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ and the angle $\theta_{\rm T3}$ for quark (left) and gluon (right) jets.
+Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
+The observables both have similar distributions for quark and gluon jets.
+However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of jets, while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude.
+Note the difference in the subjet axis definitions which affects the angles among them: soft-drop uses C/A axes whereas TD uses WTA axes.
+Also, the subjet radii in TD provide an extra handle on the area of coverage by subjets therefore soft radiation can be probed with various angular resolutions.
+While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jwpy Monte Carlo simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
 
-For the $r_g$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
+%For the $r_g$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
 While this particular observable has not yet been experimentally measured at the LHC or the RHIC, the phenomenon of broadening of the jet observed in these MC calculations are consistent with measurements of the jet shape~\cite{}.
@@
 
+In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are quite insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
+They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
+To go beyond to higher-order jet features, we also study the subjet substructure, for example subjet masses, for both the subjets.
+These observables disentangle the subjet kinematic information and are more directly sensitive to the soft radiation around subjets.
+Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (solid line: pp, dotted line: PbPb) and soft subjets (dashed line: pp, dot-dashed line: PbPb) for quark (left) and gluon (right) jets.
+Here hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
+We see hints that the $\lambda$ distributions of hard subjets exhibit flavor dependence, while those of soft subjets are similar for quark and gluon jets.
+This is consistent with the physics picture that for a quark jet the hard subjet is typically a quark-initiated subjet, while the soft subjet is a gluon-initiated subjet through $q\rightarrow qg$ splitting.
+On the other hand, both the subjets of a gluon jets are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
+Also, due to the effect of quenching, hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets.
+This suggests the qualitative feature of the soft event activities in the\jwpy simulations, which persists to affect not only the subjet topology but also the subjet substructure.
+
+The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion which systematically improves quark gluon discrimination and gives insights on the jet quenching mechanisms implemented in Monte Carlo simulations.
+Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
+
+%The first point of difference is seen , i.e.
+the pp hard and soft subjet mass are comparable with the softer subjet having a larger tail.
+The quenched softer subjet as expected has a much larger tail towards larger masses indication its constituent distribution is spread out over a larger area as opposed to the quenched harder subjet.
+
 \begin{figure}[t]
@@
 
-We also study the subjet mass for both the subjets that represent the two major prongs of our jet.
-The subjet mass distribution for hard (solid - pp, dotted - PbPb) and soft (dashed - pp, dot-dashed - PbPb) is shown for quark (left) and gluon (right) jets in Fig:~\ref{fig:comp_subjet_m}.
-In this particular case, hard or soft refers to the subjet's momentum with harder meaning the one with higher transverse momentum.
-The first point of difference is seen due to the effect of quenching, i.e.
-the pp hard and soft subjet mass are comparable with the softer subjet having a larger tail.
-The quenched softer subjet as expected has a much larger tail towards larger masses indication its constituent distribution is spread out over a larger area as opposed to the quenched harder subjet.
-
-WRITE THE CONCLUSION HERE FOR THIS PARTICULAR SECTION - WHAT DO WE LEARN
-
 \section{Quark and Gluon Jet Modification and Discrimination}
@@
 
-We discussed jet representations using multiple physics-motivated variables, jet images and telescoping deconstruction.
-In this section we will use these information in the study of jet modification and quark gluon discrimination.
+We discussed jet representations using multiple physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
+In this section we will use these information collectively and exhaustively in the study of jet modification and quark gluon discrimination, employing multivariate methods.
+
 For quark gluon discrimination, the performances are shown using Receiver Operating Characteristic (ROC) curves which plot the gluon jet efficiency as a function of the quark jet efficiency.
-Each ROC curve is labeled in the following figures with the corresponding area under the curve (auc) where the best performance is noted by large values of quark efficiency with small values of gluon efficiency leading to a small auc.
+Each ROC curve is labeled with the corresponding area under the curve (auc).
+Note that a smaller auc corresponds to a better performance, with larger values of quark efficiency and smaller values of gluon efficiency.
 
@@
 	\includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
-	\caption{ROC curves for quark vs gluon efficiency for particle jets in pp (left) and PbPb (right).
-The different colored curves show the variety of classification models studied in this paper and the quantity in the brackets marked by auc represents the area under the curve.
-The green solid and dotted lines are for the MLP with physics variables w/ and w/o the pixel multiplicity.
-The blue solid line is with the DCNN on jet images and the brown, orange and teal colors are for the TD with T1, T2 and T3 orders respectively.
-(color online)}
+	\caption{ROC curves of quark v.s.
+gluon efficiency for particle jets in pp (left panel) and PbPb (right panel) collisions at $\sqrt s=2.76$ TeV.
+The different colored curves correspond to various classification models studied in this paper, and the quantity in the brackets marked by auc represents the area under the curve.
+The green solid and dotted lines are for the MLP with physics-motivated variables with and without the pixel multiplicity, respectively.
+The blue solid line corresponds to the DCNN trained on jet images, while the brown, orange and teal lines are for the TD at T1, T2 and T3 orders, respectively.}
 \label{fig:ROC_all}
@@
 
-Figure:~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and PbPb (right panel) collisions for particle jets i.e jets reconstructed with particle momenta without $\eta=\phi$ discretization and background subtraction.
-We see an overall compatibility amongst the methods used leading to similar classification performances, implying that the three jet representations capture sufficient jet information.
-Note that there is monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$ to $\rm T1+T2+T3$ with increasing input information of higher-order subleading subjets.
-Is it important to note that the TD classification performance saturates quickly at the T2 order with additional higher order offering very little improvement in the quark efficiency given a gluon efficiency.
-In the PbPb case (right of Fig:~\ref{fig:ROC_all}), the MLP with physics variables gives better performance due to the high particle multiplicity in the \jwpy-simulated PbPb events.
-Measuring the jet multiplicity in heavy ion collisions is very tricky due to the underlying and its correlations with the jet.
-In nature of this difficulty, we also show the ROC curve for the MLP with physics variables without jet multiplicity denoted by the dotted green line in Fig:~\ref{fig:ROC_all} which reduces the overall auc by $0.02$ and becomes consistent with our DCNN.
-Even though the TD with radius scan in steps of 0.08 is unable to accurately capture the full multiplicity information at the T3 order it still within a few percent difference of the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
-In order to directly compare with data, we apply the grid discretization for both pp and PbPb with its specific GridSub background subtraction technique and compare the various orders of TD on the left of Fig:~\ref{fig:ROC_TD_grid}.
+Figure~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and PbPb (right panel) collisions for particle jets, i.e.
+jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction.
+We see an overall compatibility among the methods with similar classification performances, implying that the three jet representations capture comparable jet information.
+However, there are noticeable performance differences.
+Note the monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$, to $\rm T1+T2+T3$ with increasing amount of input information of higher-order subleading subjets.
+It is important to note that the TD classification performance saturates quickly at the T2 order, with additional higher-order features offering little improvement in the quark efficiency given a gluon efficiency.
+In pp collisions (left panel of Figure~\ref{fig:ROC_all}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
+This implies that in the pp environment the $0.08\times 0.08$ $\eta-\phi$ discretization still keeps much jet information, and the TD performance quickly saturates to the DCNN performance.
+In PbPb collisions (right panel of Figure~\ref{fig:ROC_all}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
+Moreover, the $\eta-\phi$ discretization results in a poorer DCNN performance than TD with radius scan in steps of 0.08.
+Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets.
+With this difficulty, we also show the ROC curve (the dotted green line in the right panel of Figure~\ref{fig:ROC_all}) for the MLP with physics-motivated variables except for the pixel multiplicity, which reduces the auc by $0.02$ and has a comparable performance as DCNN.
+Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
+Comparing left and right panels we see hints of a degradation of the classification performance in PbPb collisions, while the huge underlying event background may play a role.
 
@@
 
-We clearly see that the performances drops in PbPb collisions along with a reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
-This suggests that the information carried in higher-order subleading subjets can be washed out in heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions in \jwpy.
-On the right of Fig:~\ref{fig:ROC_TD_grid} we plot the TD ROC curves for jet with inputs with or without $\eta-\phi$ discretization and background subtraction.
-Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance.
-On the other hand, the performance drops with the background subtraction performed (slight increase in the auc).
-This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction quantitatively bringing the two flavors of jets together in their fragmentation basis.
+In order to have a fair comparison between jets in pp and PbPb collisions and directly compare with data, we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in the left panel of Figure~\ref{fig:ROC_TD_grid}.
+We clearly see that the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
+This suggests that the information carried in higher-order subleading subjets can be washed out in \jwpy-simulated heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions.
+In the right panel of Figure~\ref{fig:ROC_TD_grid} we plot the TD ROC curves using jets with or without $\eta-\phi$ discretization and background subtraction.
+Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance using particle jets or grid jets.
+On the other hand, the performance drops with the background subtraction performed.
+This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction, bringing the two flavors of jets together in their fragmentation basis.
 
@@
 For a given jet flavor, i.e.
-quark or gluon jet, one can also study with the various aforementioned methods, how the medium modifies it by comparing a vacuum jet with a quenched jet.
-Figure:~\ref{fig:ROC_qq_gg} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
-We see huge difference between jets and all methods perform significantly better at the task of identifying if a jet is quenched or not as opposed to if its a quark or gluon jet and therefore we plot the ROC curve using log scale in the y-axis.
-We observe that the increase of the pixel multiplicity is a key feature in aiding the MLP in identifying quenched jets with the MLP only slightly worse than the DCNN.
-The better performance in the gluon jet discrimination (right panel of Fig:~\ref{ROC_qq_gg}) suggests that gluon jets are significantly more modified than quark jets due to medium induced quenching.
+quark or gluon jets, one can also study with various aforementioned methods how the medium modifies jets by comparing vacuum jets with quenched jets in the context of proton-heavy ion jet discrimination.
+Figure~\ref{fig:ROC_qq_gg} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
+We see huge difference between jets in pp and \jwpy-simulated PbPb collisions.
+All the methods perform significantly better in the task of identifying if a jet is quenched or not, compared to quark gluon discrimination.
+Therefore we plot the ROC curves using the log scale in the $y$-axis.
+We observe that the increase of the pixel multiplicity is a key feature in identifying quenched jets, with the DCNN performs slightly better than the MLP.
+The better performance in the proton-heavy ion jet discrimination for gluon jets (right panel of Figure~\ref{fig:ROC_qq_gg}) suggests that gluon jets are modified more significantly than quark jets due to medium-induced quenching.
 
-From our studies of the ROC curve in comparing quarks vs gluons and in comparing vacuum vs quenched jets, we can begin to qualitatively understand the effect of quenching as implemented in the \jw MC.
-Quenched jets end up with significantly more jet constituents which are distributed away from the principal axis of the jet as seen in the saturation of TD classification performance from first to third order.
-This modification appears to affect gluon jets significantly more than quark jets since a gluon jet with its increased multiplicity and wider parton shower due to the color factor, can all undergo quenching, or in the case of \jw, interact with the scattering centers leading to a broader more quenched jet.
-For a systematic study of such differences in quenched quark vs gluon, we employ the use of Lund Diagram as we shall discuss in the next section.
+From our studies of the ROC curves for comparing quark jets and gluon jets, as well as comparing vacuum jets and quenched jets, we begin to qualitatively understand the effects of quenching as implemented in the \jwpy Monte Carlo.
+Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis as seen in the saturation of the TD classification performance from T1 to T3 order.
+This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can all interact with the scattering centers leading to broader and more quenched jets.
+In the next section we will employ the use of Lund Diagrams to study the differences in quenched quark and gluon jets.
 

### 2018-02-21 qg_ML.tex
YT updated draft.

@@
 quark or gluon jets, one can also study with various aforementioned methods how the medium modifies jets by comparing vacuum jets with quenched jets in the context of proton-heavy ion jet discrimination.
-Figure~\ref{fig:ROC_qq_gg} shows the ROC curves for discriminating quark jets (left panel) and gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
-We see huge difference between jets in pp and \jwpy-simulated PbPb collisions.
+Figure~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficency as a function of the pp jet efficiency, for discriminating quark jets (left panel) or gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
+We see huge differences between jets in pp and \jwpy-simulated PbPb collisions.
 All the methods perform significantly better in the task of identifying if a jet is quenched or not, compared to quark gluon discrimination.
@@
 
-From our studies of the ROC curves for comparing quark jets and gluon jets, as well as comparing vacuum jets and quenched jets, we begin to qualitatively understand the effects of quenching as implemented in the \jwpy Monte Carlo.
-Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis as seen in the saturation of the TD classification performance from T1 to T3 order.
+From our studies of the ROC curves for comparing quark jets and gluon jets, as well as comparing vacuum jets and quenched jets, we understand qualitatively in more details the effects of quenching as implemented in the \jwpy Monte Carlo.
+Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order.
 This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can all interact with the scattering centers leading to broader and more quenched jets.
-In the next section we will employ the use of Lund Diagrams to study the differences in quenched quark and gluon jets.
+In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions
 
@@
 
-Any jet observable on the market is designed as some function of the momentum and the distance scale that is used in the jet reconstruction.
-These two scales can be represented as $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$, where $z$ is the transverse momentum fraction of the softer subjet relative to sum of the two subjets, $\Delta R$ is the angular scale between the two subjets in a given branch and $R$ is the jet radius.
-Such a representation of a jet is known as the Lund Diagram~\cite{Salam:2016yht} and the prescription to create such a diagram for a given jet is as follows:
+Each jet observable is defined as a function of the momenta and angles of jet constituents, which probes the entire jet formation with a single number.
+One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches all faithfully encode the jet information.
+In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
+It has also been recently used in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
+The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling limit for jets in pp collisions because of the infrared structure of QCD.
+
+Given a jet reconstructed using the anti-$k_{T}$ algorithm with $R = 0.4$, the procedure of creating the Lund diagram is as follows:
 \begin{itemize}
-	\item Cluster your jet with a jet R say $0.4$ and a particular jet algorithm, in our case anti-k$_{t}$
-	\item Take the jet constituents and re-cluster with the C/A algorithm which combines the nearest particles with each other
-	\item Walk backwards along the clustering tree and for each branch calculate the $z=p^{2}_{T}/(p^{1}_{T}+p^{2}_{T})$ and $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
-	\item Fill the Lund diagram at each split
-	\item Continue in walking backwards along the hard branch till you reach only one particle
+	\item Recluster the jet using the C/A algorithm
+	\item Start from the root of the tree and move along the hard branches.
+At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
+	\item Fill in the Lund diagram for each branching
+	\item Continue along the hard branches until reaching only one particle
 \end{itemize}
@@
 
-The beauty of a Lund diagram is its convenient separation of wide soft angle radiation close to the y-axis, collinear hard radiation along the diagonal and with contribution from hadronization and underlying event in a line parallel to the x-axis.
-These Lund diagrams have also been recently used in distinguishing between QCD and boosted heavy objects due to their unique feature of featuring jets of a constant mass along the main diagonal (opposite to the collinear hard radiation).
-We show the Lund diagrams for quark (left) and gluon (right)jets in pp (top) and PbPb (bottom) in Fig:~\ref{fig:Lund_full} where the z is calculated relative to each split.
-We see an enhanced region in PbPb collisions close to the y-axis in comparison to pp which corresponds to increased wide soft angle in quenched jets.
-Comparing left to right (as we discussed in Fig:~\ref{fig:quenching_discrimination}) one also sees gluon jets include a larger fraction of the soft radiation and these are susceptible to hadronization and underlying event effects.
-The hard collinear radiation region of the Lund diagram, that is dominated in pp is shown to reduce in its significance for quenched jets.
-The core of the jet remains similar between pp and PbPb, but the secondary sub-jet ultimately gets pushed to wider angles and to softer momenta.
-Such a picture is exactly what we discussed in Fig:~\ref{fig:comp-zg}-\ref{fig:comp_subjet_m} but this time, visible via the Lund diagram.
+In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal.
+Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions.
+We see a significantly enhanced region close to the $y$-axis in PbPb collisions which corresponds to increased wide angle, soft radiation in quenched jets.
+Also, gluon jets have a larger fraction of the soft radiation.
+On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
+As we discussed in Figure~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
+Such picture emerges and is directly visible in the Lund diagram.
 
@@
 
-One can now study the effect of grooming with the help of the Lund diagrams by rejecting the soft radiation ($z < 0.1$) that is typically groomed a way the softdrop method.
-Fig:~\ref{fig:Lund_bkg} shows the Lund diagrams for our jets in pp and PbPb  before the one which passes the soft-drop condition.
-We see that the soft, wide-angle radiation is isolated and the hard collinear core is essentially removed.
-The bottom right panel of Fig:~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft wide angle radiation.
-These soft jet constituents in the jet's periphery are precisely the ones that show up as a modification of the jet shape.
-Within the framework of this diagrammatic representation, one can now build up these observables and follow the harder branch leading from the first split.
-Fig:~\ref{fig:Lund_hard} shows the Lund diagrams after the soft branch (background) has been removed via the soft-drop procedure.
-We see that the vacuum structure starts to emerge in the soft-drop PbPb jets.
-However, soft-drop can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation.
-We understand this behavior as the correlated medium response that belongs to the jet but is often removed due to a statistical treatment of background in experiments.
+We can also directly study the effect of soft-drop using the Lund diagram.
+Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark and gluon jets in pp and PbPb collisions, with only the branchings before the one defining the $z_g$ and $r_g$ observables.
+That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
+We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
+The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
+These soft jet constituents near the jet boundary are precisely the ones contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
+Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
+We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
+However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation.
+This can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
 
@@
 
-We have shown a systematic study of modification to the jet structure due to the effects of quenching in heavy ion collisions as implemented within the \jw MC generator.
-The medium modification is studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets.
-We compare a MLP with physics motivated jet variables and a state of the art DCNN applied on jet images to a our method of Telescoping Deconstruction and find the classification performance worsens for quenched jets when compared with vacuum pp jets.
-All of the models studied perform consistently with each other giving us the confidence that the TD to third order is extracting fundamental fragmentation patters of a given jet.
+We present a systematic study of modifications to quark and gluon jet structure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
+Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems.
+We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction.
+We find that the quark gluon classification performance worsens for quenched jets.
+All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns.
+Through multiple methods and observables we observe consistently that the dominant feature in the \jwpy-simulated PbPb collisions is the increase of soft particle multiplicity throughout the jet region.
+This results in the lost of information in subleading subjets which is a characteristic feature of the \jw simulation.
 
-DISCUSSION - THIS IS A VERY IMPORTANT PART
+The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations.
+Therefore our work serves as an example how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
+The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization.
+In the future the comparisons between theoretical calculations and simulations \cite{}, and with experimental measurements, will become standard practice to identify robust features of jet-medium interactions.
+Our work suggests a systematically improvable framework which will serve this purpose.
 
@@
 The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix.
-The authors also thank Benjamin Nachman, Joern Putschke, Konrad Tywoniuk and Sevil Salur for helpful discussions.
-We thank Gavin Salam and the organizers of CERN jet workshop (2017) for bringing to our attention the concept of Lund Diagrams and their usage in studying jet modifications.
+The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Konrad Tywoniuk and Sevil Salur for helpful discussions.
+We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.

### 2018-02-21 qg_ML.tex
YT updated the draft

@@
 We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
-We exploit jet features to highlight the differences between quark jets and gluon jets and how differently they are modified in the medium.
+We exploit jet features to highlight differences between quark jets and gluon jets and how differently they are modified in the medium.
 We perform quark-gluon and proton-heavy ion jet discriminations for jets in proton-proton and heavy ion collisions.
@@
 We draw connections to the soft-drop subjet distribution and illuminate jet modifications using Lund diagrams.
-The quark and gluon jet samples generated from \jwpy ~are used as an example to establish our general method.
+The quark and gluon jet samples generated from \jwpy are used as an example to establish our general method.
 We find that the quark gluon discrimination performance worsens in the \jwpy simulated heavy ion collisions.
-The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy ~simulation.
+The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy simulation.
 Our work suggests a systematically improvable framework for testing and comparing jet simulations and measurements in heavy ion collisions.
@@
 Sirunyan:2017isk,Sirunyan:2017oug,Sirunyan:2017lzi,Sirunyan:2017jic,Sirunyan:2017xss,Sirunyan:2017qhf}
-has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions (see ~\cite{Connors:2017ptx} for a review on heavy ion jet measurements).
+has since become an essential hard probe of the strongly interacting medium produced in heavy ion collisions (see \cite{Connors:2017ptx} for a review on heavy ion jet measurements).
 The dramatic suppression of hadron and jet cross sections is understood using the medium-induced energy loss picture \cite{Gyulassy:1993hr,Wang:1994fx,Zakharov:1996fv,Zakharov:1997uu,Baier:1996kr,Baier:1998kq,
@@
 Arnold:2001ba,Arnold:2001ms,Arnold:2002ja,Casalderrey-Solana:2014bpa}.
-In order to quantitatively extract key features of the jet-medium interaction and the underlying medium dynamics, a precise understanding of the redistribution of jet energy in heavy ion collisions is becoming essential.
+In order to quantitatively extract key features of jet-medium interactions and the underlying medium dynamics, a precise understanding of the redistribution of jet energy in heavy ion collisions is becoming essential.
 With the proliferation of accurate comparisons between theoretical calculations and measurements \cite{Chatrchyan:2013kwa,Khachatryan:2016tfj,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec,Khachatryan:2016erx,
@@
 From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, we have established that both the transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructure depends strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments.
+Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments.
 It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
@@
 Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease.
-On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize the contributions from medium responses \cite{Tachibana:2017syd}.
+On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}.
 Quark and gluon jets will provide the opportunity for studying how differently the medium responds to different probes.
@@
 The lowest-order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
-The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$ respectively, and the larger color charge of gluons results in the broader spread of the radiations inside gluon jets.
+The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons results in broader spread of the radiation inside gluon jets.
 Extra distinctive features can be included to improve the tagging by combining various jet substructure variables in a multivariate analysis, and advances in machine learning techniques have recently been incorporated which helps achieving higher tagging efficiency \cite{Komiske:2016rsd}.
@@
 We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}.
-However, the method is easily applicable to other heavy ion jet simulations \cite{} and experimental data as well.
+However, the method is easily applicable to other heavy ion jet simulations \cite{Armesto:2009fj,Cao:2017zih} and experimental data as well.
 
@@
 Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method using subjets with multiple angular resolutions.
-At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axis which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
+At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axes which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
 The corresponding subjet expansion can be truncated at any finite order which allows us to order-by-order examine the information carried by subleading subjets.
@@
 Within the catchment area of jets the significant background radiation needs to be subtracted before meaningful jet observable measurements can be made.
-Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Berta:2014eza,Komiske:2017ubm} is preferred.
+Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm} is preferred.
 Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables.
@@
 The rest of the paper is organized as follows.
-In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy ~simulation.
-In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as distributions of jet substructure observables.
+In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy simulation.
+In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as the corresponding jet substructure distributions.
 In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations.
-In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space separated by the soft-drop procedure.
+In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space in jet formation separated in the soft-drop procedure.
 We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.
@@
 We use the \jwpy Monte Carlo simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
-\jwpy ~is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
+\jwpy is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
 We use prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
@@
 
-In each event, jets are reconstructed using the anti-$k_{t}$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$.
+In each event, jets are reconstructed using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$.
 We impose the following kinematic cuts,
 \begin{equation}
-    p^{\gamma}_T > 100~{\rm GeV}, ~~~~ p^{\rm jet}_T > 50~{\rm GeV}, ~~~~ |\eta^{\rm jet}|<2, ~~~~\Delta \phi_{\gamma, {\rm jet}} > 2\pi/3\;,
+    p^{\gamma}_T > 100~{\rm GeV}, ~~~ |\eta^{\rm \gamma}|<1.5, ~~~ p^{\rm jet}_T > 50~{\rm GeV}, ~~~ |\eta^{\rm jet}|<1.5, ~~~\Delta \phi_{\gamma, {\rm jet}} > 2\pi/3\;,
 \end{equation}
@@
 The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
-We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHApdf}~\cite{Whalley:2005nh}.
+We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHAPDF}~\cite{Whalley:2005nh}.
 We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons.
 It was shown that recoil contributions are important in \jwpy to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+In order to remove the thermal component of recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 The \jw simulation produces HepMC~\cite{} files which are processed using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.
@@
 This can also help disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments.
-A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Fig:~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
+A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Figure~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
 
@@
 It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features.
-More details about the telescoping deconstruction algorithm and examples comparing QCD, W and top jets are provided in the Appendix.
+More details about the telescoping deconstruction algorithm and examples comparing QCD, $W$ and top jets are provided in the Appendix.
+We will show and discuss the classification performances of all the methods in the next section.
 
@@
 	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets.
-The top panels of each individual distribution show the pp ones and the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
+The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
 	   \label{fig:jetdistributons_pp_pbpb}
@@
 Multivariate analyses have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study.
-Given a jet reconstructed using the anti-$k_{t}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
+Given a jet reconstructed using the anti-$k_{T}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
 	\begin{itemize}
-		\item Jet mass: $m = \sqrt{E^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
-        		\item First radial moment (girth): $\sum_{i \in {\rm jet}} p^i_{T} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
-		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p^i_{T} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
-        		\item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^{i}_{T}}^2}/p^{\rm jet}_{T}$.
+		\item Jet mass: $m = \sqrt{(\sum_{i\in {\rm jet}} E_i)^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
+        \item First radial moment (girth): $\sum_{i \in {\rm jet}} p_{T_i} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
+		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p_{T_i} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
+        \item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^2_{T_i}}}/p^{\rm jet}_{T}$.
         \item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
@@
 \includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
-\caption{Average quark (left) and gluon (right) pre-processed jet images shown in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \jw.
-The color scale in the z axis represents the average pixel \pt in units of GeV/c.
+\caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}.
+The color scale represents the average transverse momenta deposited in each pixel in units of GeV/c.
 }
@@
 The color of each image pixel represents the transverse momentum deposited in the pixel.
-We see the broader energy distribution around the center pixel for gluon jets compared to quark jets in pp collisions, and the significantly broader distributions for jets in PbPb collisions.
+We see the broader energy distribution around central pixels for gluon jets compared to quark jets in pp collisions, and the significantly broader distributions for jets in PbPb collisions.
 Note that in experiment it may be very challenging to directly analyze jet images due to detector and measurement considerations.
@@
 
-A deep convolutional neural network is implemented in Keras with the TensorFlow \cite{} backend.
+A deep convolutional neural network is implemented in Keras with the TensorFlow \cite{DBLP:journals/corr/AbadiBCCDDDGIIK16} backend.
 The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer.
@@
 The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
-The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions.
+The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla,Chien:2017xrb}.
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction and then reaching out to include wide-angle, soft radiation.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_z.pdf}
-	   \caption{Transverse momentum fraction of the sub-leading subjet for Quark (left) and Gluon (right) jets in pp (solid) with soft drop, pp with TD T3 (dashed), quenched PbPb soft drop (dotted) and quenched PbPb TD T3 (dot dashed) lines.
-Subjet fraction is enclosed between 0.1 and 0.5 due to the $z_{cut}$ applied while grooming.}
+	   \caption{Distributions of groomed momentum sharing $z_g$ and TD subjet momentum fraction $z_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
+Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work.}
 \label{fig:comp_z}
@@
 We use telescoping deconstruction to analyze the two soft-drop branches defining the $z_g$ observable.
-As a first attempt, at the T2 order we choose the subjet radius proportional to the angle between the two deconstruction axes for subjet reconstruction.
-The proportionality constant $C$ is chosen to be smaller than $0.5$ in order for the two subjets be separated in angle.
+As a first attempt, at the T2 order we choose the subjet radius $r$ proportional to the angle $\theta$ between the two deconstruction axes for subjet reconstruction: $r=C \theta$.
+The proportionality constant $C$ is set to be 0.3, a number smaller than $0.5$, in order for the two subjets be separated in angle.
 We impose a similar condition for the momentum fraction $z_{\rm T2}$ of the softer subjet to be larger than $z_{\rm cut}$,
@@
 
-The situation is significantly improved at the T3 order where three axes $\{\hat n_i\}$ are used to capture the soft branch in soft-drop.
-Similarly, the subjet radii $R_i$ for the $i$-th subjet are chosen to be proportional to the angles among the three axes: $R_i = C \max_{j} \{\theta_{ij}\}$ where $\theta_{ij}=\cos^{-1}(\hat n_i\cdot\hat n_j)$.
+The situation is significantly improved at the T3 order where three axes $\{\hat n_1,\hat n_2,\hat n_3\}$ are used to capture the soft branch in soft-drop.
+Similarly, the subjet radii $r_i$ for the $i$-th subjet are chosen to be proportional to the angles among the three axes: $r_i = C \max_{j} \{\theta_{ij}\}$ where $\theta_{ij}=\cos^{-1}(\hat n_i\cdot\hat n_j)$.
 We combine the two closest subjets and construct the momentum fraction of the softer of the resulting two subjets with a similar kinematic cut,
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg.pdf}
-	   \caption{Angular distance between the leading and subleading subjets for Quark (left) and Gluon (right) jets in pp (solid) with soft drop, pp with TD T3 (dashed), quenched PbPb soft drop (dotted) and quenched PbPb TD T3 (dot dashed) lines.
-The quenched $R_{g}$ distribution is significantly modified compared the vacuum case.}
+	   \caption{Distributions of groomed jet radius $r_g$ and TD angular separation between subjets $\theta_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
+The quenched $r_{g}$ distribution is significantly modified.}
 \label{fig:comp_rg}
@@
 Also, due to the effect of quenching, hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets.
-This suggests the qualitative feature of the soft event activities in the\jwpy simulations, which persists to affect not only the subjet topology but also the subjet substructure.
+This suggests the qualitative feature of the soft event activities in the \jwpy simulations, which persists to affect not only the subjet topology but also the subjet substructure.
 
-The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion which systematically improves quark gluon discrimination and gives insights on the jet quenching mechanisms implemented in Monte Carlo simulations.
+The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion, and it can systematically improve quark gluon discrimination and give insights on the jet quenching mechanisms implemented in Monte Carlo simulations.
 Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_subjet_m.pdf}
-	   \caption{Leading subjet mass for Quark (left) and Gluon (right) jets in pp (solid) and  quenched PbPb (dotted) in comparison with the subleading subjet mass in pp (dashed) and quenched PbPb (dot dashed) lines.
+	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$, for quark (left) and gluon (right) jets in pp (solid and dashed lines) and centra PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
 }
@@
 
-We discussed jet representations using multiple physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
+We discussed jet representations using multiple, physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
 In this section we will use these information collectively and exhaustively in the study of jet modification and quark gluon discrimination, employing multivariate methods.
@@
 	\caption{ROC curves of quark v.s.
-gluon efficiency for particle jets in pp (left panel) and PbPb (right panel) collisions at $\sqrt s=2.76$ TeV.
+gluon efficiency for particle jets in pp (left panel) and central PbPb (right panel) collisions in \textsc{Jewel}.
 The different colored curves correspond to various classification models studied in this paper, and the quantity in the brackets marked by auc represents the area under the curve.
@@
 In pp collisions (left panel of Figure~\ref{fig:ROC_all}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
-This implies that in the pp environment the $0.08\times 0.08$ $\eta-\phi$ discretization still keeps much jet information, and the TD performance quickly saturates to the DCNN performance.
+This implies that in the pp environment the $\eta-\phi$ discretization still keeps much jet information, and the TD performance quickly saturates to the DCNN performance.
 In PbPb collisions (right panel of Figure~\ref{fig:ROC_all}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
-Moreover, the $\eta-\phi$ discretization results in a poorer DCNN performance than TD with radius scan in steps of 0.08.
+%Moreover, the $\eta-\phi$ discretization results in a poorer DCNN performance than TD with radius scan in steps of 0.08.
 Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets.
-With this difficulty, we also show the ROC curve (the dotted green line in the right panel of Figure~\ref{fig:ROC_all}) for the MLP with physics-motivated variables except for the pixel multiplicity, which reduces the auc by $0.02$ and has a comparable performance as DCNN.
+With this difficulty, we also show the ROC curve (the dotted green line in the right panel of Figure~\ref{fig:ROC_all}) for the MLP with physics-motivated variables except for the pixel multiplicity, which reduces the auc by $0.02$.
+% and has a comparable performance as DCNN.
 Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
@@
         \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
-	\caption{(Left) ROC curves for quark vs gluon efficiency with the TD method for grid jets in pp (dotted lines) and PbPb (solid lines) (w/ GridSub).
-The different colors are for different orders (T1 - blue, T2 - red, T3 - green; color online).
-(Right) ROC curves for quark vs gluon efficiency with the TD model at the third order for pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid no bkg-sub (solid red) and PbPb grid sub (solid blue) jets.
+	\caption{(Left panel) ROC curves of quark v.s.
+gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions with GridSub subtraction.
+The different colors represent different TD orders (T1: blue, T2: red, T3: green).
+(Right panel) ROC curves of quark v.s.
+gluon efficiency using TD at the T3 order for jets with pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid (solid red) and PbPb grid with GridSub subtraction (solid blue) constituents.
 The corresponding area under the curves are shown in the brackets.}
@@
 	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
-	   \caption{ROC curves for a given jet to be pp-like (in vacuum) or PbPb-like (quenched) for quark (left) and gluon jets on the grid (right).
-As before, the MLP with physics variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at first, second and third order are in solid maroon, orange and cyan lines.}
+	   \caption{ROC curves of pp-jet v.s.
+PbPb-jet efficiency for quark (left panel) and gluon (right panel) jets with grid discretization.
+The different colored curves correspond to various classification models studied in this paper.}
+%As before, the MLP with physics-motivated variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at T1, T1+T2 and third order are in solid maroon, orange and cyan lines.}
 \label{fig:ROC_qq_gg}
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams plotted with the relative $z$ of the soft subjet in the harder branch for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
-(See text for definition of the axis)}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
 \label{fig:Lund_full}
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	\caption{Lund diagrams specifically showing subjets that are removed and tagged as background ($z<0.1$) for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
-(See text for definition of the axis)}
+	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+See text for the definition of the axes.}
 \label{fig:Lund_bkg}
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
-	   \caption{Lund diagrams created by following the first hard branch selected by $z>0.1$ for quark (left) and gluon jets (right) in pp (top) and central PbPb simulations with \jw.
-(See text for definition of the axis)}
+	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+See text for the definition of the axes.}
 \label{fig:Lund_hard}

### 2018-02-22 qg_ML.tex
small update to text before pulling
added the new figures to the text, still need to change the captions and text

@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_z.pdf}
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_zg_ppcomp_SD_TD2_new.pdf}
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_ppcomp_SD_TD2_new.pdf}
+	   \caption{Top: Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum  $z_{\rm T2}$ (right); Bottom: Distributions of groomed jet radius $r_g$ and the TD angular separation between subjets $\theta_{\rm T2}$ for of quark (blue) and gluon (red) jets in pp \jw simulations.}
+\label{fig:comp_z}
+\end{figure}
+
+\begin{figure}[t]
+	   \centering
+	   \caption{Distributions of groomed jet radius $r_g$ and TD subjet momentum  $z_{\rm T2}$ of quark (blue) and gluon (red) jets in pp.}
+\label{fig:comp_z}
+\end{figure}
+
+\begin{figure}[t]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_z_new.pdf}
 	   \caption{Distributions of groomed momentum sharing $z_g$ and TD subjet momentum fraction $z_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg.pdf}
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_new.pdf}
 	   \caption{Distributions of groomed jet radius $r_g$ and TD angular separation between subjets $\theta_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_subjet_m.pdf}
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
 	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$, for quark (left) and gluon (right) jets in pp (solid and dashed lines) and centra PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
@@
 We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.
+RKE also thanks for organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.

### 2018-02-24 qg_ML.tex
text update before updating the root files

@@
 	   \caption{Top: Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum  $z_{\rm T2}$ (right); Bottom: Distributions of groomed jet radius $r_g$ and the TD angular separation between subjets $\theta_{\rm T2}$ for of quark (blue) and gluon (red) jets in pp \jw simulations.}
-\label{fig:comp_z}
+\label{fig:comp_z_pp_wT2}
 \end{figure}
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_z_new.pdf}
-	   \caption{Distributions of groomed momentum sharing $z_g$ and TD subjet momentum fraction $z_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
-Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work.}
+	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD (right) subjet momentum fraction $z_{\rm T3}$ of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
+Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work.
+(color online)}
 \label{fig:comp_z}
@@
 where $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two exclusively constructed subjets at the T2 order.
-The resulting $z_{\rm T2}$ distribution is quite different from the $z_g$ distribution, especially for gluon jets.
+The resulting $z_{\rm T2}$ distribution is quite different from the $z_g$ distribution, especially for gluon jets as shown in the top panels of Fig:~\ref{fig:comp_z_pp_wT2}.
 Note that the axes used in the telescoping deconstruction procedure are defined with the winner-take-all scheme which favors energetic particles.
@@
 as well as the angle between the two subjets $\theta_{\rm T3} > \Delta$ where $\Delta$ is chosen to be 0.1 \cite{Sirunyan:2017bsd}.
-This identifies the kinematics of the branching at the root of the C/A clustering of the three subjets.
-The distributions of the momentum fraction $z_{\rm T3}$ for quark (left) and gluon (right) jets are very similar as shown in Figure~\ref{fig:comp_z}.
-The solid and dotted lines represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy-simulated PbPb collisions \cite{Milhano:2017nzm}.
-On the other hand, the momentum fraction $z_{\rm T3}$ for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
-See the comparison between the dashed and the dot-dashed lines in Figure~\ref{fig:comp_z}.
+The angular separation at the second order on the other hand is also able to reproduce the softdrop groomed jet radius as shown in the bottom panels of Fig:~\ref{fig:comp_z_pp_wT2}.
+This identifies the kinematics of the branching at the root of the C/A clustering of the two/three subjets.
+The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Figure~\ref{fig:comp_z}.
+The top-left and bottom-left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy-simulated PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+On the other hand, the momentum fraction $z_{\rm T3}$ (right panels) for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
 
 In addition to the subjet momentum fraction which provides information about the longitudinal momentum distribution, the angular separation between the two subjets encodes information about the momentum distribution transverse to the jet direction.
-Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ and the angle $\theta_{\rm T3}$ for quark (left) and gluon (right) jets.
+Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
 Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
-The observables both have similar distributions for quark and gluon jets.
-However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of jets, while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude.
+The observables both have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Fig:~\ref{fig:comp_rg}.
+However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of the quenched jets (bottom panels), while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude.
 Note the difference in the subjet axis definitions which affects the angles among them: soft-drop uses C/A axes whereas TD uses WTA axes.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_new.pdf}
-	   \caption{Distributions of groomed jet radius $r_g$ and TD angular separation between subjets $\theta_{\rm T3}$ of quark (left) and gluon (right) jets in pp (solid and dashed lines) and central PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
-The quenched $r_{g}$ distribution is significantly modified.}
+	   \caption{Distributions of groomed jet radius $r_g$ (left) and TD angular separation between subjets $\theta_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
+The quenched $r_{g}$ distribution is significantly modified with quenched jets having a larger sub-structure at the edges of the jet axis.}
 \label{fig:comp_rg}
@@
 These observables disentangle the subjet kinematic information and are more directly sensitive to the soft radiation around subjets.
-Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (solid line: pp, dotted line: PbPb) and soft subjets (dashed line: pp, dot-dashed line: PbPb) for quark (left) and gluon (right) jets.
+Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (left) and gluon (right) jets.
 Here hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
@@
 On the other hand, both the subjets of a gluon jets are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
-Also, due to the effect of quenching, hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets.
+Also, due to the effect of quenching (bottom panels in Fig:~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
 This suggests the qualitative feature of the soft event activities in the \jwpy simulations, which persists to affect not only the subjet topology but also the subjet substructure.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
-	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$, for quark (left) and gluon (right) jets in pp (solid and dashed lines) and centra PbPb (dotted and dot-dashed lines) collisions in \textsc{Jewel}.
+	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$, for quark (left) and gluon (right) jets in pp (solid and dashed lines) and centra PbPb (dotted and dot-dashed lines) collisions in \jwpy.
 }

### 2018-02-26 qg_ML.tex
text revision and figure updates

@@
 We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
-We exploit jet features to highlight differences between quark jets and gluon jets and how differently they are modified in the medium.
-We perform quark-gluon and proton-heavy ion jet discriminations for jets in proton-proton and heavy ion collisions.
-Several physics-motivated jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity are examined in a multivariate analysis.
-In comparison, we perform the classifications using image-recognition methods by training a deep convolutional neutral network on jet images.
-To systematically extract jet substructure information, we develop the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
-We draw connections to the soft-drop subjet distribution and illuminate jet modifications using Lund diagrams.
-The quark and gluon jet samples generated from \jwpy are used as an example to establish our general method.
-We find that the quark gluon discrimination performance worsens in the \jwpy simulated heavy ion collisions.
-The information carried by subleading subjets can be washed out due to soft event activities, a characteristic feature of the \jwpy simulation.
-Our work suggests a systematically improvable framework for testing and comparing jet simulations and measurements in heavy ion collisions.
+We exploit jet and sub-jet features to highlight differences between quark and gluon jets in vacuum and with the jet-quenching model implemented in \jw.
+We begin with a physics-motivated analysis of jet structure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity examined in a multivariate analysis.
+In comparison, we employ state of the art image-recognition techniques by training a deep convolutional neutral network on jet images.
+To systematically extract high level jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
+We draw connections to the soft-drop subjet distribution and illuminate jet medium induced modifications using Lund diagrams.
+We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity on the softer part of the jet structure.
+Out work suggests a systematically improvable framework for studying quark and gluon classification in direct comparisons of jet simulations with measurements in heavy ion collisions.
 }
@@
 Jet substructure measurements provide concrete and consistent physics information of the modification of jets in a heavy ion environment.
-From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, we have established that both the transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change of the fraction of quark-initiated jets and gluon-initiated jets may contribute to all substructure modifications observed in experiments.
+From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified.
+Since jet substructures and jet fragmentation depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may contribute to a significant fraction of substructure modifications observed in experiments.
 It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
-With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the change of quark and gluon jet fractions caused by jet-medium interactions.
-This then allows one to focus on studying how differently quark and gluon jets are modified and use them as independent jet quenching probes.
+With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the changes in quark and gluon jet fractions caused by jet-medium interactions.
+This then allows one to focus on studying how quark and gluon jets are differently modified and use them as independent jet quenching probes.
 Note that the above jet modification picture implies that, with the high-purity quark jets in photon-tagged jet samples, the narrowing effect of jet cores should decrease.
 On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}.
-Quark and gluon jets will provide the opportunity for studying how differently the medium responds to different probes.
+Selecting pure samples of quark and gluon jets will provide the golden opportunity for studying how the medium responds to different hard probes.
 
-The use of quark jets and gluon jets as different hard probes of heavy ion collisions is closely related to the tagging of these two probes, which is an outstanding problem \cite{Gras:2017jty,Frye:2017yrw}.
+The use of quark jets and gluon jets as independent hard probes of heavy ion collisions is closely related to the tagging of these two probes, which is an outstanding problem \cite{Gras:2017jty,Frye:2017yrw}.
 The lowest-order feature that separates quark jets from gluon jets is the different color charges carried by the jet-initiating partons.
-The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons results in broader spread of the radiation inside gluon jets.
-Extra distinctive features can be included to improve the tagging by combining various jet substructure variables in a multivariate analysis, and advances in machine learning techniques have recently been incorporated which helps achieving higher tagging efficiency \cite{Komiske:2016rsd}.
+The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons is expected to result in broader spread of the radiation inside such jets.
+Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis using a Boosted Decision Tree (BDT), similar to how heavy flavor jets are identified.
+Driven by the need to identify and study boosted object searches at the LHC, the community has had significant advances on the use of machine learning techniques leading to significantly higher tagging efficiency at large kinematic ranges \cite{Komiske:2016rsd, Larkoski:2017jix}.
 
-Besides the conventional wisdom in jet quenching studies by directly comparing jets in proton-proton and heavy ion collisions, we will systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
-This indirect, complimentary approach studies the modification of the differences between quark jets and gluon jets which disentangles modifications of their common features.
+Jet quenching studies of the past fifteen years has made compelling qualitative strides by comparing jets in proton-proton to heavy ion collisions.
+In this paper, we systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
+This indirect, complimentary approach studies the  differences in the modification of quark vs gluon jets thus disentangling modifications of their common features.
 By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements.
@@
 We will use three classes of methods and gain insights from comparing their performances.
-We combine in a multivariate analysis five physics-motivated and representative jet substructure variables which are effective in quark gluon discrimination \cite{Gallicchio:2012ez}.
+We combine in a multivariate analysis, five physics-motivated and representative jet substructure variables which are effective in quark gluon discrimination \cite{Gallicchio:2012ez}.
 They include the jet mass, two radial moments including the girth, the $p_T^D$ and the pixel multiplicity.
-Subsets of the variables are combined and studied in a multi-layer perceptron (MLP) with two hidden layers to extract correlations among the variables.
-We also employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN).
-The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a representation of the jet energy distribution.
+These variables are combined and studied in a multi-layer perceptron (MLP) with two hidden layers to extract correlations among the variables.
+Next, we employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN).
+The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a fixed-representation of the jet energy distribution.
 Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}.
-Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method using subjets with multiple angular resolutions.
-At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axes which are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
-The corresponding subjet expansion can be truncated at any finite order which allows us to order-by-order examine the information carried by subleading subjets.
+Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method.
+At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axes that are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
+The corresponding subjet expansion can be truncated at any finite order thus allowing us to order-by-order examine the information carried by subleading subjets.
 These three methods are used in discriminating quark and gluon jets, as well as discriminating jets in proton-proton and heavy ion collisions.
-We will see that new insights can be gained from comprehensively comparing different probes across different collision systems.
 
-Because of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}.
+Due of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically well understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}.
 In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$.
-Within the catchment area of jets the significant background radiation needs to be subtracted before meaningful jet observable measurements can be made.
-Note that multiple jet substructure observables are analyzed simultaneously therefore a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm} is preferred.
-Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables.
-We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the telescoping deconstruction framework.
-We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent jets and illustrate the regions of phase space in the context of soft drop grooming procedure.
+Within the catchment area of such jets, the significant background contamination needs to be subtracted before meaningful jet observable measurements can be made.
+In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}.
+Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlights contributions from hard, collinear particles to jet observables.
+We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the context of the telescoping deconstruction framework.
+We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of the soft drop grooming procedure.
 
@@
 In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations.
-In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space in jet formation separated in the soft-drop procedure.
-We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook of future studies.
-
-% Different simulations may produce differently characteristic quark and gluon jets, and this method can help compare simulations and be extended to using the data from experiments.
+In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space in jet formation separated by the soft-drop procedure.
+We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook to future studies.
 
@@
 To use quark and gluon jets as independent probes of heavy ion collisions, we need to first define and identify the respective probes.
-In this paper, we define jet flavor based on the parton from hard scattering which initiates the parton shower evolution.
-We use the \jwpy Monte Carlo simulation to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
-\jwpy is a perturbative QCD Monte Carlo framework for simulating jets in proton-proton and heavy ion collisions.
-We use prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
-It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
-$Z$+jet) to enhance either the gluon or the quark jet fraction.
-This can then be directly applied to analyze other jet simulations and experimental data in the future.
+In this paper, we define jet flavor based on the parton from hard scattering that initiates the parton shower evolution.
+We use the \jw Monte Carlo simulation, a perturbative jet evolution framework to generate proton-proton and central (0-20\%) lead-lead collisions at $\sqrt{s}=2.76$ TeV using the standard setup \cite{Zapp:2013zya}.
+We use the prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
+While It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
+$Z$+jet) to enhance either the gluon or the quark jet fraction, we do not follow this path in this current analysis.
+Weak and unsupervised learning algorithms can be employed in the future on such samples inorder to treat both data and simulations on the same footing.
 
-In each event, jets are reconstructed using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp} implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with $R = 0.4$.
+In each simulated event, jets are reconstructed using the anti-$k_t$ algorithm~\cite{Cacciari:2008gp} as implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} framewith with a jet distance parameter of $R = 0.4$.
 We impose the following kinematic cuts,
@@
 where $p_T$, $\eta$ and $\phi$ are the transverse momentum, pseudo rapidity and azimuthal angle, respectively, to select high $p_T$ jets recoiling against the prompt photon.
-The medium-induced energy loss in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion.
-The medium formation time $\tau_\text{i}=0.6 $ fm and initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
+Jet quenching in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion.
+The medium formation time is set to $\tau_\text{i}=0.6 $ fm along with an initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
 We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHAPDF}~\cite{Whalley:2005nh}.
@@
 In order to remove the thermal component of recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-The \jw simulation produces HepMC~\cite{} files which are processed using a \textsc{Rivet} analysis framework~\cite{} to reconstruct the jets.
-
-%For a detailed description of the MC implementation of the jet production and its elastic/inelastic interactions with the scattering centers along with the LPM effect in \jw, please refer~\cite{}.
-The primary event collection was split into two halves, with one utilized for training the model(s) and the other for testing \textcolor{red}{additional to a di-jet sample, with a mixture of quark and gluon jets utilized for further evaluation of the model performance}.
-The question of what constitutes a "quark" or "gluon" jet is a topic of current discussion~\cite{} but for this paper, we choose the conventional definition of a quark/gluon jet with its progenitor hard scattered parton.
-There are two operational modes of generating \jw events in heavy ion collisions; with and without including the recoil partons, those originating from interactions of the propagating hard scattered parton with the scattering centers.
-In recent studies, \jw with recoils was shown to accurately reproduce several of the important qualitative features in jet structure modifications measured by LHC experiments with the addition of background subtraction techniques introduced here~\cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-The jet-background in \jw consists of the scattering center's thermal momenta i.e, their momentum before the interaction.
-In studying these different jet collections, we extract the dependence of the tagging on key operations performed on jets in data, primarily discretization and background subtraction and discuss their impact on the physics.
+The \jw simulation produces HepMC~\cite{Dobbs:2001ck} files which are processed using a \textsc{Rivet} analysis framework~\cite{Buckley:2010ar}.
 
@@
 \end{itemize}
-For jets in proton-proton collisions, we use the sets of jet samples with or without discretization.
+For jets in proton-proton collisions, we use the sets of jet samples with and without discretization.
 In experiments, the fluctuating, underlying event background is subtracted via specific background subtraction techniques.
-Since \jw does not produce full underlying events in quenched samples, we can not directly apply those background subtraction methods.
-With the application of the grid subtraction (GridSub) technique, the collection of jets represent ideally subtracted jets in experiment.
+Since \jw does not generate full underlying events in quenched samples, we can not directly apply those background subtraction methods.
+With the application of the grid subtraction (GridSub) technique, the collection of jets in \jw represent ideally subtracted jets in the experiment.
 This enables us to focus on and highlight the effects of jet quenching, as opposed to additional, nontrivial smearing effects.
@@
 We use quark jets and gluon jets as different probes of heavy ion collisions and compare their modification patterns.
-In doing so one attempts to identify qualitatively different jet features which are sensitive to aspects of quark and gluon jet quenching mechanisms.
-Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which show differences between them.
-The idea is that one can compare quark and gluon jet substructures in the same collision system and use their differences as a way to quantify the jet modification.
-New insights can be gained by putting jet substructures in multiple contexts.
-This can also help disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments.
+In doing so one attempts to identify qualitatively different jet features that are sensitive to aspects of quark and gluon quenching mechanisms.
+Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which emphasize the differences.
+By comparing quark and gluon jet substructures in the same collision system, one can quantify the respective jet modification.
+This method also helps to disentangle jet quenching effects on common features between quark and gluon jets, most notably the jet transverse momentum, and cancel systematic uncertainties in experiments.
 A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Figure~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
 
-Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures \cite{}.
+% this whole paragraph seems unnecessary and repetitive since we mentioned a lot of it, some almost word for word, in the introduction. 
+%Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures.
 With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful selections \cite{Gallicchio:2011xq,Gallicchio:2012ez}.
@@
 
-Multivariate analyses have been successfully employed in object classification and selection by the LHC experiments \cite{} and thus are an effective, physics-motivated baseline in our study.
-Given a jet reconstructed using the anti-$k_{T}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
+Multivariate analyses have been successfully employed in object classification and selections in high energy physics \cite{Bhat:2010zz, 1742-6596-608-1-012058} and thus are an effective, physics-motivated baseline in our study.
+Given a jet reconstructed using the anti-$k_{t}$ algorithm, we choose the following set of five jet substructure observables \cite{Gallicchio:2012ez} to highlight the differences between quark jets and gluon jets,
 	\begin{itemize}
@@
 
-Figure ~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
+Figure:~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
 Both pp and PbPb jets include the grid discretization.
@@
 Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions.
-In vacuum, they reflect the fact that gluon jets have larger jet mass and radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$.
-In the \jwpy-simulated PbPb collisions, for both quark and gluon jets the distributions are modified towards larger values.
-On the other hand, $p_T^D$ and the pixel multiplicity are IRC unsafe.
-We see that there is significant increase of the pixel multiplicity and decrease of the values of $p_T^D$.
-This is due to the presence of the increased number of soft particles inside jets from the \jw medium recoil contributions \cite{KunnawalkamElayavalli:2017hxo}.
+In vacuum, they reflect the fact that gluon jets have larger jet mass and larger radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$.
+In the \jwpy-simulated PbPb collisions, both quark and gluon jets have distributions that are modified towards larger values.
+On the other hand, $p_T^D$ and the pixel multiplicity are IRC unsafe observables that show a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets respectively.
+%This is due to the presence of the increased number of soft particles inside jets from the \jw medium recoil contributions \cite{KunnawalkamElayavalli:2017hxo}.
 
@@
 The unfolding procedure for jet observables increases in dimensionality and complexity as the correlations among observables become strong.
-For example, the jet mass will have to be unfolded in the 4-dimensional distribution of the generated and reconstructed jet \pt ~and mass.
+For example, the jet mass will have to be unfolded with a 4-dimensional response matrix consisting of the generated and reconstructed jet \pt ~and mass.
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgQuarkJet}
-\includegraphics[width=0.45\textwidth]{plots/jewel_pp_avgGluonJet}
-\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgQuarkJet}
-\includegraphics[width=0.45\textwidth]{plots/jewel_pbpb020_avgGluonJet}
+\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
+\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
+\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
+\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
 \caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}.
@@
 
-The multivariate analysis discussed previously is a constructive approach to collect and examine the usefulness of jet features which may help distinguish quark jets from gluon jets, or jets in different collision systems.
-An alternative approach is to use image-recognition techniques which optimize comprehensive multivariate models and identify all possible, useful features.
-The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane of transverse momentum deposit in detector calorimeter cells, much like a digital camera.
+The multivariate analysis discussed previously is a constructive approach to collect and examine the usefulness of jet features that may help distinguish quark jets from gluon jets, or jets in different collision systems.
+An alternative approach is to use image-recognition techniques which optimize comprehensive multivariate models and identify all possible useful features.
+The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane with \pt deposits in the detector cells, much like a digital camera.
 The jet image thus provides a fixed-dimensional representation of jet fragmentation.
-Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the highest pixel centers at the origin.
-Also, the image is rotated so that the principal component of the energy density is along the $\eta=0, \phi<0$ direction.
+Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the highest pixel is centered at the origin.
+Also, the image is rotated so that the principal component of the energy density lies along the $\eta=0, \phi<0$ direction.
 
-Figure~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
+Figure:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
 The color of each image pixel represents the transverse momentum deposited in the pixel.
-We see the broader energy distribution around central pixels for gluon jets compared to quark jets in pp collisions, and the significantly broader distributions for jets in PbPb collisions.
-Note that in experiment it may be very challenging to directly analyze jet images due to detector and measurement considerations.
+We see a broader energy distribution around central pixels for gluon jets as compared to quark jets in pp collisions, with both flavors exhibiting significantly broader distributions for jets in PbPb collisions.
+Note that in experiment, it may be very challenging to directly analyze jet images due to detector and measurement considerations.
 Here we mainly use the jet image approach as a useful and state-of-the-art comparison of the classification performance.
@@
 The multivariate analysis and jet image recognition discussed previously are two characteristically different methods.
-The multivariate analysis uses physics-motivated observables and is a ``bottom-up" approach to collect useful jet features.
-However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve the method.
-On the other hand, jet images represent low-level, comprehensive jet information.
+The former uses physics-motivated observables and is a ``bottom-up" approach to collect useful jet features.
+However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve on the method.
+On the other hand, the latter jet images method represent low-level, comprehensive jet information.
 The training of a deep neural network is a ``top-down" approach to identify all useful features through model optimization.
@@
 We develop and apply the telescoping deconstruction (TD) framework in our study of quark and gluon jet substructures in pp and PbPb collisions.
-Telescoping deconstruction aims to embrace both advantages of multivariate analysis and jet image recognition.
+TD aims to embrace both advantages of multivariate analysis and jet image recognition.
 It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized fragmentation basis of jet substructures.
-It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables.
-The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual telescoping deconstruction observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
-The telescoping procedure probes radiation around dominant energy flows in a jet with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla,Chien:2017xrb}.
-It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction and then reaching out to include wide-angle, soft radiation.
-Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders.
-Details about the telescoping deconstruction procedures are provided in the Appendix, including the applications to boosted $W$ and top tagging in high energy physics.
+It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables i.e the subjet \pt and mass.
+The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual TD observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
+The telescoping procedure probes radiation around dominant energy flow in a jet with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla,Chien:2017xrb}.
+It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
+Figure:~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders (left to right).
+Details about the telescoping deconstruction procedures are provided in the Appendix A, including the applications to boosted $W$ and top tagging in high energy physics.
 
 For $R=0.4$ jets, we telescope around the subjets axes using radii from 0.08 to 0.4 with $N_{\rm steps} = 5$ steps of 0.08.
-The subjet observables up to the $N$-th order are combined using dense neural networks to maximize the classification performance.
 We consider $N=\{1,2,3\}$ in this study and find the performance quickly saturates at $N=3$.
@@
 %Thus, third order TD includes by definition, all variables from the second and first order TD expansion of a jet.
-These telescoping variables are cumulatively combined and input into a multi-layer perceptron, much like the one we used in the previous multivariate analysis, with two hidden dense layers and a final output layer of activation and optimization functions.
+These telescoping variables are cumulatively combined and input to a multi-layer perceptron, much like the one we used in the previous multivariate analysis, with two hidden dense layers and a final output layer with the same activation and optimization functions.
 
-Beside the overall classification performance using telescoping deconstruction which we will present in the next section, subjet variables within this framework can reveal many aspects of jet dynamics.
+Beside the overall classification performance using telescoping deconstruction we present in the next section, subjet variables within this framework can reveal many aspects of jet dynamics.
 In particular, there is rich information contained in subjet topology and subjet substructure.
@@
 Here $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two branches, and $z_{\rm cut}$ is the soft-drop parameter which is set to $0.1$ in our analysis.
-Note that the Cambridge/Aachen clustering tree is angular-ordered, therefore the above procedure identifies the most wide angle, soft subjet which carries a significant fraction of the jet transverse momentum larger than $z_{\rm cut}$.
+Note that the C/A clustering tree is angular-ordered and therefore the above procedure identifies the most wide angle, soft subjet that carries a significant fraction of the jet transverse momentum larger than $z_{\rm cut}$.
 The size of the soft subjet is dynamically determined and related to the angle $r_g$ between the two soft-drop branches, defined as the groomed jet radius.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_ppcomp_SD_TD2_new.pdf}
-	   \caption{Top: Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum  $z_{\rm T2}$ (right); Bottom: Distributions of groomed jet radius $r_g$ and the TD angular separation between subjets $\theta_{\rm T2}$ for of quark (blue) and gluon (red) jets in pp \jw simulations.}
+	   \caption{Top: Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum  $z_{\rm T2}$ (right); Bottom: Distributions of groomed jet radius $r_g$ and the second order TD angular separation between subjets $\theta_{\rm T2}$ for of quark (blue) and gluon (red) jets in pp \jw simulations.}
 \label{fig:comp_z_pp_wT2}
-\end{figure}
-
-\begin{figure}[t]
-	   \centering
-	   \caption{Distributions of groomed jet radius $r_g$ and TD subjet momentum  $z_{\rm T2}$ of quark (blue) and gluon (red) jets in pp.}
-\label{fig:comp_z}
 \end{figure}
@@
 
-We use telescoping deconstruction to analyze the two soft-drop branches defining the $z_g$ observable.
+% This next paragraph could use a bit of a rewrite. 
+We use telescoping deconstruction to analyze the two soft-drop branches that define the $z_g$ observable.
 As a first attempt, at the T2 order we choose the subjet radius $r$ proportional to the angle $\theta$ between the two deconstruction axes for subjet reconstruction: $r=C \theta$.
-The proportionality constant $C$ is set to be 0.3, a number smaller than $0.5$, in order for the two subjets be separated in angle.
+The proportionality constant $C$ is set to be 0.3, to force the two subjets separated in angle.
 We impose a similar condition for the momentum fraction $z_{\rm T2}$ of the softer subjet to be larger than $z_{\rm cut}$,
@@
 This identifies the kinematics of the branching at the root of the C/A clustering of the two/three subjets.
-The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Figure~\ref{fig:comp_z}.
-The top-left and bottom-left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy-simulated PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Fig:~\ref{fig:comp_z}.
+The top-left and bottom-left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 On the other hand, the momentum fraction $z_{\rm T3}$ (right panels) for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
@@
 In addition to the subjet momentum fraction which provides information about the longitudinal momentum distribution, the angular separation between the two subjets encodes information about the momentum distribution transverse to the jet direction.
-Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
-Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
+Figure:~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
+Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$, the jet radius.
 The observables both have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Fig:~\ref{fig:comp_rg}.
@@
 While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jwpy Monte Carlo simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
-
-%For the $r_g$ observable, the modification induced to a jet due to quenching is clearly visible in that the subleading subjet gets pushed to a larger distance away.
-While this particular observable has not yet been experimentally measured at the LHC or the RHIC, the phenomenon of broadening of the jet observed in these MC calculations are consistent with measurements of the jet shape~\cite{}.
-Both the soft-drop $r_g$ and the T3 $\theta$ show deviations due to quenching, but the quenched distributions do not match up with each other.
-Softdrop shows a narrow peak for quenched jets at 0.3, where as the T3 distribution is broader since the sub-jet axis definition are not the same (Softdrop uses C/A whereas TD uses WTA).
-Thus these are similar observables that in comparison, we extract the physics due to quenching but these do not provide an apples to apples comparison.
 
@@
 
+\begin{figure}[t]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
+	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (top) and soft (bottom panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \jwpy.
+}
+\label{fig:comp_subjet_m}
+\end{figure}
+
 In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are quite insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
 They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
-To go beyond to higher-order jet features, we also study the subjet substructure, for example subjet masses, for both the subjets.
+To go beyond the higher-order jet features, we also study the subjet substructure, for example subjet masses.
 These observables disentangle the subjet kinematic information and are more directly sensitive to the soft radiation around subjets.
-Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (left) and gluon (right) jets.
+Figure:~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (blue dashed) and gluon (red dashed) jets.
 Here hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
@@
 On the other hand, both the subjets of a gluon jets are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
-Also, due to the effect of quenching (bottom panels in Fig:~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
-This suggests the qualitative feature of the soft event activities in the \jwpy simulations, which persists to affect not only the subjet topology but also the subjet substructure.
-
-The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion, and it can systematically improve quark gluon discrimination and give insights on the jet quenching mechanisms implemented in Monte Carlo simulations.
-Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
-
-%The first point of difference is seen , i.e.
-the pp hard and soft subjet mass are comparable with the softer subjet having a larger tail.
-The quenched softer subjet as expected has a much larger tail towards larger masses indication its constituent distribution is spread out over a larger area as opposed to the quenched harder subjet.
+Also, due to the effect of quenching (right panels in Fig:~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
+This suggests that the qualitative feature of the soft event activities in the \jwpy simulations, persists to affect not only the subjet topology but also the subjet substructure.
 
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
-	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$, for quark (left) and gluon (right) jets in pp (solid and dashed lines) and centra PbPb (dotted and dot-dashed lines) collisions in \jwpy.
-}
-\label{fig:comp_subjet_m}
+	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
+	   \caption{Distributions of mass difference between softdrop groomed mass and invariant jet mass for quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \jwpy.}
+\label{fig:comp_delta_m2}
 \end{figure}
+
+% ADD DISCUSSION HERE ABOUT THE MASS DIFFERENCE
+
+The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion.
+It systematically improves quark gluon discrimination, leading to insights about jet quenching mechanisms as implemented in Monte Carlo simulations.
+Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
 
@@
 
-We discussed jet representations using multiple, physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
-In this section we will use these information collectively and exhaustively in the study of jet modification and quark gluon discrimination, employing multivariate methods.
+In the previous section we discussed jet representations using multiple, physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
+In this section we will use these information collectively and exhaustively in the study of jet modification and quark gluon discrimination.
 
-For quark gluon discrimination, the performances are shown using Receiver Operating Characteristic (ROC) curves which plot the gluon jet efficiency as a function of the quark jet efficiency.
+For quark gluon discrimination, the performances are shown using Receiver Operating Characteristic (ROC) curves which plot the gluon jet efficiency or fake rate as a function of the quark jet efficiency or purity.
 Each ROC curve is labeled with the corresponding area under the curve (auc).
-Note that a smaller auc corresponds to a better performance, with larger values of quark efficiency and smaller values of gluon efficiency.
+Note that a smaller auc corresponds to a better performance, with larger values of quark purity and smaller values of gluon fake rate.
 
-\begin{figure}[t]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_pp_2p76TeV_genLevel_QvsG.pdf}
-	\includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
+	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_pp_2p76TeV_genLevel_QvsG.pdf}
 	\caption{ROC curves of quark v.s.
-gluon efficiency for particle jets in pp (left panel) and central PbPb (right panel) collisions in \textsc{Jewel}.
+gluon efficiency for particle jets in pp collisions in \textsc{Jewel}.
 The different colored curves correspond to various classification models studied in this paper, and the quantity in the brackets marked by auc represents the area under the curve.
@@
 The blue solid line corresponds to the DCNN trained on jet images, while the brown, orange and teal lines are for the TD at T1, T2 and T3 orders, respectively.}
-\label{fig:ROC_all}
+\label{fig:ROC_pp}
 \end{figure}
 
-Figure~\ref{fig:ROC_all} shows the ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp (left panel) and PbPb (right panel) collisions for particle jets, i.e.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
+	\caption{Similar to Fig:~\ref{fig:ROC_pp} ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}.}
+\label{fig:ROC_pbpb}
+\end{figure}
+
+Figures~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions respectively for particle jets, i.e.
 jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction.
-We see an overall compatibility among the methods with similar classification performances, implying that the three jet representations capture comparable jet information.
-However, there are noticeable performance differences.
+We see a remarkable overall compatibility between the methods with similar classification performances, implying that the three jet representations capture comparable jet information.
+However, there are noticeable performance differences especially in their ordering.
 Note the monotonic increase of the telescoping deconstruction performances from $\rm T1$ to $\rm T1+T2$, to $\rm T1+T2+T3$ with increasing amount of input information of higher-order subleading subjets.
 It is important to note that the TD classification performance saturates quickly at the T2 order, with additional higher-order features offering little improvement in the quark efficiency given a gluon efficiency.
-In pp collisions (left panel of Figure~\ref{fig:ROC_all}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
-This implies that in the pp environment the $\eta-\phi$ discretization still keeps much jet information, and the TD performance quickly saturates to the DCNN performance.
-In PbPb collisions (right panel of Figure~\ref{fig:ROC_all}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
-%Moreover, the $\eta-\phi$ discretization results in a poorer DCNN performance than TD with radius scan in steps of 0.08.
+In pp collisions (Fig:~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
+This implies that in the pp environment the $\eta-\phi$ discretization doesnt smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance.
+In PbPb collisions (Fig:~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
 Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets.
-With this difficulty, we also show the ROC curve (the dotted green line in the right panel of Figure~\ref{fig:ROC_all}) for the MLP with physics-motivated variables except for the pixel multiplicity, which reduces the auc by $0.02$.
-% and has a comparable performance as DCNN.
-Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
-Comparing left and right panels we see hints of a degradation of the classification performance in PbPb collisions, while the huge underlying event background may play a role.
+To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$.
+Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference amongst other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
+%Comparing Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}, we see a stark degradation in the classification performance in PbPb collisions which we shall proceed to study in detail.
 
-\begin{figure}[t]
+\begin{figure}[h]
 	\centering
-        \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
-        \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
-	\caption{(Left panel) ROC curves of quark v.s.
+        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
+	\caption{ROC curves of quark v.s.
 gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions with GridSub subtraction.
 The different colors represent different TD orders (T1: blue, T2: red, T3: green).
-(Right panel) ROC curves of quark v.s.
-gluon efficiency using TD at the T3 order for jets with pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid (solid red) and PbPb grid with GridSub subtraction (solid blue) constituents.
 The corresponding area under the curves are shown in the brackets.}
@@
 
-In order to have a fair comparison between jets in pp and PbPb collisions and directly compare with data, we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in the left panel of Figure~\ref{fig:ROC_TD_grid}.
-We clearly see that the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
-This suggests that the information carried in higher-order subleading subjets can be washed out in \jwpy-simulated heavy ion collisions due to the huge underlying-event soft activities or, more interestingly, the thermal randomization caused by the medium interactions.
-In the right panel of Figure~\ref{fig:ROC_TD_grid} we plot the TD ROC curves using jets with or without $\eta-\phi$ discretization and background subtraction.
+\begin{figure}[h]
+	\centering
+        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
+	\caption{ROC curves of quark v.s.
+gluon efficiency using TD at the T3 order for jets with pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid (solid red) and PbPb grid with GridSub subtraction (solid blue) constituents.
+The corresponding area under the curves are shown in the brackets.}
+	\label{fig:ROC_TD3_grid}
+\end{figure}
+
+In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Fig~\ref{fig:ROC_TD_grid}.
+We clearly see the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
+This suggests that the information carried in higher-order subleading subjets can be washed out in \jw heavy ion collisions due to the medium interactions.
+In Fig~\ref{fig:ROC_TD3_grid} we plot the ROC curves for TD at third order and compare jets with or without $\eta-\phi$ discretization and background subtraction.
 Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance using particle jets or grid jets.
-On the other hand, the performance drops with the background subtraction performed.
-This suggests that some intrinsic differences between quark jets and gluon jets are removed in the subtraction, bringing the two flavors of jets together in their fragmentation basis.
+On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed.
+This suggests some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.
 
@@
 	   \centering
-	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
-	   \includegraphics[width=0.49\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
+	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
+	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
 	   \caption{ROC curves of pp-jet v.s.
-PbPb-jet efficiency for quark (left panel) and gluon (right panel) jets with grid discretization.
+PbPb-jet efficiency for quark (top panel) and gluon (left panel) jets with grid discretization.
 The different colored curves correspond to various classification models studied in this paper.}
@@
 For a given jet flavor, i.e.
-quark or gluon jets, one can also study with various aforementioned methods how the medium modifies jets by comparing vacuum jets with quenched jets in the context of proton-heavy ion jet discrimination.
-Figure~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficency as a function of the pp jet efficiency, for discriminating quark jets (left panel) or gluon jets (right panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
-We see huge differences between jets in pp and \jwpy-simulated PbPb collisions.
+quark or gluon jets, one can also study with the aforementioned methods, how the medium modifies jets by comparing vacuum jets with quenched jets in the context of proton-heavy ion jet discrimination.
+Figure:~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficiency as a function of the pp jet efficiency, for discriminating quark jets (top panel) or gluon jets (bottom panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
+We see huge differences between jets in pp and \jwpy PbPb collisions.
 All the methods perform significantly better in the task of identifying if a jet is quenched or not, compared to quark gluon discrimination.
@@
 We observe that the increase of the pixel multiplicity is a key feature in identifying quenched jets, with the DCNN performs slightly better than the MLP.
-The better performance in the proton-heavy ion jet discrimination for gluon jets (right panel of Figure~\ref{fig:ROC_qq_gg}) suggests that gluon jets are modified more significantly than quark jets due to medium-induced quenching.
+The better performance in the proton-heavy ion jet discrimination for gluon jets (bottom panel of Figure~\ref{fig:ROC_qq_gg}) suggests that gluon jets are modified more significantly than quark jets due to medium-induced quenching.
 
-From our studies of the ROC curves for comparing quark jets and gluon jets, as well as comparing vacuum jets and quenched jets, we understand qualitatively in more details the effects of quenching as implemented in the \jwpy Monte Carlo.
+From our comprehensive comparisons of quark vs gluon jets, as well as comparing vacuum vs quenched jets, we can qualitatively understand the effects of quenching as implemented in the \jw Monte Carlo.
 Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order.
-This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can all interact with the scattering centers leading to broader and more quenched jets.
+This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets.
 In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \jw. See text for the definition of the axes.}
 \label{fig:Lund_full}
@@
 
-Each jet observable is defined as a function of the momenta and angles of jet constituents, which probes the entire jet formation with a single number.
-One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches all faithfully encode the jet information.
-In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
-It has also been recently used in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+Each jet observable is defined as a function of the momenta and angles between jet axis/constituents so as to probe the entire jet formation/structure with a single number.
+One can choose a set of jet observables to be a representation of jets, and we discussed previously, several approaches all faithfully encode the jet information.
+In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation as it was recently shown to be efficient in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
 As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
 The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling limit for jets in pp collisions because of the infrared structure of QCD.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling limit for jets in pp collisions due to the infrared structure of QCD.
 
-Given a jet reconstructed using the anti-$k_{T}$ algorithm with $R = 0.4$, the procedure of creating the Lund diagram is as follows:
+Given a jet reconstructed using the anti-$k_{T}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
 \begin{itemize}
@@
 	\item Fill in the Lund diagram for each branching
-	\item Continue along the hard branches until reaching only one particle
+	\item Continue along the subsequent hard branches until reaching only one particle
 \end{itemize}
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw.
 See text for the definition of the axes.}
@@
 
-In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal.
-Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions.
-We see a significantly enhanced region close to the $y$-axis in PbPb collisions which corresponds to increased wide angle, soft radiation in quenched jets.
-Also, gluon jets have a larger fraction of the soft radiation.
+In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
+Figure:~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions.
+We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
+Comparing quark to gluon jets, we see that gluon jets (right panel) end up with a larger fraction of the soft radiation.
 On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
-As we discussed in Figure~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
-Such picture emerges and is directly visible in the Lund diagram.
+As we discussed in Fig:~\ref{fig:comp_z}, \ref{fig:comp_rg} and Fig:~\ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
+The main advantage of the Lund diagram is that one can directly see the emergence of such qualitative features.
 
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
-	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw.
 See text for the definition of the axes.}
@@
 
-We can also directly study the effect of soft-drop using the Lund diagram.
-Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark and gluon jets in pp and PbPb collisions, with only the branchings before the one defining the $z_g$ and $r_g$ observables.
-That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+We can also study the effect of soft-drop using the Lund diagrams.
+Figure:~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
+That is, Fig:~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
 Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
 We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
-The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
-These soft jet constituents near the jet boundary are precisely the ones contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
-Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
+The bottom right panel of Fig:~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
+These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
+Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Fig:~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
 We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
-However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation.
+However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation, even though now the angular scale has significantly reduced.
 This can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
@@
 
-We present a systematic study of modifications to quark and gluon jet structure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
+We presented a systematic study of modifications to quark and gluon jet structure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
 Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems.
 We compare the performances of a MLP with physics-motivated jet variables, a state-of-the-art DCNN trained on jet images, and a newly developed method of telescoping deconstruction.
-We find that the quark gluon classification performance worsens for quenched jets.
+We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets.
 All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns.
-Through multiple methods and observables we observe consistently that the dominant feature in the \jwpy-simulated PbPb collisions is the increase of soft particle multiplicity throughout the jet region.
+Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region.
 This results in the lost of information in subleading subjets which is a characteristic feature of the \jw simulation.
@@
 The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization.
-In the future the comparisons between theoretical calculations and simulations \cite{}, and with experimental measurements, will become standard practice to identify robust features of jet-medium interactions.
-Our work suggests a systematically improvable framework which will serve this purpose.
+In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice to identify robust features of jet-medium interactions.
+Our work suggests a systematically improvable framework which will serve this important purpose.
 
@@
 We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.
-RKE also thanks for organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
+RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.

### 2018-02-26 root/plot.h
added plotting code and updated some plots

@@ -1,0 +1,618 @@
+// Standard library
+#include <math.h>
+#include <iostream>
+#include <fstream>
+
+// ROOT Library
+#include <TROOT.h>
+#include <TStyle.h>
+#include <TLegend.h>
+#include <TH1F.h>
+#include <TH2F.h>
+#include <TH3F.h>
+#include <TProfile.h>
+#include <TFile.h>
+#include <TCanvas.h>
+#include <TNtuple.h>
+#include <TTree.h>
+#include <TLine.h>
+#include <TF1.h>
+#include <TCut.h>
+#include <TPad.h>
+#include "TLatex.h"
+#include <TColor.h>
+#include <TGraphErrors.h>
+#include <TGraphAsymmErrors.h>
+#include <TGaxis.h>
+#include <TMultiGraph.h>
+
+#include <vector>
+#include <string>
+#include <set>
+#include <cmath>
+#include <exception>
+
+
+
+TLegend *getLegend(double x1, double y1, double x2, double y2)
+{
+  TLegend *leg = new TLegend(x1,y1,x2,y2,NULL,"BRNDC");
+  leg->SetHeader("");
+  leg->SetBorderSize(0);
+  leg->SetTextFont(42);
+  leg->SetTextSize(0.06);
+  leg->SetLineColor(1);
+  leg->SetLineStyle(1);
+  leg->SetLineWidth(1);
+  leg->SetFillColor(10);
+  leg->SetFillStyle(1001);
+  return leg;
+}
+
+
+// divide by bin width
+void divideBinWidth(TH1 *h)
+{
+  h->Sumw2();
+  for (int i=1;i<=h->GetNbinsX();i++)
+    {
+      Float_t val = h->GetBinContent(i);
+      Float_t valErr = h->GetBinError(i);
+      val/=h->GetBinWidth(i);
+      valErr/=h->GetBinWidth(i);
+      h->SetBinContent(i,val);
+      h->SetBinError(i,valErr);
+    }
+}
+
+// multiply by bin width
+void multiplyBinWidth(TH1 *h)
+{
+  h->Sumw2();
+  for (int i=1;i<=h->GetNbinsX();i++)
+    {
+      Float_t val = h->GetBinContent(i);
+      Float_t valErr = h->GetBinError(i);
+      val*=h->GetBinWidth(i);
[…]

### 2018-02-26 root/plot_SubJetVars.C

@@ -1,0 +1,1186 @@
+//! macro to plot the Lund Images from JEWEL samples 
+
+#include "plot.h"
+//#include "utilities.h"
+
+using namespace std;
+
+void plot_SubJetVars(){
+
+  //! Load the files for the lund images and create the plots
+  TH2::SetDefaultSumw2();
+  TH1::SetDefaultSumw2();
+  gStyle->SetOptStat(0);
+  bool dologz = false;
+  bool dorebin = false;
+  bool dozratioplot = false;
+  //! Load files in histograms 
+
+  std::string finLoc = ".";
+  std::string plotoutLoc = "../plots/";
+  
+  TFile * fin_AAjet_Gluon_rg_5R_gridsub = TFile::Open(Form("%s/AAjet_Gluon_rg_5R_gridsub.root", finLoc.c_str()));
+  TH1D * hPbPb_Gluon_rg_5R = (TH1D*)fin_AAjet_Gluon_rg_5R_gridsub->Get("Histogram");
+  hPbPb_Gluon_rg_5R->SetName("hPbPb_Gluon_rg_5R");
+  TFile * fin_AAjet_Quark_rg_5R_gridsub = TFile::Open(Form("%s/AAjet_Quark_rg_5R_gridsub.root", finLoc.c_str()));
+  TH1D * hPbPb_Quark_rg_5R = (TH1D*)fin_AAjet_Quark_rg_5R_gridsub->Get("Histogram");
+  hPbPb_Quark_rg_5R->SetName("hPbPb_Quark_rg_5R");
+
+  TFile * fin_AAjet_Gluon_delta_m2_5R_gridsub = TFile::Open(Form("%s/AAjet_Gluon_delta_m2_5R_gridsub.root", finLoc.c_str()));
+  TH1D * hPbPb_Gluon_delta_m2_5R = (TH1D*)fin_AAjet_Gluon_delta_m2_5R_gridsub->Get("Histogram");
+  hPbPb_Gluon_delta_m2_5R->SetName("hPbPb_Gluon_delta_m2_5R");
[…]

### 2018-02-26 qg_ML.tex
YT updated the draft.

@@
 We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
-We exploit jet and sub-jet features to highlight differences between quark and gluon jets in vacuum and with the jet-quenching model implemented in \jw.
-We begin with a physics-motivated analysis of jet structure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity examined in a multivariate analysis.
-In comparison, we employ state of the art image-recognition techniques by training a deep convolutional neutral network on jet images.
-To systematically extract high level jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
-We draw connections to the soft-drop subjet distribution and illuminate jet medium induced modifications using Lund diagrams.
-We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity on the softer part of the jet structure.
-Out work suggests a systematically improvable framework for studying quark and gluon classification in direct comparisons of jet simulations with measurements in heavy ion collisions.
+We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with jet-quenching model implemented in \textsc{Jewel}.
+We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity.
+In comparison, we employ state-of-the-art image-recognition techniques by training a deep convolutional neutral network on jet images.
+To systematically extract jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
+We draw connections to the soft-drop subjet distribution and illuminate medium-induced jet modifications using Lund diagrams.
+We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure.
+Our work suggests a systematically improvable framework for studying quark and gluon jet substructures in direct comparisons between jet simulations and measurements in heavy ion collisions.
 }
@@
 From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructures and jet fragmentation depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may contribute to a significant fraction of substructure modifications observed in experiments.
+Since jet substructures %and jet fragmentation
+depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
 It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
@@
 On the other hand, certain jet quenching models predict a universal narrowing of jet cores \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm,Casalderrey-Solana:2016jvj,Brewer:2017fqy} and emphasize contributions from medium responses \cite{Tachibana:2017syd}.
-Selecting pure samples of quark and gluon jets will provide the golden opportunity for studying how the medium responds to different hard probes.
+Selecting purer samples of quark and gluon jets will provide the opportunity for directly studying how the medium responds to different hard probes.
 
@@
 Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis using a Boosted Decision Tree (BDT), similar to how heavy flavor jets are identified.
-Driven by the need to identify and study boosted object searches at the LHC, the community has had significant advances on the use of machine learning techniques leading to significantly higher tagging efficiency at large kinematic ranges \cite{Komiske:2016rsd, Larkoski:2017jix}.
+Driven by the need to identify and study boosted objects at the LHC, the community has made significant advances on the use of machine-learning techniques leading to significantly higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
 
-Jet quenching studies of the past fifteen years has made compelling qualitative strides by comparing jets in proton-proton to heavy ion collisions.
+Jet quenching studies of the past fifteen years have made compelling, qualitative strides by comparing jets in proton-proton and heavy ion collisions.
 In this paper, we systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
-This indirect, complimentary approach studies the  differences in the modification of quark vs gluon jets thus disentangling modifications of their common features.
+This indirect, complimentary approach studies the  differences in the modification of quark v.s.
+gluon jets thus disentangling modifications of their common features.
 By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements.
 We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}.
-However, the method is easily applicable to other heavy ion jet simulations \cite{Armesto:2009fj,Cao:2017zih} and experimental data as well.
+However, the method is easily applicable to other heavy ion jet simulations \cite{Armesto:2009fj,Casalderrey-Solana:2016jvj,Cao:2017zih} and experimental data as well.
 
@@
 Next, we employ a modern 2-dimensional image recognition technique involving a deep convolutional neural network (DCNN).
-The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a fixed-representation of the jet energy distribution.
+The network is trained on discretized images of quark jets and gluon jets on pseudo rapidity-azimuthal angle $(\eta,\phi)$ plane which forms a fixed-dimensional representation of the jet energy distribution.
 Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}.
@@
 We use the prompt photon production channels \cite{KunnawalkamElayavalli:2016ttl} $q +\gamma$ and $g +\gamma$ to generate quark and gluon jet samples, respectively, as an attempt to prepare two sets of jets with distinct quark and gluon jet fractions.
-While It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
-$Z$+jet) to enhance either the gluon or the quark jet fraction, we do not follow this path in this current analysis.
-Weak and unsupervised learning algorithms can be employed in the future on such samples inorder to treat both data and simulations on the same footing.
+It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
+$Z$+jet) to enhance either the gluon or the quark jet fraction.
+On the other hand, weak and unsupervised learning algorithms \cite{Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the future on such samples in order to treat both data and simulations on equal footing.
 
-In each simulated event, jets are reconstructed using the anti-$k_t$ algorithm~\cite{Cacciari:2008gp} as implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} framewith with a jet distance parameter of $R = 0.4$.
+In each simulated event, jets are reconstructed using the anti-$k_t$ algorithm~\cite{Cacciari:2008gp} as implemented in \textsc{FastJet}~\cite{Cacciari:2011ma} with jet radius $R = 0.4$.
 We impose the following kinematic cuts,
@@
 
-Figure:~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
+Figure~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms).
 Both pp and PbPb jets include the grid discretization.
@@
 An alternative approach is to use image-recognition techniques which optimize comprehensive multivariate models and identify all possible useful features.
-The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane with \pt deposits in the detector cells, much like a digital camera.
-The jet image thus provides a fixed-dimensional representation of jet fragmentation.
+The radiation pattern of a jet can be quantified as an image in the $\eta-\phi$ plane with $p_T$ deposited in detector cells, much like a digital camera.
+The jet image thus provides a fixed-dimensional representation of jet energy distribution.
 Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the highest pixel is centered at the origin.
@@
 
-Figure:~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
+Figure~\ref{fig:qgjetimages} shows the average quark (left) and gluon (right) jet images in pp (top) and central PbPb (bottom) collisions.
 The color of each image pixel represents the transverse momentum deposited in the pixel.
@@
 \caption{Illustration of telescoping deconstruction at T1 (left panel), T2 (middle panel) and T3 (right panel) orders for a QCD jet.
-The crosses are the Winner-Take-All $k_T$ axes.
+The crosses are the Winner-Take-All $k_t$ axes.
 The straight lines are the exclusive subjet boundaries.
@@
 However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve on the method.
-On the other hand, the latter jet images method represent low-level, comprehensive jet information.
+On the other hand, the jet images method represent low-level, comprehensive jet information.
 The training of a deep neural network is a ``top-down" approach to identify all useful features through model optimization.
@@
 It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized fragmentation basis of jet substructures.
-It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables i.e the subjet \pt and mass.
+It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables, i.e.
+the subjet $p_T$ and mass.
 The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual TD observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
@@
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
-Figure:~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders (left to right).
+Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders (left to right).
 Details about the telescoping deconstruction procedures are provided in the Appendix A, including the applications to boosted $W$ and top tagging in high energy physics.
@@
 Note that the C/A clustering tree is angular-ordered and therefore the above procedure identifies the most wide angle, soft subjet that carries a significant fraction of the jet transverse momentum larger than $z_{\rm cut}$.
-The size of the soft subjet is dynamically determined and related to the angle $r_g$ between the two soft-drop branches, defined as the groomed jet radius.
+The sizes of the subjets are dynamically determined and related to the angle $r_g$ between the two soft-drop branches, defined as the groomed jet radius.
 The observables $z_g$ and $r_g$ thus encode the kinematic information of the two subjets which is in the category of subjet topology.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_ppcomp_SD_TD2_new.pdf}
-	   \caption{Top: Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum  $z_{\rm T2}$ (right); Bottom: Distributions of groomed jet radius $r_g$ and the second order TD angular separation between subjets $\theta_{\rm T2}$ for of quark (blue) and gluon (red) jets in pp \jw simulations.}
+	   \caption{Top panels: Distributions of groomed momentum sharing $z_g$ (left) and the T2 order TD subjet momentum fraction $z_{\rm T2}$ (right); Bottom panels: Distributions of groomed jet radius $r_g$ and the T2 order TD angular separation between subjets $\theta_{\rm T2}$ for quark (blue) and gluon (red) jets in pp \jw simulations.}
 \label{fig:comp_z_pp_wT2}
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_z_new.pdf}
-	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD (right) subjet momentum fraction $z_{\rm T3}$ of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
+	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum fraction $z_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
 Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work.
@@
 where $p_{T_1}$ and $p_{T_2}$ are the transverse momenta of the two exclusively constructed subjets at the T2 order.
-The resulting $z_{\rm T2}$ distribution is quite different from the $z_g$ distribution, especially for gluon jets as shown in the top panels of Fig:~\ref{fig:comp_z_pp_wT2}.
+The resulting $z_{\rm T2}$ distribution is quite different from the $z_g$ distribution, especially for gluon jets as shown in the top panels of Figure~\ref{fig:comp_z_pp_wT2}.
 Note that the axes used in the telescoping deconstruction procedure are defined with the winner-take-all scheme which favors energetic particles.
@@
 as well as the angle between the two subjets $\theta_{\rm T3} > \Delta$ where $\Delta$ is chosen to be 0.1 \cite{Sirunyan:2017bsd}.
-The angular separation at the second order on the other hand is also able to reproduce the softdrop groomed jet radius as shown in the bottom panels of Fig:~\ref{fig:comp_z_pp_wT2}.
+%The angular separation at the second order on the other hand is also able to reproduce the softdrop groomed jet radius as shown in the bottom panels of Figure~\ref{fig:comp_z_pp_wT2}.
 This identifies the kinematics of the branching at the root of the C/A clustering of the two/three subjets.
-The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Fig:~\ref{fig:comp_z}.
-The top-left and bottom-left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Figure~\ref{fig:comp_z}.
+The left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 On the other hand, the momentum fraction $z_{\rm T3}$ (right panels) for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
@@
 In addition to the subjet momentum fraction which provides information about the longitudinal momentum distribution, the angular separation between the two subjets encodes information about the momentum distribution transverse to the jet direction.
-Figure:~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
-Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$, the jet radius.
-The observables both have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Fig:~\ref{fig:comp_rg}.
-However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of the quenched jets (bottom panels), while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude.
+Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
+Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
+The observables both have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Figure~\ref{fig:comp_rg}.
+However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of the quenched jets, while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude (bottom panels).
 Note the difference in the subjet axis definitions which affects the angles among them: soft-drop uses C/A axes whereas TD uses WTA axes.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
-	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (top) and soft (bottom panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \jwpy.
+	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (top) and soft (bottom panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.
 }
@@
 They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
-To go beyond the higher-order jet features, we also study the subjet substructure, for example subjet masses.
+To go beyond to higher-order jet features, we also study the subjet substructure, for example subjet masses.
 These observables disentangle the subjet kinematic information and are more directly sensitive to the soft radiation around subjets.
-Figure:~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (blue dashed) and gluon (red dashed) jets.
+Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (blue dashed) and gluon (red dashed) jets.
 Here hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
@@
 On the other hand, both the subjets of a gluon jets are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
-Also, due to the effect of quenching (right panels in Fig:~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
+Also, due to the effect of quenching (right panels in Figure~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
 This suggests that the qualitative feature of the soft event activities in the \jwpy simulations, persists to affect not only the subjet topology but also the subjet substructure.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
-	   \caption{Distributions of mass difference between softdrop groomed mass and invariant jet mass for quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \jwpy.}
+	   \caption{Distributions of invariant mass difference between soft-drop groomed mass and ungroomed mass for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
 \label{fig:comp_delta_m2}
@@
 
-% ADD DISCUSSION HERE ABOUT THE MASS DIFFERENCE
+Using the general idea of telescoping jets \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2_{\rm jet}-m^2_{\rm g-jet}$.
+Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ gives another way of disentangling hard, collinear radiation and probe the soft contribution to jet mass.
+Figure~\ref{fig:comp_delta_m2} shows the $\delta m^2$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.
+We see that $\delta m^2$ in pp collisions exhibits differences between quark jets and gluon jets.
+On the other hand, the $\delta m^2$ distributions for quark and gluon jets become very similar in PbPb collisions.
+This confirms again the previous observation that the soft event activities in \textsc{Jewel} washes out differences between quark and gluon jets so that the classification performance drops.
 
@@
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
-RKE acknowledges partial support from the NSF under Grant No.1067907 \& 1352081 and DOE (need to ask Joern).
+RKE acknowledges supported in part by the U.S.
+Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.
 

### 2018-02-27 qg_ML.tex
YT updated the draft.

@@
 Since jet substructures %and jet fragmentation
-depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
+depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
 It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
@@
 Due of the huge underlying event background in heavy ion collisions, jets with smaller radii are more reliably reconstructed and theoretically well understood \cite{Dasgupta:2014yra,Chien:2015cka,Becher:2015hka,Kang:2016mcy}.
-In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_T$ algorithm \cite{Cacciari:2008gp} with $R=0.4$.
+In order to minimize the background contamination within the experimentally allowed angular resolution, we study jets reconstructed using the anti-$k_t$ algorithm \cite{Cacciari:2008gp} with $R=0.4$.
 Within the catchment area of such jets, the significant background contamination needs to be subtracted before meaningful jet observable measurements can be made.
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
-	   \caption{Distributions of invariant mass difference between soft-drop groomed mass and ungroomed mass for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
+	   \caption{Distributions of $\sqrt{\delta m^2}=\sqrt{m^2-m^2_{\rm soft~drop}}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
 \label{fig:comp_delta_m2}
@@
 
-Using the general idea of telescoping jets \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2_{\rm jet}-m^2_{\rm g-jet}$.
-Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ gives another way of disentangling hard, collinear radiation and probe the soft contribution to jet mass.
-Figure~\ref{fig:comp_delta_m2} shows the $\delta m^2$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.
-We see that $\delta m^2$ in pp collisions exhibits differences between quark jets and gluon jets.
-On the other hand, the $\delta m^2$ distributions for quark and gluon jets become very similar in PbPb collisions.
-This confirms again the previous observation that the soft event activities in \textsc{Jewel} washes out differences between quark and gluon jets so that the classification performance drops.
+Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2-m^2_{\rm soft~drop}$.
+Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ disentangles hard, collinear radiation and probes the soft radiation that is dropped.
+Figure~\ref{fig:comp_delta_m2} shows the $\sqrt{\delta m^2}$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}.
+We see that the observable $\sqrt{\delta m^2}$ in pp collisions exhibits differences between quark jets and gluon jets.
+On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions.
+This hints that soft event activities in \textsc{Jewel} can wash out differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section.
 
@@
 	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
-	\caption{Similar to Fig:~\ref{fig:ROC_pp} ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}.}
+	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}, similar to Figure~\ref{fig:ROC_pp}.}
 \label{fig:ROC_pbpb}
@@
 
-Figures~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions respectively for particle jets, i.e.
+Figures~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions, respectively, for particle jets, i.e.
 jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction.
@@
 It is important to note that the TD classification performance saturates quickly at the T2 order, with additional higher-order features offering little improvement in the quark efficiency given a gluon efficiency.
-In pp collisions (Fig:~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
+In pp collisions (Figure~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
 This implies that in the pp environment the $\eta-\phi$ discretization doesnt smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance.
-In PbPb collisions (Fig:~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
+In PbPb collisions (Figure~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
 Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets.
-To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$.
-Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference amongst other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
+To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Figure~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$.
+Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference among other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
 %Comparing Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}, we see a stark degradation in the classification performance in PbPb collisions which we shall proceed to study in detail.
@@
 	\caption{ROC curves of quark v.s.
-gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions with GridSub subtraction.
+gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions in \textsc{Jewel} with GridSub subtraction.
 The different colors represent different TD orders (T1: blue, T2: red, T3: green).
@@
 
-In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Fig~\ref{fig:ROC_TD_grid}.
+In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Figure~\ref{fig:ROC_TD_grid}.
 We clearly see the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
 This suggests that the information carried in higher-order subleading subjets can be washed out in \jw heavy ion collisions due to the medium interactions.
-In Fig~\ref{fig:ROC_TD3_grid} we plot the ROC curves for TD at third order and compare jets with or without $\eta-\phi$ discretization and background subtraction.
+In Figure~\ref{fig:ROC_TD3_grid} we plot the ROC curves for TD at the T3 order and compare jets with or without $\eta-\phi$ discretization and background subtraction.
 Since the radius scan in TD respects the $\eta-\phi$ discretization resolution, we don't see the change of the TD performance using particle jets or grid jets.
 On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed.
-This suggests some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.
+This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.
 
@@
 	   \caption{ROC curves of pp-jet v.s.
-PbPb-jet efficiency for quark (top panel) and gluon (left panel) jets with grid discretization.
+PbPb-jet efficiency for quark (top panel) and gluon (bottom panel) jets with grid discretization.
 The different colored curves correspond to various classification models studied in this paper.}
@@
 
-From our comprehensive comparisons of quark vs gluon jets, as well as comparing vacuum vs quenched jets, we can qualitatively understand the effects of quenching as implemented in the \jw Monte Carlo.
+From our comprehensive comparisons of quark v.s.
+gluon jets, as well as comparing vacuum v.s.
+quenched jets, we can qualitatively understand the effects of quenching as implemented in the \jw Monte Carlo.
 Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order.
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \jw. See text for the definition of the axes.}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
 \label{fig:Lund_full}
@@
 
-Each jet observable is defined as a function of the momenta and angles between jet axis/constituents so as to probe the entire jet formation/structure with a single number.
-One can choose a set of jet observables to be a representation of jets, and we discussed previously, several approaches all faithfully encode the jet information.
-In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation as it was recently shown to be efficient in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
+One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches all faithfully encode the jet information.
+In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
+It was shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
 As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
 The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling limit for jets in pp collisions due to the infrared structure of QCD.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD.
 
-Given a jet reconstructed using the anti-$k_{T}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
+Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
 \begin{itemize}
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw.
+	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
 See text for the definition of the axes.}
@@
 In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
-Figure:~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions.
+Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
 We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
-Comparing quark to gluon jets, we see that gluon jets (right panel) end up with a larger fraction of the soft radiation.
+Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
 On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
-As we discussed in Fig:~\ref{fig:comp_z}, \ref{fig:comp_rg} and Fig:~\ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
-The main advantage of the Lund diagram is that one can directly see the emergence of such qualitative features.
+As we discussed in Figures~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
+Using the Lund diagram one can directly see the emergence of such qualitative features.
 
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
-	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \jw.
+	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
 See text for the definition of the axes.}
@@
 We can also study the effect of soft-drop using the Lund diagrams.
-Figure:~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
-That is, Fig:~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
+That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
 Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
 We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
-The bottom right panel of Fig:~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
+The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
 These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
-Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Fig:~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
+Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
 We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
-However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions as shown by the enhanced wide angle radiation, even though now the angular scale has significantly reduced.
-This can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
+However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions.
+A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
+This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
+The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
 
@@
 
-We presented a systematic study of modifications to quark and gluon jet structure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
+We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
 Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems.
@@
 Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region.
-This results in the lost of information in subleading subjets which is a characteristic feature of the \jw simulation.
+This is closely related to the lost of information in subleading subjets, which is a characteristic feature of the \jw simulation.
 
 The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations.
-Therefore our work serves as an example how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
+Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
 The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization.
@@
 The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix.
-The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Konrad Tywoniuk and Sevil Salur for helpful discussions.
+The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions.
 We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.
 RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
+YTC also thank the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
-RKE acknowledges supported in part by the U.S.
+RKE acknowledges supports in part by the U.S.
 Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.
@@
 
-Jets are collimated sprays of particles observed at high energy colliders such as the Large Hadron Collider (LHC).
+Jets are collimated sprays of particles observed at high energy colliders such as the LHC.
 Hard collisions produce energetic quarks and gluons, which shower and hadronize through the strong interaction into the final state particles captured by the detectors.
@@
 
-In this appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation.
+In this Appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation.
 By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern.
@@
 In addition to being a useful jet representation, the telescoping deconstruction allows physical information to be easily extracted.
-\Fig{masses} shows the heavy subjet mass distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background.
+\Fig{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background.
 As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass.

### 2018-02-27 qg_ML.tex
text edits including appendix

@@
 \DeclareRobustCommand{\Tabs}[2]{Tables~\ref{#1} and \ref{#2}}
-\DeclareRobustCommand{\Fig}[1]{Fig.~\ref{#1}}
-\DeclareRobustCommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
-\DeclareRobustCommand{\Figss}[3]{Figs.~\ref{#1}, \ref{#2} and \ref{#3}}
+\DeclareRobustCommand{\Fig}[1]{Fig:~\ref{#1}}
+\DeclareRobustCommand{\Figs}[2]{Figs:~\ref{#1} and \ref{#2}}
+\DeclareRobustCommand{\Figss}[3]{Figs:~\ref{#1}, \ref{#2} and \ref{#3}}
 \DeclareRobustCommand{\Eq}[1]{Eq.~(\ref{#1})}
@@
 We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
-We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with jet-quenching model implemented in \textsc{Jewel}.
+We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with the jet-quenching model as implemented implemented in \textsc{Jewel}.
 We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity.
@@
 We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure.
-Our work suggests a systematically improvable framework for studying quark and gluon jet substructures in direct comparisons between jet simulations and measurements in heavy ion collisions.
+Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between analytic calculations, simulations and measurements in heavy ion collisions.
 }
@@
 From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructures %and jet fragmentation
-depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
-It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
+Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
+It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores as suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
 With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the changes in quark and gluon jet fractions caused by jet-medium interactions.
@@
 The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons is expected to result in broader spread of the radiation inside such jets.
-Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis using a Boosted Decision Tree (BDT), similar to how heavy flavor jets are identified.
-Driven by the need to identify and study boosted objects at the LHC, the community has made significant advances on the use of machine-learning techniques leading to significantly higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
+Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC.
+Driven by the need to identify and study boosted objects at the LHC, the high energy community has made significant advances on the use of machine-learning techniques leading to higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
 
@@
 By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements.
-We will demonstrate our general method using Monte Carlo simulations and take \jwpy as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}.
+We will demonstrate our general method using Monte Carlo simulations and take \jw as a concrete example \cite{Zapp:2013zya,KunnawalkamElayavalli:2016ttl}.
 However, the method is easily applicable to other heavy ion jet simulations \cite{Armesto:2009fj,Casalderrey-Solana:2016jvj,Cao:2017zih} and experimental data as well.
@@
 The rest of the paper is organized as follows.
-In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jwpy simulation.
+In Sec.~\ref{sec:sample} we describe the jet samples we generate using the \jw simulation.
 In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as the corresponding jet substructure distributions.
@@
 $Z$+jet) to enhance either the gluon or the quark jet fraction.
-On the other hand, weak and unsupervised learning algorithms \cite{Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the future on such samples in order to treat both data and simulations on equal footing.
+On the other hand, weak and unsupervised learning algorithms \cite{Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
 
@@
 Jet quenching in central PbPb collisions in \jw is calculated based on a medium model consisting of thermal scattering centers undergoing Bjorken expansion.
-The medium formation time is set to $\tau_\text{i}=0.6 $ fm along with an initial temperature $T_\text{i}=485$ MeV are set based on the hydrodynamic calculation~\cite{Shen:2012vn,Shen:2014vra}.
-We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHAPDF}~\cite{Whalley:2005nh}.
+The medium formation time is set to $\tau_\text{i}=0.6 $ fm along with an initial temperature $T_\text{i}=485$ MeV based on hydrodynamic calculations~\cite{Shen:2012vn,Shen:2014vra}.
+We use the \textsc{CTEQ6LL} \cite{Pumplin:2002vw} and the \textsc{EPS09}~\cite{Eskola:2009uj} parton distribution functions in \textsc{LHAPDF-5}~\cite{Whalley:2005nh}.
 We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons.
-It was shown that recoil contributions are important in \jwpy to accurately describe several qualitative features in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
-In order to remove the thermal component of recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+It was recently shown that the recoil contributions is important in \jw to accurately describe several qualitative features as seen in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 The \jw simulation produces HepMC~\cite{Dobbs:2001ck} files which are processed using a \textsc{Rivet} analysis framework~\cite{Buckley:2010ar}.
@@
 In doing so one attempts to identify qualitatively different jet features that are sensitive to aspects of quark and gluon quenching mechanisms.
-Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which emphasize the differences.
+Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which emphasize these differences.
 By comparing quark and gluon jet substructures in the same collision system, one can quantify the respective jet modification.
@@
 A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Figure~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
-
-% this whole paragraph seems unnecessary and repetitive since we mentioned a lot of it, some almost word for word, in the introduction.
-%Quark gluon discrimination has been studied extensively in the recent past, particularly in higher kinematic regimes at the LHC in order to highlight standard model and beyond the standard model signatures.
-With powerful tools for analyzing multiple variables, a large number of physics-motivated jet substructure observables have been examined systematically to identify the most useful selections \cite{Gallicchio:2011xq,Gallicchio:2012ez}.
-On the other hand, using modern machine-learning, image recognition techniques to process the raw input of the angular distribution of jet energy as jet images \cite{Komiske:2016rsd}, the discrimination performance has been shown to significantly improve at high transverse momentum.
-This suggests that additional jet features have been identified by the deep neural network to help the tagging task.
-We use both methods to gain insights on how jet quenching is manifested in these two jet representations.
-Furthermore, we develop the telescoping deconstruction framework which represents the jet information using subjets with varying angular resolution.
-It introduces an organization in the decomposition of jet substructure which allows order-by-order examination of jet fragmentation features.
-More details about the telescoping deconstruction algorithm and examples comparing QCD, $W$ and top jets are provided in the Appendix.
-We will show and discuss the classification performances of all the methods in the next section.
 
@@
 	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets.
-The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jwpy simulations.}
+The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jw simulations.}
 	   \label{fig:jetdistributons_pp_pbpb}
@@
         		\item $p_{T}^{D}$: $p^{D}_{T} = \sqrt{\sum_{i \in {\rm jet}} {p^2_{T_i}}}/p^{\rm jet}_{T}$.
-       		\item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. %including charged and neutral; given by the anti-k$_{t}$ algorithm
+       		\item Pixel multiplicity: the number of $(\eta,\phi)$ pixels with nonzero energy deposit. 
 	\end{itemize}
@@
 In vacuum, they reflect the fact that gluon jets have larger jet mass and larger radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$.
-In the \jwpy-simulated PbPb collisions, both quark and gluon jets have distributions that are modified towards larger values.
-On the other hand, $p_T^D$ and the pixel multiplicity are IRC unsafe observables that show a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets respectively.
-%This is due to the presence of the increased number of soft particles inside jets from the \jw medium recoil contributions \cite{KunnawalkamElayavalli:2017hxo}.
+In the \jw PbPb simulations, both quark and gluon jets have distributions that are modified towards larger values.
+On the other hand, pixel multiplicity and the $p_T^D$ are IRC unsafe observables that show a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets respectively.
 
@@
 We perform cross validation using Monte Carlo samples which are split into two random halves for training and testing.
-
-%increased low \pt particle multiplicity in the periphery regions of the jet leading to a broadening of the jet shape.
-Such an effect is captured in \jw, with the introduction of the medium induced recoils and introduces a further complexity in the question of quark vs gluon classification in PbPb.
 
@@
 \centering
-\includegraphics[width=.32\columnwidth]{figures/QCDjetT1_telescoped2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/QCDjetT2_telescoped2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/QCDjetT3_telescoped2.pdf}
-\caption{Illustration of telescoping deconstruction at T1 (left panel), T2 (middle panel) and T3 (right panel) orders for a QCD jet.
+\includegraphics[width=.49\columnwidth]{figures/QCDjetT1_telescoped2.pdf}
+\includegraphics[width=.49\columnwidth]{figures/QCDjetT2_telescoped2.pdf}
+\includegraphics[width=.49\columnwidth]{figures/QCDjetT3_telescoped2.pdf}
+\caption{Illustration of telescoping deconstruction at T1 (top left panel), T2 (top right panel) and T3 (bottom panel) orders for a random pp jet.
 The crosses are the Winner-Take-All $k_t$ axes.
@@
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
-Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a QCD jet at T1, T2 and T3 orders (left to right).
-Details about the telescoping deconstruction procedures are provided in the Appendix A, including the applications to boosted $W$ and top tagging in high energy physics.
+Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a random pp jet at T1, T2 and T3 orders (clockwise from top left).
+Details about the telescoping deconstruction procedures are provided in the Appendix, including the applications to boosted $W$ and top tagging in high energy physics.
 
@@
 
-% This next paragraph could use a bit of a rewrite.
 We use telescoping deconstruction to analyze the two soft-drop branches that define the $z_g$ observable.
@@
 The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Figure~\ref{fig:comp_z}.
-The left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jwpy PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+The left panels represent the soft-drop $z_{g}$ distributions in pp and PbPb collisions, and we see an enhancement of low-$z_g$ jets in the \jw PbPb collisions \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 On the other hand, the momentum fraction $z_{\rm T3}$ (right panels) for both quark and gluon jets has a similar enhancement at low $z_{\rm T3}$ values, but its distribution is less modified by quenching.
@@
 Figure~\ref{fig:comp_rg} shows the distributions of the groomed jet radius $r_g$ (left panels) and the angle $\theta_{\rm T3}$ (right panels) for quark (blue) and gluon (red) jets.
-Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
-The observables both have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Figure~\ref{fig:comp_rg}.
+%Note that the value of $r_g$ is effectively cutoff by the jet radius $R$, while the value of $\theta_{\rm T3}$ can be larger than $R$.
+Both the observables have similar distributions for quark and gluon jets in vacuum as shown in the top panels of Figure~\ref{fig:comp_rg}.
 However, we see a significant enhancement of the $r_g$ distribution at large angle ($r_g\approx0.3$) near the edge of the quenched jets, while the modification of $\theta_{\rm T3}$ has a similar enhancement but with smaller magnitude (bottom panels).
@@
 Also, the subjet radii in TD provide an extra handle on the area of coverage by subjets therefore soft radiation can be probed with various angular resolutions.
-While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jwpy Monte Carlo simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
+While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jw Monte Carlo simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
 
@@
 
-In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are quite insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
+In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are relatively insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
 They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
@@
 This is consistent with the physics picture that for a quark jet the hard subjet is typically a quark-initiated subjet, while the soft subjet is a gluon-initiated subjet through $q\rightarrow qg$ splitting.
-On the other hand, both the subjets of a gluon jets are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
-Also, due to the effect of quenching (right panels in Figure~\ref{fig:comp_subjet_m}), hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets towards larger values with an enhanced tail.
-This suggests that the qualitative feature of the soft event activities in the \jwpy simulations, persists to affect not only the subjet topology but also the subjet substructure.
+On the other hand, both the subjets of a gluon jet are mostly gluon-initiated subjets through the $g\rightarrow gg$ splitting.
+We also see the effect of quenching (right panels in Figure~\ref{fig:comp_subjet_m}), where the hard and soft subjet $\lambda$ distributions get modified significantly, especially for soft subjets which tend towards larger values with an enhanced tail.
+This suggests that the qualitative feature of the soft event activities in the \jw simulations, persists to affect not only the subjet topology but also the subjet substructure.
 
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
-	   \caption{Distributions of $\sqrt{\delta m^2}=\sqrt{m^2-m^2_{\rm soft~drop}}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
+	   \caption{Distributions of $\delta m=\sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
 \label{fig:comp_delta_m2}
@@
 
-Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop jets: $\delta m^2 \equiv m^2-m^2_{\rm soft~drop}$.
-Since soft-drop removes soft, wide angle radiation from jets, $\delta m^2$ disentangles hard, collinear radiation and probes the soft radiation that is dropped.
-Figure~\ref{fig:comp_delta_m2} shows the $\sqrt{\delta m^2}$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}.
-We see that the observable $\sqrt{\delta m^2}$ in pp collisions exhibits differences between quark jets and gluon jets.
-On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions.
-This hints that soft event activities in \textsc{Jewel} can wash out differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section.
+Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$.
+Since soft-drop removes soft, wide angle radiation from jets, $\delta m$ disentangles hard, collinear radiation and probes the soft radiation that is dropped.
+Figure:~\ref{fig:comp_delta_m2} shows the $\delta m$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}.
+We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets due to the asymmetric impact of softdrop on gluon jet mass and structure.
+On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently a broader distribution as compared to pp collisions.
+This hints that soft event activities in \textsc{Jewel} can end up smearing the differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section.
+Another interesting feature highlighted by the $\delta m$ distributions is the peak at zero, i.e.
+the fraction of jets who's mass is essentially unaffected by grooming.
+As shown in Fig:~\ref{fig:comp_delta_m2}, the distribution at x-axis $\approx 0$ is higher for gluon jets as opposed to quark jets and this trend persists even after quenching.
 
@@
 
-Figures~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions, respectively, for particle jets, i.e.
+Figure~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb} show ROC curves for quark gluon discrimination using physics-motivated variables, jet images and telescoping deconstruction in pp and PbPb collisions, respectively, for particle jets, i.e.
 jets reconstructed with particle momenta without $\eta-\phi$ discretization and background subtraction.
@@
 In pp collisions (Figure~\ref{fig:ROC_pp}) DCNN outperforms the multivariate analysis using physics-motivated variables and has a similar performance as $\rm T1+T2+T3$.
-This implies that in the pp environment the $\eta-\phi$ discretization doesnt smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance.
-In PbPb collisions (Figure~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jwpy-simulated PbPb events.
+This implies that in the pp environment the $\eta-\phi$ discretization doesn't smear the relevant jet information, and the TD performance quickly saturates to the DCNN performance.
+In PbPb collisions (Figure~\ref{fig:ROC_pbpb}), the combination of physics-motivated variables gives better performance than telescoping deconstruction at the T3 order and DCNN due to the large particle multiplicity in the \jw PbPb events.
 Measuring the particle multiplicity in heavy ion collisions is highly nontrivial due to the huge underlying event contribution and its correlation to jets.
-To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Figure~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables exclusing the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$.
+To take into account the difficulty of such an observable, we also show the ROC curve (the dotted green line in Figure~\ref{fig:ROC_pp} and Figure~\ref{fig:ROC_pbpb}) for the MLP with physics-motivated variables excluding the pixel multiplicity, which reduces the auc by a few percent $0.01 - 0.02$.
 Even though the TD with radius scan in steps of 0.08 is not able to accurately capture the full multiplicity information at the T3 order, it is still within a few percent difference among other methods in the auc, providing confidence in its ability to capture the entirety of the jet's flavor dependent fragmentation.
-%Comparing Fig:~\ref{fig:ROC_pp} and Fig:~\ref{fig:ROC_pbpb}, we see a stark degradation in the classification performance in PbPb collisions which we shall proceed to study in detail.
 
@@
 The different colored curves correspond to various classification models studied in this paper.}
-%As before, the MLP with physics-motivated variables with and without pixel multiplicity in the green solid and dotted lines, DCNN on jet images in solid blue, TD at T1, T1+T2 and third order are in solid maroon, orange and cyan lines.}
 \label{fig:ROC_qq_gg}
@@
 Figure:~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficiency as a function of the pp jet efficiency, for discriminating quark jets (top panel) or gluon jets (bottom panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
-We see huge differences between jets in pp and \jwpy PbPb collisions.
+We see huge differences between jets in pp and \jw PbPb collisions.
 All the methods perform significantly better in the task of identifying if a jet is quenched or not, compared to quark gluon discrimination.
@@
 Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
-One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches all faithfully encode the jet information.
+One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches that faithfully encode the jet information.
 In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
-It was shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
 As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
 The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
 
@@
 
-We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jwpy Monte Carlo simulations as an example.
+We presented a systematic study of modifications to quark and gluon jet substructure in heavy ion collisions, taking the \jw Monte Carlo simulations as an example.
 Jet modifications are studied via the novel method of comparing and contrasting the classification of quark jets from gluon jets in different collision systems.
@@
 RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
-YTC also thank the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented.
+YTC also thanks the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
@@
 
-Jets are collimated sprays of particles observed at high energy colliders such as the LHC.
-Hard collisions produce energetic quarks and gluons, which shower and hadronize through the strong interaction into the final state particles captured by the detectors.
-This parton shower picture has been extremely successful for understanding and modeling jets.
-Physics at all energy scales enters the parton shower and affects the final-state particle distribution.
-To extract specific features of the jet formation, many physically motivated jet substructure observables, defined in terms of the observed particles, have been studied extensively \cite{Abdesselam:2010pt,Altheimer:2012mn,Altheimer:2013yza,Adams:2015hiv,Larkoski:2017jix,ATLAS-CONF-2017-064,Khachatryan:1955546}.
-
-In this Appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation.
+In this Appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation via its fragmentation basis.
 By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern.
-We demonstrate the use of the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging, and we show that the framework captures relevant physics of jets.
+We demonstrated the use of the framework in the main text for quark/gluon discrimination and to study the medium modification in heavy ion collisions and in this appendix, we take the approach to a much different problem at the LHC: identifying and tagging boosted quark/gluon jets,  boosted $W$ and tops at the LHC.
 Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies.
 Explicit examples of the $W$ isolation \cite{Chien:2017xrb} and exposing the $W$ boson in a top jet are presented.
-This highlights the physically meaningful nature of each telescoping deconstruction observables and demonstrates their collective power as a representation.
+This highlights the physically meaningful nature of each TD observables and demonstrates their collective power as a representation.
 
 Recently, there has been progress in utilizing the complete information in a jet in an unbiased way using advanced machine-learning techniques.
-Powerful multivariate function approximators, such as neural networks, are capable of extracting useful features of the data relevant for a specific task.
+Powerful multivariate approximators, such as neural networks, are capable of extracting useful features of the data relevant for a specific task.
 Examples include jet images with convolutional neural networks (CNNs)~\cite{Cogan:2014oua,deOliveira:2015xxd,Komiske:2016rsd,Kasieczka:2017nvn}, clustering histories with recurrent neural networks~\cite{Louppe:2017ipp}, complete sets of high-level observables with dense neural networks (DNNs)~\cite{Datta:2017rhs,Datta:2017lxt,Aguilar-Saavedra:2017rzt} and linear basis of energy flow polynomials \cite{Komiske:2017aww}.
 See \cite{Larkoski:2017jix} for a more complete summary and discussion of recent progress.
-%While machine-learning methods can provide large increases in performance, extracting physical messages from the trained neural network parameters is a complicated task.
 
-Telescoping deconstruction aims to encapsulate relevant physics information in simple, physical observables.
-The variables are close to the perturbative expansion of QCD and the parton shower picture therefore perturbative and nonperturbative physics information can be systematically extracted.
+We aims to encapsulate relevant physics information in simple, physical observables derived with the TD framework.
+These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encapsulate perturbative and non-perturbative physics information.
 Similar to fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets.
 At each order, $N$ axes are determined by finding the dominant energy flow directions in the rapidity-azimuth plane~\cite{Stewart:2010tn,Chien:2013kca,Stewart:2015waa,Thaler:2015xaa}, which is partitioned into energy flow regions determined by the nearest axis.
-Jets at multiple angular resolutions are probed simply by the {\sl kinematics} of subjets consisting of particles within different distances $R_T$ from the energy flow axes.
+Jets at multiple angular resolutions are probed simply by the kinematics of subjets consisting of particles within different distances $R_T$ from the energy flow axes.
 
@@
 \includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
-\caption{\label{fig:tjet}(Top row) The telescoping deconstruction of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders.
+\caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders.
 The crosses are the Winner-Take-All $k_T$ axes.
@@
 
-The telescoping deconstruction procedure is:
+The TD procedure is prescribed as follows:
 \begin{itemize}
@@
 
-The telescoping deconstruction observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses.
-{\sl Subjet charge}~\cite{Krohn:2012fg} information can be included in this framework as well.
+The TD observables fall into two categories~\cite{Chien:2017xrb}: the {\sl subjet topology}, which is described by the axes and subjet transverse momenta, and the {\sl subjet substructure}, quantified by the subjet masses~\footnote{{\sl Subjet charge}~\cite{Krohn:2012fg} information can be included in this framework as well}.
 As the telescoping subjet order $N$ (T$N$ order) increases, more jet energy is covered by the subjets and the number of subjet radii sampled can be systematically decreased.
@@
 The telescoping deconstruction allows one to exploit features both within each subjet and among all the subjets.
-See \Fig{fig:tjet} for the telescoping deconstruction of a top jet for $N = 1$, $2$, and $3$ where the $W$ resonance can clearly be seen.
+See Fig:~\ref{fig:tjet} for the TD of a top jet for $N = 1$, $2$, and $3$ where the $W$ mass resonance can clearly be seen in the bottom panels.
 
-To demonstrate the efficacy of telescoping deconstruction in capturing the full jet information,
-we apply the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging. Each of these problems have
-signal jets with a different characteristic number of prongs: one, two, and three prongs, respectively.
+To demonstrate the efficacy of TD in capturing the full jet information, we apply the framework in quark/gluon discrimination, boosted $W$ tagging and boosted top tagging.
+Each of these problems have signal jets with a different characteristic number of prongs: one, two, and three prongs, respectively.
 Events were generated from Monte Carlo simulation of proton-proton collisions at 14 TeV using \textsc{Pythia} 8.226~\cite{Sjostrand:2007gs}.
-Final-state non-neutrino particles are clustered into jets with \textsc{FastJet} 3.3.0~\cite{Cacciari:2011ma} using the anti-$k_T$ algorithm~\cite{Cacciari:2008gp}.
+Final-state non-neutrino particles are clustered into jets with \textsc{FastJet} 3.3.0~\cite{Cacciari:2011ma} using the anti-$k_t$ algorithm~\cite{Cacciari:2008gp}.
 We consider the boosted regime with the jet $p_T$ between 800 GeV and 900 GeV.
-For quark/gluon discrimination, quark jets were generated by $pp\to q+Z(\to\nu\bar\nu)$ and gluon jets %were generated
-by $pp\to g + Z(\to\nu\bar\nu)$, clustered into $R=0.4$ jets with rapidity $|y|<1.5$.
+For quark/gluon discrimination, quark jets were generated by $pp\to q+Z(\to\nu\bar\nu)$ and gluon jets by $pp\to g + Z(\to\nu\bar\nu)$, clustered into $R=0.4$ jets with rapidity $|y|<1.5$.
 
@@
 Information from the hard process about the overall jet kinematics is eliminated by translating each jet to a frame where $y$ and $\phi$ of the jet are both zero.
-For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets from QCD dijets.
+For $W$ and top tagging, $W$ jets were generated by $pp\to WW (\to \text{hadrons})$ and top jets were generated by $pp\to t\bar t (\to \text{hadrons})$, with background jets taken from QCD dijets.
 Final-state non-neutrino particles are clustered into $R=1.0$ jets with rapidity $|y|<1.5$.
 Signal and background jets identically populate a five-bin $1/p_T^4$ histogram in transverse momentum and a three-bin uniform histogram in mass between 75 GeV and 85 GeV for $W$ tagging and 160 GeV to 180 GeV for top tagging.
-Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in
-quark/gluon discrimination, and between 0.05 to 0.95 in
-$W$ and top tagging.
+Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in quark/gluon discrimination, and between 0.05 to 0.95 in $W$ and top tagging.
 Reducing the number of radius sampling is left for future work.
@@
 One could also use boosted decision trees to combine variables \cite{Chien:2017xrb}.
-We compare the performance of our method to jet images using a CNN, which requires a graphics processing unit (GPU) to train.
+We compare the performance of our method to jet images using a CNN as done in the main text.
 The jet images are size $33\times 33$ and span a $2R\times 2R$ patch of the rapidity-azimuth plane with the intensity of each pixel corresponding to the total $p_T$ of particles in the pixel.
-The jet images are standardized according to the procedure in \Ref{Komiske:2016rsd}.
+The jet images are pre-processed and standardized according to the procedure in \Ref{Komiske:2016rsd}.
 The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer.
@@
 
-The performance of the trained models can be captured in a Receiver Operating Characteristic (ROC) curve which plots the inverse of the background mistag rate at different signal efficiencies, where a higher curve indicates better classification performance.
+The performance of the trained models can be captured in a slightly different ROC curve (compared to the main text) where we plot the inverse of the background mistag rate at different signal efficiencies, thus a higher curve indicates better classification performance.
 \Fig{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images.
 The T$N$ performance converges quickly and is comparable to the performance of the jet images approach.
-% Note that since
 The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance.
-%This is highlighted in the top case, where the telescoping deconstruction DNN outperforms the CNN at T2 order and above.
+
 \begin{figure}[h]
@@
 The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders.
-At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in %the case of
-top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
+At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in top subjets than the wider (more gluon-like) QCD subjets.
+These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
 \end{figure}
@@
 For $W$ tagging, T1 order is sufficient to achieve most of the classification performance, which unambiguously confirms the $W$ isolation feature in the boosted regime compared to the QCD background~\cite{Chien:2017xrb}.
-Clearly, the T1 order probes the depletion of the radiation at large angles within $W$ jets, whereas the QCD background jets continue to accrue mass from radiation at large angles.
+Clearly, the T1 order probes the depletion of the radiation at large angles within $W$ jets, whereas the QCD background jets continue to acquire mass from radiation at large angles.
 For quark/gluon discrimination, there is a significant increase in performance from T1 to T2 and a smaller increase from T2 to T3 where the performance saturates, suggesting the usefulness of T2 subjet substructure and its sensitivity to subjet flavors.
-This confirms that the T$N$ expansion converges efficiently and telescoping deconstruction faithfully represents the jet information.
+This confirms that the T$N$ expansion converges efficiently and TD faithfully represents the jet information.
 
-In addition to being a useful jet representation, the telescoping deconstruction allows physical information to be easily extracted.
+In addition to being a useful jet representation, the TD allows physical information to be easily extracted.
 \Fig{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background.
 As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass.
-In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they %continue to
-accrue mass at larger radii.
+In contrast, the background QCD jets do not peak at the $W$ mass for any $R_T$ and transition from QCD-like to more top-like as they acquire mass at larger radii.
 The top jets at T2 order clearly and automatically show the $W$ peak which is completely absent in the background distributions.
-At T3 order, both top and background distributions appear QCD-like, with the wider background distributions %slightly wider
-due to the prevalence of gluon subjets. This highlights again subjet substructure in telescoping deconstruction.
+At T3 order, both top and background distributions appear QCD-like, with the wider background distributions due to the prevalence of gluon subjets.
+Thus we find that TD extracts the relevant and significant subjet substructure observables in order to efficiently perform at its task, i.e.
+quark jet vs gluon jet or boosted W/tops vs QCD jets.
 

### 2018-02-27 qg_ML.tex
YT updated the draft.

@@
 We study the phenomenon of jet quenching utilizing quark and gluon jet substructures as independent probes of heavy ion collisions.
-We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with the jet-quenching model as implemented implemented in \textsc{Jewel}.
+We exploit jet and subjet features to highlight differences between quark and gluon jets in vacuum and in a medium with the jet-quenching model implemented in \textsc{Jewel}.
 We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity.
@@
 We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure.
-Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between analytic calculations, simulations and measurements in heavy ion collisions.
+Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between theoretical calculations, simulations and measurements in heavy ion collisions.
 }
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
-	   \caption{Distributions of $\delta m=\sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
+	   \caption{Distributions of $\delta m=\sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
 \label{fig:comp_delta_m2}
@@
 
-Using the general idea of telescoping jet substructure \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{Jet}^2-m_{groomed-Jet}^2}$.
+Using the general idea of telescoping jet substructure and variability \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$.
 Since soft-drop removes soft, wide angle radiation from jets, $\delta m$ disentangles hard, collinear radiation and probes the soft radiation that is dropped.
-Figure:~\ref{fig:comp_delta_m2} shows the $\delta m$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}.
-We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets due to the asymmetric impact of softdrop on gluon jet mass and structure.
-On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently a broader distribution as compared to pp collisions.
+Figure~\ref{fig:comp_delta_m2} shows the $\delta m$ distributions for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions simulated in \textsc{Jewel}.
+We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets.
+%due to the asymmetric impact of softdrop on gluon jet mass and structure.
+On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently become broader as compared to pp collisions.
 This hints that soft event activities in \textsc{Jewel} can end up smearing the differences between quark and gluon jets, which leads to the lower classification performance as we will show in the next section.
 Another interesting feature highlighted by the $\delta m$ distributions is the peak at zero, i.e.
-the fraction of jets who's mass is essentially unaffected by grooming.
-As shown in Fig:~\ref{fig:comp_delta_m2}, the distribution at x-axis $\approx 0$ is higher for gluon jets as opposed to quark jets and this trend persists even after quenching.
+there is significant fraction of jets of which the masses are essentially unaffected by soft-drop.
+In other words, for such jets no particles are removed by soft-drop.
+As shown in Figure~\ref{fig:comp_delta_m2}, the distribution at $\delta m \approx 0$ is higher for gluon jets as opposed to quark jets.
+This trend persists even after quenching.
 
@@
 RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
-YTC also thanks the organizers of the BOOST2015 conference where the idea of telescoping deconstruction was first presented.
+YTC also thanks the organizers of the BOOST2015 conference where the ideas of telescoping deconstruction and $\delta m$ were first presented.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
@@
 
-In this Appendix, we present the framework of telescoping deconstruction to systematically probe aspects of jet formation via its fragmentation basis.
+In this Appendix, we present the framework of telescoping deconstruction (TD) to systematically probe aspects of jet formation via its fragmentation basis.
 By scanning around dominant energy flows with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla}, one can efficiently quantify the radiation pattern.
-We demonstrated the use of the framework in the main text for quark/gluon discrimination and to study the medium modification in heavy ion collisions and in this appendix, we take the approach to a much different problem at the LHC: identifying and tagging boosted quark/gluon jets,  boosted $W$ and tops at the LHC.
+We demonstrated the use of the framework in the main text for quark/gluon discrimination and the study of medium modifications in heavy ion collisions.
+Here, to show the generality of TD we apply the method to different problems at the LHC: identifying and tagging boosted quark/gluon jets,  boosted $W$ and tops at the LHC.
 Crucially, the framework involves a fixed-order organization of individual observables which allows systematically improvable jet studies.
@@
 We aims to encapsulate relevant physics information in simple, physical observables derived with the TD framework.
-These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encapsulate perturbative and non-perturbative physics information.
+These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encode perturbative and non-perturbative physics information.
 Similar to fixed-order perturbative expansions and parton shower splitting kernels \cite{Nagy:2017ggp}, the expansion is ordered by the number $N$ of exclusively reconstructed subjets.
@@
 \caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders.
-The crosses are the Winner-Take-All $k_T$ axes.
+The crosses are the Winner-Take-All $k_t$ axes.
 The straight lines are the exclusive subjet boundaries.
@@
 The telescoping deconstruction allows one to exploit features both within each subjet and among all the subjets.
-See Fig:~\ref{fig:tjet} for the TD of a top jet for $N = 1$, $2$, and $3$ where the $W$ mass resonance can clearly be seen in the bottom panels.
+See Figure~\ref{fig:tjet} for the TD of a top jet for $N = 1$, $2$, and $3$ where the $W$ mass resonance can clearly be seen in the bottom panels.
 
@@
 Telescoping subjets are constructed with $R_T$ in steps of 0.05 between 0.05 and 0.4 in quark/gluon discrimination, and between 0.05 to 0.95 in $W$ and top tagging.
-Reducing the number of radius sampling is left for future work.
+%Reducing the number of radius sampling is left for future work.
 For each problem, 200k events are generated for signal and background, with 10\% used for validation, 15\% for testing, and the remaining 75\% for training.
@@
 The training can be performed on a typical laptop CPU in fewer than five minutes.
-One could also use boosted decision trees to combine variables \cite{Chien:2017xrb}.
+One could also use boosted decision trees (BDT) to combine variables \cite{Chien:2017xrb}.
 We compare the performance of our method to jet images using a CNN as done in the main text.
@@
 
-The performance of the trained models can be captured in a slightly different ROC curve (compared to the main text) where we plot the inverse of the background mistag rate at different signal efficiencies, thus a higher curve indicates better classification performance.
-\Fig{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images.
-The T$N$ performance converges quickly and is comparable to the performance of the jet images approach.
-The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance.
-
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 
+The performance of the trained models can be captured in a slightly different ROC curve (compared to the main text) where we plot the inverse of the background mistag rate at different signal efficiencies, thus a higher curve indicates better classification performance.
+Figure~\ref{ROCs} shows the ROC curves for the three tagging problems of the models trained on the cumulant telescoping deconstruction observables up to T$N$ order for $N\in\{1,2,3,4\}$ and the CNNs trained on jet images.
+The T$N$ performance converges quickly and is comparable to the performance of the jet images approach.
+The CNN architecture has not been tuned exhaustively, therefore its ROC curves serve to give a general sense of performance.
+
 For top tagging, there is a significant increase in performance from T1 to T2 order in \Fig{ROCs}, and the performance saturates beyond this order due to the sensitivity of T2 to the $W$ in top jets.
@@
 In addition to being a useful jet representation, the TD allows physical information to be easily extracted.
-\Fig{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background.
+Figure~\ref{masses} shows the heavy subjet mass \cite{Chien:2010kc} distributions at T1, T2, and T3 orders, scanned over different telescoping radii $R_T$, for top jets and their QCD background.
 As $R_T$ is increased, the top jet T1 subjet mass distributions transition from QCD-like, to peaked at the $W$ mass, to peaked at the top mass.
@@
 At T3 order, both top and background distributions appear QCD-like, with the wider background distributions due to the prevalence of gluon subjets.
-Thus we find that TD extracts the relevant and significant subjet substructure observables in order to efficiently perform at its task, i.e.
-quark jet vs gluon jet or boosted W/tops vs QCD jets.
+Thus we find that TD extracts relevant and significant subjet substructure information in order to efficiently perform at its task, i.e.
+quark jet v.s.
+gluon jet or boosted $W$/top jets v.s.
+QCD jets.
 

### 2018-02-28 qg_ML.tex
update from raghav, edited conclusions and fig5 update

@@
 We find that the quark gluon classification performance worsens for quenched jets, with a significant fraction of quark jets now masquerading as gluon jets.
-All the methods studied perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns.
-Through multiple methods and observables, we observe consistently that the dominant feature in the \jw PbPb collisions is the increase of soft particle multiplicity due to medium recoils throughout the jet region.
-This is closely related to the lost of information in subleading subjets, which is a characteristic feature of the \jw simulation.
+All the methods studied in this paper perform consistently, suggesting that the TD can extract fundamental jet fragmentation patterns.
+Through multiple methods and observables, we consistently find the dominant feature of the \jw jet quenching model to be the increase of soft particle multiplicity due to medium recoils throughout the jet region.
+This is closely related to the loss of information in subleading subjets, which is a characteristic feature of \jw.
 
@@
 Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
-The telescoping deconstruction method suggests a framework of jet substructure observables for jet physics studies, which we show to give qualitatively new insights order-by-order in its observable organization.
-In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice to identify robust features of jet-medium interactions.
-Our work suggests a systematically improvable framework which will serve this important purpose.
+By studying the telescoping deconstruction framework, we suggest new jet substructure observables for jet quenching studies, that we hope will provide qualitatively new insights order-by-order when measured experimentally.
+In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice in order to identify robust features of jet-medium interactions.
+Our work represents a thorough systematic framework that we hope will serve this important purpose.
 

### 2018-02-28 qg_ML.tex
YT updated the draft.

@@
 We find that the quark gluon discrimination performance worsens in heavy ion jets due to significant soft event activity affecting the soft jet substructure.
-Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and to facilitate direct comparisons between theoretical calculations, simulations and measurements in heavy ion collisions.
+Our work suggests a systematically improvable framework for studying modifications to quark and gluon jet substructures and facilitating direct comparisons between theoretical calculations, simulations and measurements in heavy ion collisions.
 }
@@
 The expansion is ordered by $N$, the number of subjets exclusively reconstructed, and the individual TD observables are physically motivated to encode the hard splitting of jets and relevant non-perturbative physics within each subjet variable.
-The telescoping procedure probes radiation around dominant energy flow in a jet with multiple angular resolutions \cite{Chien:2013kca,Chien:2014hla,Chien:2017xrb}.
+The telescoping procedure probes radiation around dominant energy flow in a jet with multiple angular resolutions \cite{Chien:2013kca,Adams:2015hiv,Chien:2014hla,Chien:2017xrb}.
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
@@
 quark or gluon jets, one can also study with the aforementioned methods, how the medium modifies jets by comparing vacuum jets with quenched jets in the context of proton-heavy ion jet discrimination.
-Figure:~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficiency as a function of the pp jet efficiency, for discriminating quark jets (top panel) or gluon jets (bottom panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
+Figure~\ref{fig:ROC_qq_gg} shows the ROC curves plotting the PbPb jet efficiency as a function of the pp jet efficiency, for discriminating quark jets (top panel) or gluon jets (bottom panel) in pp and PbPb collisions using physics-motivated variables, jet images and telescoping deconstruction.
 We see huge differences between jets in pp and \jw PbPb collisions.
@@
 The future of heavy ion jet physics program necessarily moves toward precise, simultaneous understanding of multiple jet substructure observables and their correlations.
-Therefore our work serves as an example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
+Therefore our work serves as a first example of how multivariate techniques can help illuminate the underlying mechanism of jet modifications.
 By studying the telescoping deconstruction framework, we suggest new jet substructure observables for jet quenching studies, that we hope will provide qualitatively new insights order-by-order when measured experimentally.
@@
             \begin{equation}
-                {\rm subjet}_{i,m} = \{p^\mu_j~|~d_{ij}<R_{T,m}~{\rm and}~d_{ij}<d_{kj} \forall i\neq k\}.
+                {\rm subjet}_{i,m} = \{p^\mu_j~|~d_{ij}<R_{T,m}~{\rm and}~d_{ij}<d_{kj}, \forall i\neq k\}.
             \end{equation}
@@
 
-For top tagging, there is a significant increase in performance from T1 to T2 order in \Fig{ROCs}, and the performance saturates beyond this order due to the sensitivity of T2 to the $W$ in top jets.
+For top tagging, there is a significant increase in performance from T1 to T2 order in Figure~\ref{ROCs}, and the performance saturates beyond this order due to the sensitivity of T2 to the $W$ in top jets.
 For $W$ tagging, T1 order is sufficient to achieve most of the classification performance, which unambiguously confirms the $W$ isolation feature in the boosted regime compared to the QCD background~\cite{Chien:2017xrb}.

### 2018-03-02 qg_ML.tex
YT updated the draft.

@@
-\documentclass[notoc]{JHEP3}
+\documentclass[notoc,preprintnumbers]{JHEP3}
 \usepackage{cite,color,psfrag,url,latexsym,graphicx,epstopdf,slashed,xspace,hyperref,enumitem,lineno}
@@
 
+\preprint{MIT-CTP 4947\\ WSU-HEP 1802}
 \title{Probing heavy ion collisions using quark and gluon jet substructure}
@@
 The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons is expected to result in broader spread of the radiation inside such jets.
-Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC.
+Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC \cite{}.
 Driven by the need to identify and study boosted objects at the LHC, the high energy community has made significant advances on the use of machine-learning techniques leading to higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
@@
 In doing so one attempts to identify qualitatively different jet features that are sensitive to aspects of quark and gluon quenching mechanisms.
-Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which emphasize these differences.
+Such direct approaches are closely related to the discrimination of quark and gluon jets, where the goal is to exploit jet features which emphasize differences between quark jets and gluon jets.
 By comparing quark and gluon jet substructures in the same collision system, one can quantify the respective jet modification.
@@
 In the \jw PbPb simulations, both quark and gluon jets have distributions that are modified towards larger values.
-On the other hand, pixel multiplicity and the $p_T^D$ are IRC unsafe observables that show a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets respectively.
+On the other hand, pixel multiplicity and the $p_T^D$ are IRC unsafe observables.
+We see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively.
 
@@
 
-The multivariate analysis and jet image recognition discussed previously are two characteristically different methods.
+\footnote{We thank Patrick Komiske and Eric Metodiev for making the beautiful, ``{\sl Ripples in Jets}" illustration in Figure~\ref{fig:tjetQCD} and the one in the Appendix, and we congratulate again the Nobel prize award to gravitational wave detection.} The multivariate analysis and jet image recognition discussed previously are two characteristically different methods.
 The former uses physics-motivated observables and is a ``bottom-up" approach to collect useful jet features.
-However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve on the method.
-On the other hand, the jet images method represent low-level, comprehensive jet information.
+However, the set of observables identified may be highly correlated, and it is not clear how to systematically improve the method.
+On the other hand, jet images represent low-level, comprehensive jet information.
 The training of a deep neural network is a ``top-down" approach to identify all useful features through model optimization.
@@
 as well as the angle between the two subjets $\theta_{\rm T3} > \Delta$ where $\Delta$ is chosen to be 0.1 \cite{Sirunyan:2017bsd}. 
-%The angular separation at the second order on the other hand is also able to reproduce the softdrop groomed jet radius as shown in the bottom panels of Figure~\ref{fig:comp_z_pp_wT2}.
-This identifies the kinematics of the branching at the root of the C/A clustering of the two/three subjets.
+This identifies the kinematics of the branching at the root of the C/A clustering of the three telescoping subjets.
 The distributions of the momentum fraction $z_{\rm T3}$ for quark (blue) and gluon (red) jets are very similar as shown in Figure~\ref{fig:comp_z}.
@@
 	   \caption{Distributions of groomed jet radius $r_g$ (left) and TD angular separation between subjets $\theta_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
-The quenched $r_{g}$ distribution is significantly modified with quenched jets having a larger sub-structure at the edges of the jet axis.}
+The quenched $r_{g}$ distribution is significantly modified with enhancement at the edge of jets.}
 \label{fig:comp_rg}
@@
 
-In order to have a fair comparison between jets in pp and PbPb collisions (and to later on directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Figure~\ref{fig:ROC_TD_grid}.
+In order to have a fair comparison between jets in pp and PbPb collisions (and to directly compare with data), we apply the grid discretization for jets in both pp and PbPb collisions, the later with the specific GridSub background subtraction technique, and compare with various TD orders in Figure~\ref{fig:ROC_TD_grid}.
 We clearly see the performance drops in PbPb collisions, along with a significant reduction in the performance improvement from T1 to $\rm T1+T2$ in PbPb collisions.
@@
 	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$) for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
 See text for the definition of the axes.}
@@
 In the future, comparisons between theoretical calculations, simulations and experimental measurements will become standard practice in order to identify robust features of jet-medium interactions.
-Our work represents a thorough systematic framework that we hope will serve this important purpose.
+Our work represents a thorough, systematic framework that we hope will serve this important purpose.
 

### 2018-03-06 qg_ML.tex
text and ref additions

@@
 The Casimir factors of quark jets and gluon jets are $C_F = 4/3$ and $C_A=3$, respectively, and the larger color charge of gluons is expected to result in broader spread of the radiation inside such jets.
-Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC \cite{}.
+Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC \cite{Chatrchyan:2012jua, Aad:2015ydr}.
 Driven by the need to identify and study boosted objects at the LHC, the high energy community has made significant advances on the use of machine-learning techniques leading to higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
@@
 $Z$+jet) to enhance either the gluon or the quark jet fraction.
-On the other hand, weak and unsupervised learning algorithms \cite{Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
+On the other hand, weak and unsupervised learning algorithms \cite{Dery:2017fap, Cohen:2017exh, Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
 
@@
 
-\footnote{We thank Patrick Komiske and Eric Metodiev for making the beautiful, ``{\sl Ripples in Jets}" illustration in Figure~\ref{fig:tjetQCD} and the one in the Appendix, and we congratulate again the Nobel prize award to gravitational wave detection.} The multivariate analysis and jet image recognition discussed previously are two characteristically different methods.
+The multivariate analysis and jet image recognition discussed previously are two characteristically different methods.
 The former uses physics-motivated observables and is a ``bottom-up" approach to collect useful jet features.
@@
 It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized fragmentation basis of jet substructures.
-It is a complete and systematic subjet expansion consisting simply of subjet kinematic variables, i.e.
+It is a complete and systematic subjet expansion~\footnote{Orthogonal frameworks utilizing jet information have been published recently~\cite{Datta:2017rhs, Komiske:2017aww}} consisting simply of subjet kinematic variables, i.e.
 the subjet $p_T$ and mass.
@@
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
-Figure~\ref{fig:tjetQCD} illustrates the telescoping deconstruction of a random pp jet at T1, T2 and T3 orders (clockwise from top left).
+Figure~\ref{fig:tjetQCD}\footnote{We thank Patrick Komiske and Eric Metodiev for making the beautiful, ``{\sl Ripples in Jets}" illustration in Figure~\ref{fig:tjetQCD} and the one in the Appendix, inspired by the nobel prize awarded to the detection of gravitational waves i.e.
+{\sl Ripples in Space-Time}.} illustrates the telescoping deconstruction of a random pp jet at T1, T2 and T3 orders (clockwise from top left).
 Details about the telescoping deconstruction procedures are provided in the Appendix, including the applications to boosted $W$ and top tagging in high energy physics.

### 2018-03-06 qg_ML.tex
added raghav acknowledgements

@@
 We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.
-RKE also thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
+RKE thanks the organizers of the DS@HEP 2017 workshop at FermiLab for an excellent introduction on machine learning techniques.
 YTC also thanks the organizers of the BOOST2015 conference where the ideas of telescoping deconstruction and $\delta m$ were first presented.
 YTC was supported by the LHC Theory Initiative Postdoctoral Fellowship under the National Science Foundation grant PHY-1419008.
-RKE acknowledges supports in part by the U.S.
+RKE also acknowledges support in part from the National Science Foundation under Grant No.1067907 \& 1352081 and in part by the U.S.
 Department of Energy Office of Science, Office of Nuclear Physics under Award Number DE-FG02-92ER-40713.

### 2018-03-07 qg_ML.tex
YT updated the draft

@@
 	   \centering
-	   \includegraphics[width=0.6\textwidth]{plots/qg_HI}
+	   \includegraphics[width=0.7\textwidth]{plots/qg_HI}
 	   \caption{Illustration of the interplay between jet quenching (vertical) and quark gluon discrimination (horizontal). }
@@
 It decomposes jets using physics-motivated, comprehensive sets of observables which form an organized fragmentation basis of jet substructures.
-It is a complete and systematic subjet expansion~\footnote{Orthogonal frameworks utilizing jet information have been published recently~\cite{Datta:2017rhs, Komiske:2017aww}} consisting simply of subjet kinematic variables, i.e.
+It is a complete and systematic subjet expansion~\footnote{Other useful basis of jet substructure observables have also been explored recently~\cite{Datta:2017rhs, Komiske:2017aww}.} consisting simply of subjet kinematic variables, i.e.
 the subjet $p_T$ and mass.
@@
 It efficiently quantifies the radiation distribution by first capturing the dominant energy in the subjet reconstruction followed by reaching out to include wide-angle, soft radiation.
-Figure~\ref{fig:tjetQCD}\footnote{We thank Patrick Komiske and Eric Metodiev for making the beautiful, ``{\sl Ripples in Jets}" illustration in Figure~\ref{fig:tjetQCD} and the one in the Appendix, inspired by the nobel prize awarded to the detection of gravitational waves i.e.
-{\sl Ripples in Space-Time}.} illustrates the telescoping deconstruction of a random pp jet at T1, T2 and T3 orders (clockwise from top left).
+Figure~\ref{fig:tjetQCD}\footnote{We thank Patrick Komiske and Eric Metodiev for making the beautiful, ``{\sl Ripples in Jets}" illustration in Figure~\ref{fig:tjetQCD} and the one in the Appendix, celebrating the Nobel prize awarded to the detection of gravitational waves, i.e., {\sl Ripples in Spacetime}.} illustrates the telescoping deconstruction of a random pp jet at T1, T2 and T3 orders (clockwise from top left).
 Details about the telescoping deconstruction procedures are provided in the Appendix, including the applications to boosted $W$ and top tagging in high energy physics.
@@
 
+\subsection{Lund Diagrams}
+\label{sec:lund}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
+\label{fig:Lund_full}
+\end{figure}
+
+Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
+One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches that faithfully encode the jet information.
+Here we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
+As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
+The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+
+Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
+\begin{itemize}
+	\item Recluster the jet using the C/A algorithm
+	\item Start from the root of the tree and move along the hard branches.
+At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
+	\item Fill in the Lund diagram for each branching
+	\item Continue along the subsequent hard branches until reaching only one particle
+\end{itemize}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
+	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+See text for the definition of the axes.}
+\label{fig:Lund_bkg}
+\end{figure}
+
+In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
+Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
+We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
+Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
+On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
+As we discussed in Figures~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
+Using the Lund diagram one can directly see the emergence of such qualitative features.
+
+\begin{figure}[t]
+	   \centering
+	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
+	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+See text for the definition of the axes.}
+\label{fig:Lund_hard}
+\end{figure}
+
+We can also study the effect of soft-drop using the Lund diagrams.
+Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
+That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
+We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
+The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
+These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
+Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
+We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
+However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions.
+A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
+This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
+The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
+
 \section{Quark and Gluon Jet Modification and Discrimination}
@@
 
-In the previous section we discussed jet representations using multiple, physics-motivated variables, jet images and telescoping deconstruction, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
+In the previous section we discussed jet representations using multiple, physics-motivated variables, jet images, telescoping deconstruction and Lund diagrams, and we showed how individual observables can be used to highlight the differences between quark jets and gluon jets, as well as between jets in proton-proton and heavy ion collisions.
 In this section we will use these information collectively and exhaustively in the study of jet modification and quark gluon discrimination.
@@
 This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets.
-In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions
-
-\section{Quark and Gluon Jets in Lund Diagrams}
-\label{sec:lund}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
-\label{fig:Lund_full}
-\end{figure}
-
-Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
-One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches that faithfully encode the jet information.
-In this section we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
-As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
-The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
-
-Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
-\begin{itemize}
-	\item Recluster the jet using the C/A algorithm
-	\item Start from the root of the tree and move along the hard branches.
-At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
-	\item Fill in the Lund diagram for each branching
-	\item Continue along the subsequent hard branches until reaching only one particle
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
-	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
-See text for the definition of the axes.}
-\label{fig:Lund_bkg}
-\end{figure}
-
-In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
-Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
-We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
-Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
-On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
-As we discussed in Figures~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
-Using the Lund diagram one can directly see the emergence of such qualitative features.
-
-\begin{figure}[t]
-	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
-	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
-See text for the definition of the axes.}
-\label{fig:Lund_hard}
-\end{figure}
-
-We can also study the effect of soft-drop using the Lund diagrams.
-Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
-That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
-Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
-We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
-The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
-These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
-Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
-We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
-However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions.
-A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
-This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
-The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
+%In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions
 

### 2018-03-07 qg_ML.tex
YT updated the draft

@@
 	   \centering
-	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
-	   \includegraphics[width=0.59\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
+	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
+	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
 	   \caption{ROC curves of pp-jet v.s.

### 2018-03-07 qg_ML.tex
updated the text - moved around the lund diagrams

@@
 Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
-It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can constitute the narrowing of jet cores as suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
+It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can contribute to the narrowing of jet cores as suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
 With jet samples consisting of different quark and gluon jet fractions, for example comparing inclusive jet and photon-tagged jet measurements, one can disentangle the effect on jet substructure modifications due to the changes in quark and gluon jet fractions caused by jet-medium interactions.
@@
 
-Jet quenching studies of the past fifteen years have made compelling, qualitative strides by comparing jets in proton-proton and heavy ion collisions.
+Jet quenching studies of the past sixteen years have made compelling, qualitative strides by comparing jets in proton-proton and heavy ion collisions.
 In this paper, we systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
 This indirect, complimentary approach studies the  differences in the modification of quark v.s.
-gluon jets thus disentangling modifications of their common features.
+gluon jets, thus disentangling modifications of their common features in a multivariate analysis.
 By first comparing probes in the same collision system, it can also help reduce the systematic uncertainty in experimental measurements.
@@
 Each jet image undergoes standard pre-processing as described in \cite{deOliveira:2015xxd}.
-Moreover, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method.
+Finally, in order to systematically extract the complete jet substructure information, we apply the newly developed telescoping deconstruction (TD) method.
 At each order $N$, TD deconstructs a jet into its fragmentation basis with regard to $N$ axes that are defined using the winner-take-all (WTA) recombination scheme \cite{Bertolini:2013iqa}.
@@
 In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}.
-Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlights contributions from hard, collinear particles to jet observables.
+Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables.
 We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the context of the telescoping deconstruction framework.
-We will also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of the soft drop grooming procedure.
+During the course of the paper, we use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of jet quenching in \jw and with regards to the soft drop grooming procedure.
 
@@
 In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations.
-In Sec.~\ref{sec:lund} we use Lund diagrams to illustrate the regions of phase space in jet formation separated by the soft-drop procedure.
 We conclude in Sec.~\ref{sec:conc} and raise awareness of such methods by giving an outlook to future studies.
@@
 It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
-$Z$+jet) to enhance either the gluon or the quark jet fraction.
-On the other hand, weak and unsupervised learning algorithms \cite{Dery:2017fap, Cohen:2017exh, Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
+$Z$+jet~\footnote{It is important to note that dijets in heavy ion events are both expected to be quenched by the medium, which can in turn lead to artificially altered quark and gluon fractions.}) to enhance either the gluon or the quark jet fraction.
+On the other hand, unsupervised or weakly supervised learning algorithms \cite{Dery:2017fap, Cohen:2017exh, Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
 
@@
 We include recoil partons in our PbPb simulations that originate from elastic and inelastic scattering with thermal scattering centers due to interactions with propagating hard scattered partons.
-It was recently shown that the recoil contributions is important in \jw to accurately describe several qualitative features as seen in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
+It was recently shown that the recoil contributions are important in \jw to accurately describe several qualitative features as seen in jet substructure modifications measured at the LHC \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
 In order to remove the thermal component of the recoil partons, we adopt the background subtraction procedure introduced in \cite{KunnawalkamElayavalli:2017hxo,Milhano:2017nzm}.
@@
 
-To emulate finite detector effects on the jet angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of hadronic calorimeter at the LHC.
+To emulate finite detector effects on the jet angular resolution, we discretize jets using $\eta-\phi$ grids of size $0.08 \times 0.08$ respecting the current angular resolution of the CMS experiment's inherent hadronic calorimeter resolution at the LHC.
 To study how such discretization as well as background subtraction affect the analysis, we use the following three different sets of jets in heavy ion collisions, all with the medium recoil contributions,
@@
 A schematic illustration of possible ways of exploiting quark and gluon jet substructure is shown in Figure~\ref{fig:quenching_discrimination} where the interplay between quenching (vertical arrows) and the jet flavor difference (horizontal arrows) are highlighted.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
+\label{fig:Lund_full}
+\end{figure}
+
+Every jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
+One can choose a set of jet observables to be a representation of jets and we discus several approaches in the coming sections that faithfully encode the jet information.
+We use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as diagrammatic jet representation through which one can comprehend the quenching mechanism in \textsc{Jewel}.
+The Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
+The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+
+Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
+\begin{itemize}
+	\item Recluster the jet using the C/A algorithm
+	\item Start from the root of the tree and move along the hard branches.
+At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\Delta \phi^2_{1, 2}}$
+	\item Fill in the Lund diagram for each branching
+	\item Continue along the subsequent hard branches until reaching only one particle
+\end{itemize}
+
+In Lund diagrams.
+wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
+Figure:~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
+We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
+Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
+On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
+This modification manifests in modified jet cores in PbPb collisions where subleading subjets tend to have wider angles and softer momenta.
+These are the high level features we hope to extract with jet representations in order to effectively tag quark jets vs gluon jets in both vacuum and in the medium.
 
@@
 In the \jw PbPb simulations, both quark and gluon jets have distributions that are modified towards larger values.
-On the other hand, pixel multiplicity and the $p_T^D$ are IRC unsafe observables.
-We see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively.
+On the other hand, we see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively, with the caveat that pixel multiplicity is unreliable in a heavy ion environment and the $p_T^D$ is an IRC unsafe observable.
 
@@
 The jet image thus provides a fixed-dimensional representation of jet energy distribution.
-Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the highest pixel is centered at the origin.
+Each jet image is pre-processed \cite{deOliveira:2015xxd} by translating in the $\eta-\phi$ plane so that the most active pixel is centered at the origin.
 Also, the image is rotated so that the principal component of the energy density lies along the $\eta=0, \phi<0$ direction.
@@
 The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer.
-The convolution layers use hyperbolic tangent activation function, and the output layer uses sigmoid activation function.
+The convolution layers use a hyperbolic tangent activation function, and the output layer uses a sigmoid activation function.
 The output layer is preceded by three layers, a dropout layer with a rejection score of 0.20, a dense layer with 20 nodes and an additional dropout layer with a reduced rejection score of 0.10.
@@
 Recently, the CMS~\cite{Sirunyan:2017bsd} and STAR~\cite{Kauder:2017mhg} collaborations measured the groomed momentum sharing observable $z_g$.
-One reclusters jets using the Cambridge/Aachen (C/A) algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and sequentially removes wide angle, soft radiation until the following condition is satisfied by a branching,
+Jets are reclustered using the Cambridge/Aachen (C/A) algorithm \cite{Dokshitzer:1997in,Wobisch:1998wt} and wide angle, soft radiation are sequentially removed until the following condition is satisfied by a branching,
 \begin{equation}
@@
 Also, the subjet radii in TD provide an extra handle on the area of coverage by subjets therefore soft radiation can be probed with various angular resolutions.
-While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jw Monte Carlo simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
+While these observables have not yet been experimentally measured at the RHIC or the LHC, the observed phenomenon of jet broadening in the \jw simulations is consistent with the jet shape measurement~\cite{Chatrchyan:2013kwa}.
 
@@
 
+In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are relatively insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
+They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
+
 \begin{figure}[t]
@@
 	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
-	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (top) and soft (bottom panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.
+	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (left) and soft (right panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (top) and central PbPb (bottom) collisions in \textsc{Jewel}.
 }
@@
 
-In vacuum, we see that all the $\{z_g,r_g\}$ and $\{z_{\rm T3},\theta_{\rm T3}\}$ distributions are relatively insensitive to the partonic origin of jets because the color factors cancel in the normalization of the distributions.
-They encode information about the leading-order subjet topology which is determined by the QCD splitting kinematics.
 To go beyond to higher-order jet features, we also study the subjet substructure, for example subjet masses.
@@
 
-Using the general idea of telescoping jet substructure and variability \cite{Chien:2017xrb} we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$.
+Using the general idea of telescoping jet substructure and variability~\cite{Chien:2017xrb}, we also look at the difference between the square of jet mass for ungroomed and soft-drop groomed jets: $\delta m \equiv \sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$.
 Since soft-drop removes soft, wide angle radiation from jets, $\delta m$ disentangles hard, collinear radiation and probes the soft radiation that is dropped.
@@
 We see that the observable $\delta m$ in pp collisions exhibits differences between quark jets and gluon jets.
-%due to the asymmetric impact of softdrop on gluon jet mass and structure.
 On the other hand, the distributions for quark and gluon jets become very similar in PbPb collisions, albeit shifted towards higher values, and consequently become broader as compared to pp collisions.
@@
 In other words, for such jets no particles are removed by soft-drop.
-As shown in Figure~\ref{fig:comp_delta_m2}, the distribution at $\delta m \approx 0$ is higher for gluon jets as opposed to quark jets.
-This trend persists even after quenching.
+As shown in Figure~\ref{fig:comp_delta_m2}, the distribution at $\delta m \approx 0$ is higher for gluon jets as opposed to quark jets and this particular trend persists even after quenching.
 
@@
 Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
-
-\subsection{Lund Diagrams}
-\label{sec:lund}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
-\label{fig:Lund_full}
-\end{figure}
-
-Each jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
-One can choose a set of jet observables to be a representation of jets, and we discussed previously several approaches that faithfully encode the jet information.
-Here we will use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as another jet representation.
-As we will see, the Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
-The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
-
-Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
-\begin{itemize}
-	\item Recluster the jet using the C/A algorithm
-	\item Start from the root of the tree and move along the hard branches.
-At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\phi^2_{1, 2}}$
-	\item Fill in the Lund diagram for each branching
-	\item Continue along the subsequent hard branches until reaching only one particle
-\end{itemize}
 
@@
 
-In the Lund diagram, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
-Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
-We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
-Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
-On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
-As we discussed in Figures~\ref{fig:comp_z}, \ref{fig:comp_rg} and \ref{fig:comp_subjet_m}, jet cores are also modified in PbPb collisions, and subleading subjets tend to have wider angles and softer momenta.
-Using the Lund diagram one can directly see the emergence of such qualitative features.
-
 \begin{figure}[t]
@@
 
-We can also study the effect of soft-drop using the Lund diagrams.
-Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
-That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+We also study the effect of soft-drop on our \jw samples using the Lund diagrams in Fig:~\ref{fig:Lund_bkg} for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
+That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure, which is a diagrammatic phase-space representation of the aforementioned $\delta m$ observable.
 Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
@@
 We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
-However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions.
+However, soft-drop does not remove all the soft radiation within the hard branch in PbPb collisions.
 A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
 This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
-The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
+The radiation can be due to correlated medium responses but is often suppressed or even removed with a statistical treatment of background in experiments.
 
@@
 On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed.
-This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.
+This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction~\footnote{Underlying event subtraction methods in experiments could influence and create un-intended features in any of our jet representations discussed.
+Such effects could only be studied with a full heavy ion event simulation and we leave this particular task for the near future.}.
 
@@
 This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets.
-%In the next section we will employ the use of another jet representation, the Lund Diagram, to illuminate the differences between quark and gluon jets, as well as jets in pp and PbPb collisions
 
@@
 Through multiple methods and observables, we consistently find the dominant feature of the \jw jet quenching model to be the increase of soft particle multiplicity due to medium recoils throughout the jet region.
-This is closely related to the loss of information in subleading subjets, which is a characteristic feature of \jw.
+This is closely related to the loss of information in subleading subjets, which is a characteristic feature of \textsc{Jewel}.
 
@@
 Powerful multivariate approximators, such as neural networks, are capable of extracting useful features of the data relevant for a specific task.
-Examples include jet images with convolutional neural networks (CNNs)~\cite{Cogan:2014oua,deOliveira:2015xxd,Komiske:2016rsd,Kasieczka:2017nvn}, clustering histories with recurrent neural networks~\cite{Louppe:2017ipp}, complete sets of high-level observables with dense neural networks (DNNs)~\cite{Datta:2017rhs,Datta:2017lxt,Aguilar-Saavedra:2017rzt} and linear basis of energy flow polynomials \cite{Komiske:2017aww}.
+Examples include jet images with convolutional neural networks (CNNs)~\cite{Cogan:2014oua,deOliveira:2015xxd,Komiske:2016rsd,Kasieczka:2017nvn}, clustering histories with recurrent neural networks~\cite{Louppe:2017ipp}, complete sets of high-level observables with dense neural networks~\cite{Datta:2017rhs,Datta:2017lxt,Aguilar-Saavedra:2017rzt} and linear basis of energy flow polynomials \cite{Komiske:2017aww}.
 See \cite{Larkoski:2017jix} for a more complete summary and discussion of recent progress.
 
-We aims to encapsulate relevant physics information in simple, physical observables derived with the TD framework.
+We aim to encapsulate relevant physics information in simple, physical observables derived with the TD framework.
 These variables follow the perturbative expansion of QCD and the parton shower picture and therefore encode perturbative and non-perturbative physics information.

### 2018-03-08 qg_ML.tex
YT updated the draft.

@@
 Additional distinctive features can be included to improve the tagging by combining a variety of jet substructure variables in a multivariate analysis, similar to how heavy flavor jets are identified at the LHC \cite{Chatrchyan:2012jua, Aad:2015ydr}.
-Driven by the need to identify and study boosted objects at the LHC, the high energy community has made significant advances on the use of machine-learning techniques leading to higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
+Driven by the need to identify and study boosted objects at the LHC, the high energy physics community has made significant advances on the use of machine-learning techniques leading to higher tagging efficiency in a wide kinematic range \cite{Komiske:2016rsd, Larkoski:2017jix}.
 
-Jet quenching studies of the past sixteen years have made compelling, qualitative strides by comparing jets in proton-proton and heavy ion collisions.
+Jet quenching studies of the past decades have made compelling, qualitative strides by comparing jets in proton-proton and heavy ion collisions.
 In this paper, we systematically study the tagging of quark and gluon jets in both proton-proton and lead-lead collisions making full use of their substructure information.
@@
 Within the catchment area of such jets, the significant background contamination needs to be subtracted before meaningful jet observable measurements can be made.
-In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires a constituent-level, observable-independent background subtraction method \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}.
+In order to study multiple jet substructure observables and their correlations simultaneously, it necessarily requires constituent-level, observable-independent background subtraction methods \cite{Soyez:2012hv,Cacciari:2014gra,Berta:2014eza,Bertolini:2014bba,Komiske:2017ubm}.
 Furthermore, various jet grooming techniques \cite{Ellis:2009su,Ellis:2009me,Krohn:2009th,Dasgupta:2013ihk,Larkoski:2014wba} systematically remove soft radiation and highlight contributions from hard, collinear particles to jet observables.
 We will draw connections from telescoping deconstruction to the soft-drop groomed momentum sharing $z_g$ and groomed jet radius $r_g$ variables \cite{Larkoski:2014wba} and discuss qualitatively new jet substructure observables within the context of the telescoping deconstruction framework.
-During the course of the paper, we use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in the context of jet quenching in \jw and with regards to the soft drop grooming procedure.
+During the course of the paper, we also use Lund diagrams \cite{Andersson1989,Salam:2016yht} to represent and showcase significant regions of phase space in \jw in the context of the soft drop grooming procedure.
 
@@
 In Sec.~\ref{sec:qg} we give details about the multivariate analysis, the jet image, and the telescoping deconstruction methods, as well as the corresponding jet substructure distributions.
+We also use Lund diagrams to illustrate the regions of phase space in jet formation separated by the soft-drop procedure.
 In Sec.~\ref{sec:results} we compare the performances of the methods and discuss the physical interpretations.
@@
 It is possible to construct jet samples by performing hard process and kinematical selections (such as dijet v.s.
-$Z$+jet~\footnote{It is important to note that dijets in heavy ion events are both expected to be quenched by the medium, which can in turn lead to artificially altered quark and gluon fractions.}) to enhance either the gluon or the quark jet fraction.
+$Z$+jet) to enhance either the gluon or the quark jet fraction.
 On the other hand, unsupervised or weakly supervised learning algorithms \cite{Dery:2017fap, Cohen:2017exh, Metodiev:2017vrx,Komiske:2018oaa,Metodiev:2018ftz} can also be employed in the near future on such samples in order to treat both data and simulations on equal footing.
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
-	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
-\label{fig:Lund_full}
-\end{figure}
-
 Every jet observable is defined as a function of the constituent momenta and angles from jet axes so as to probe the entire jet substructure with a single number.
-One can choose a set of jet observables to be a representation of jets and we discus several approaches in the coming sections that faithfully encode the jet information.
-We use the Lund diagram~\cite{Andersson1989} associated with the Cambridge/Aachen clustering tree structure as diagrammatic jet representation through which one can comprehend the quenching mechanism in \textsc{Jewel}.
-The Lund diagram directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
-The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
-The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
-
-Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
-\begin{itemize}
-	\item Recluster the jet using the C/A algorithm
-	\item Start from the root of the tree and move along the hard branches.
-At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\Delta \phi^2_{1, 2}}$
-	\item Fill in the Lund diagram for each branching
-	\item Continue along the subsequent hard branches until reaching only one particle
-\end{itemize}
-
-In Lund diagrams.
-wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
-Figure:~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
-We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
-Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
-On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
-This modification manifests in modified jet cores in PbPb collisions where subleading subjets tend to have wider angles and softer momenta.
-These are the high level features we hope to extract with jet representations in order to effectively tag quark jets vs gluon jets in both vacuum and in the medium.
+One can choose a set of jet observables to be a representation of jets, and we discus several approaches which faithfully encode the jet information.
 
@@
 
-The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion.
-It systematically improves quark gluon discrimination, leading to insights about jet quenching mechanisms as implemented in Monte Carlo simulations.
-Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
+	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
+\label{fig:Lund_full}
+\end{figure}
+
+The quenching mechanism in \textsc{Jewel} probed by telescoping deconstruction variables above can be further illuminated using the Lund diagram \cite{Andersson1989}.
+The Lund diagram~\cite{Andersson1989} is a diagrammatic jet representation associated with the Cambridge/Aachen clustering tree structure.
+It directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
+The Lund diagram is then a two-dimensional histogram of variables $\ln (R/\Delta R)$ and $\ln z(\Delta R/R)$ where $R$ is the jet radius used in jet reconstruction.
+The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+
+Given a jet reconstructed using the anti-$k_{t}$ algorithm with $R = 0.4$, the procedure of creating a Lund diagram is as follows:
+\begin{itemize}
+	\item Recluster the jet using the C/A algorithm
+	\item Start from the root of the tree and move along the hard branches.
+At each branching, construct the momentum fraction of the soft branch $z=\min(p_{T_1},p_{T_2})/(p_{T_1}+p_{T_2})$ and the angular distance between the two branches $\Delta R = \sqrt{\Delta \eta^2_{1, 2}+\Delta \phi^2_{1, 2}}$
+	\item Fill in the Lund diagram for each branching
+	\item Continue along the subsequent hard branches until reaching only one particle
+\end{itemize}
+
+In Lund diagrams, wide angle, soft radiation populates close to the $y$-axis, while collinear, hard radiation is along the diagonal (with the negative slope).
+Figure~\ref{fig:Lund_full} shows the Lund diagrams for quark (left) and gluon (right) jets in pp (top) and PbPb (bottom) collisions in \textsc{Jewel}.
+We see a significantly enhanced region close to the $y$-axis in PbPb collisions that corresponds to increased wide angle, soft radiation in quenched jets.
+Comparing quark to gluon jets, we see that gluon jets (right panel) have a larger fraction of soft radiation.
+On the other hand, the hard, collinear radiation region of the Lund diagram is shown to reduce in its significance for quenched jets.
+This suggests again that jet cores are modified in \textsc{Jewel} PbPb collisions where subleading subjets tend to have wider angles and softer momenta.
+These are characteristic features we hope to extract, and using the Lund diagram one can directly see the emergence of such qualitative features.
 
@@
 
+We can also study the effect of soft-drop using the Lund diagrams.
+Figure~\ref{fig:Lund_bkg} shows the Lund diagrams for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
+That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure.
+Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
+We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
+The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
+These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
+Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
+We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
+However, soft-drop still can not remove all the soft radiation within the hard branch in PbPb collisions.
+A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
+This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
+The radiation can be due to correlated medium responses but is often removed with a statistical treatment of background in experiments.
+
 \begin{figure}[t]
@@
 
-We also study the effect of soft-drop on our \jw samples using the Lund diagrams in Fig:~\ref{fig:Lund_bkg} for quark (left) and gluon (right) jets in pp and PbPb collisions, with only the branchings right up to the one defining the $z_g$ and $r_g$ observables.
-That is, Figure~\ref{fig:Lund_bkg} corresponds to radiation that is dropped in the soft-drop procedure, which is a diagrammatic phase-space representation of the aforementioned $\delta m$ observable.
-Note the strict cutoff of $z < z_{\rm cut} = 0.1$ in the Lund diagrams.
-We see that a large portion of soft, wide-angle radiation is isolated from the hard, collinear core.
-The bottom right panel of Figure~\ref{fig:Lund_bkg} with quenched gluon jets shows the maximum density of soft, wide angle radiation.
-These soft jet constituents near the jet boundary are precisely the ones that contribute to the modification of the jet shape in \jw \cite{KunnawalkamElayavalli:2017hxo}.
-Within the framework of this diagrammatic representation, one can also examine the harder branch after the soft-drop procedure and Figure~\ref{fig:Lund_hard} shows the corresponding Lund diagram.
-We see that the vacuum structure starts to emerge in the hard branch of soft-drop PbPb jets.
-However, soft-drop does not remove all the soft radiation within the hard branch in PbPb collisions.
-A similar observation was made in \cite{Hoang:2017kmk} that hard branchings can ``protect" soft radiation from being dropped.
-This is shown by the enhanced, wide angle radiation, even though now its angular scale has significantly reduced.
-The radiation can be due to correlated medium responses but is often suppressed or even removed with a statistical treatment of background in experiments.
+The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion.
+It systematically improves quark gluon discrimination, leading to insights about jet quenching mechanisms as implemented in Monte Carlo simulations.
+Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
 
@@
 On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed.
-This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction~\footnote{Underlying event subtraction methods in experiments could influence and create un-intended features in any of our jet representations discussed.
-Such effects could only be studied with a full heavy ion event simulation and we leave this particular task for the near future.}.
+This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction~\footnote{Underlying events and their subtractions in experiments can both influence and introduce jet substructure features.
+Such effects can only be studied with full heavy ion event simulations, which we leave for future studies.}.
 

### 2018-03-08 qg_ML.tex
YT updated the draft

@@
 
-The quenching mechanism in \textsc{Jewel} probed by telescoping deconstruction variables above can be further illuminated using the Lund diagram \cite{Andersson1989}.
+The quenching mechanism in \textsc{Jewel} probed by telescoping deconstruction variables and soft-drop can be further illuminated using the Lund diagram \cite{Andersson1989}.
 The Lund diagram~\cite{Andersson1989} is a diagrammatic jet representation associated with the Cambridge/Aachen clustering tree structure.

### 2018-03-09 qg_ML.tex
paper version ready for arxiv submission

@@
 	   \centering
-	   \includegraphics[width=0.7\textwidth]{plots/qg_HI}
+	   \includegraphics[width=0.7\textwidth]{Fig1}
 	   \caption{Illustration of the interplay between jet quenching (vertical) and quark gluon discrimination (horizontal). }
@@
 	   \centering
-	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
-	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
-	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
-	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
-	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
+	   \includegraphics[width=0.48\textwidth]{Fig2a}
+	   \includegraphics[width=0.48\textwidth]{Fig2b}
+	   \includegraphics[width=0.48\textwidth]{Fig2c}
+	   \includegraphics[width=0.48\textwidth]{Fig2d}
+	   \includegraphics[width=0.48\textwidth]{Fig2e}
 	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets.
@@
 \centering
-\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
-\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
-\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
-\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
+\includegraphics[width=0.45\textwidth]{Fig3a}
+\includegraphics[width=0.45\textwidth]{Fig3b}
+\includegraphics[width=0.45\textwidth]{Fig3c}
+\includegraphics[width=0.45\textwidth]{Fig3d}
 \caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}.
@@
 \centering
-\includegraphics[width=.49\columnwidth]{figures/QCDjetT1_telescoped2.pdf}
-\includegraphics[width=.49\columnwidth]{figures/QCDjetT2_telescoped2.pdf}
-\includegraphics[width=.49\columnwidth]{figures/QCDjetT3_telescoped2.pdf}
+\includegraphics[width=.49\columnwidth]{Fig4a}
+\includegraphics[width=.49\columnwidth]{Fig4b}
+\includegraphics[width=.49\columnwidth]{Fig4c}
 \caption{Illustration of telescoping deconstruction at T1 (top left panel), T2 (top right panel) and T3 (bottom panel) orders for a random pp jet.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_zg_ppcomp_SD_TD2_new.pdf}
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_ppcomp_SD_TD2_new.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig5a}
+	   \includegraphics[width=0.9\textwidth]{Fig5b}
 	   \caption{Top panels: Distributions of groomed momentum sharing $z_g$ (left) and the T2 order TD subjet momentum fraction $z_{\rm T2}$ (right); Bottom panels: Distributions of groomed jet radius $r_g$ and the T2 order TD angular separation between subjets $\theta_{\rm T2}$ for quark (blue) and gluon (red) jets in pp \jw simulations.}
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_z_new.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig6}
 	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum fraction $z_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_new.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig7}
 	   \caption{Distributions of groomed jet radius $r_g$ (left) and TD angular separation between subjets $\theta_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig8}
 	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (left) and soft (right panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (top) and central PbPb (bottom) collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig9}
 	   \caption{Distributions of $\delta m=\sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
+	\includegraphics[width=0.9\textwidth]{Fig10}
 	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
+	\includegraphics[width=0.9\textwidth]{Fig11}
 	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
@@
 	   \centering
-	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
+	   \includegraphics[width=0.9\textwidth]{Fig12}
 	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_pp_2p76TeV_genLevel_QvsG.pdf}
+	\includegraphics[width=0.8\textwidth]{Fig13}
 	\caption{ROC curves of quark v.s.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
+	\includegraphics[width=0.8\textwidth]{Fig14}
 	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}, similar to Figure~\ref{fig:ROC_pp}.}
@@
 	\centering
-        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
+        \includegraphics[width=0.8\textwidth]{Fig15}
 	\caption{ROC curves of quark v.s.
@@
 	\centering
-        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
+        \includegraphics[width=0.8\textwidth]{Fig16}
 	\caption{ROC curves of quark v.s.
@@
 	   \centering
-	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
-	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
+	   \includegraphics[width=0.62\textwidth]{Fig17a}
+	   \includegraphics[width=0.62\textwidth]{Fig17b}
 	   \caption{ROC curves of pp-jet v.s.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
-\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
+\includegraphics[width=.32\columnwidth]{Fig18a}
+\includegraphics[width=.32\columnwidth]{Fig18b}
+\includegraphics[width=.32\columnwidth]{Fig18c}
+\includegraphics[width=.32\columnwidth]{Fig18d}
+\includegraphics[width=.32\columnwidth]{Fig18e}
+\includegraphics[width=.32\columnwidth]{Fig18f}
 \caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{figures/QG_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/W_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/Top_invROC.pdf}
+\includegraphics[width=.32\columnwidth]{Fig19a}
+\includegraphics[width=.32\columnwidth]{Fig19b}
+\includegraphics[width=.32\columnwidth]{Fig19c}
 \caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
-\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
-\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
-\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
-\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
-\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
+\includegraphics[width=.32\columnwidth]{Fig20a}
+\includegraphics[width=.32\columnwidth]{Fig20b}
+\includegraphics[width=.32\columnwidth]{Fig20c}
+\includegraphics[width=.32\columnwidth]{Fig20d}
+\includegraphics[width=.32\columnwidth]{Fig20e}
+\includegraphics[width=.32\columnwidth]{Fig20f}
 \caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row).

### 2018-04-21 qg_ML.tex
YT updated the draft for JHEP submission

@@
 We begin with a physics-motivated, multivariate analysis of jet substructure observables including the jet mass, the radial moments, the $p_T^D$ and the pixel multiplicity.
-In comparison, we employ state-of-the-art image-recognition techniques by training a deep convolutional neutral network on jet images.
+In comparison, we employ state-of-the-art image-recognition techniques by training a deep convolutional neural network on jet images.
 To systematically extract jet substructure information, we introduce the telescoping deconstruction framework exploiting subjet kinematics at multiple angular scales.
@@
 From the measurements of the jet shape \cite{Ellis:1992qq,Chatrchyan:2013kwa,Khachatryan:2016tfj,Seymour:1997kj,Li:2011hy,Li:2012bw,Vitev:2008rz,Vitev:2009rd} and jet fragmentation function \cite{Procura:2009vm,Chatrchyan:2012gw,Aad:2014wha,Chatrchyan:2014ava,Aaboud:2017bzv,Sirunyan:2018qec}, the community has recently established that both transverse and longitudinal momentum distributions inside jets are significantly modified.
-Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets at the LHC may significantly contribute to substructure modifications observed in experiments.
+Since jet substructures depend strongly on the partonic origin of jets \cite{Gallicchio:2011xq,Gallicchio:2012ez,Chien:2012ur,Chien:2015ctp,Chien:2014nsa}, a change in the fraction of quark-initiated jets and gluon-initiated jets from proton-proton to heavy ion collisions may significantly contribute to substructure modifications observed in experiments.
 It has recently been realized that an increase of the quark jet fraction in heavy ion collisions can contribute to the narrowing of jet cores as suggested in various inclusive jet substructure measurements~\cite{Chien:2015hda,Spousta:2015fca}.
@@
 	   \includegraphics[width=0.9\textwidth]{Fig12}
-	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+	   \caption{Lund diagrams for branches in the hard branch ($z>z_{\rm cut}=0.1$) after the soft-drop procedure for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}.
+Note the significant fraction of branches with $z<z_{\rm cut}$ and the increase of branches with $z\gtrsim z_{\rm cut}$ at wide angles in PbPb collisions.
 See text for the definition of the axes.}
@@
 The authors are grateful to Patrick Komiske and Eric Metodiev for collaborating on the development of telescoping deconstruction and contributions to the Appendix.
-The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions.
+The authors also thank Fr\'ed\'eric Dreyer, Benjamin Nachman, Joern Putschke, Jesse Thaler, Konrad Tywoniuk and Sevil Salur for helpful discussions and comments.
 We thank Gavin Salam and the participants of the CERN TH institute ``{\sl Novel tools and observables for jet physics in heavy ion collisions}" and the 5-th Heavy Ion Jet Workshop for bringing to our attention the Lund Diagram and its usage in studying jet modifications.

### 2018-11-19 referee_response_v2.txt
first update of referee responses and small text updates to the draft

We would like to thank the referee for the report which helps improve the manuscript, and we have revised the paper accordingly. Detailed replies can be found below.

First of all, we would like to clarify that, the main point of our work is a systematic and comprehensive study of jet quenching in heavy ion collisions using quark (q) and gluon (g) jet substructure as independent probes, instead of what has been summarized in the beginning of the referee report as discrimination between q vs g jets which is only a part of our work. We would also like to clarify that, the study of discriminating quenched and unquenched jets is in fact very natural and at the heart of all jet quenching studies. The goal is to identify and quantitatively describe various aspects of jet modifications, including the dependence on the parton flavor that can lead us towards determining the transport properties of the QGP medium produced in heavy ion collisions. Looking at the q-g classification performance plots is one way to quantify the differences regarding medium modifications to jet (sub)structure which can be captured by the different basis (physics motivated variables, jet images or the telescoping deconstructed variables) one chooses to study jet quenching.

On the other hand, telescoping deconstruction (TD) is developed and applied in this work. In comparison to the original telescoping jets method, that one of us developed a few years ago, this paper presents a crucial improvement leading to TD which is an entirely new framework consisting of jet substructure basis that can in turn be used to represent the various as aspects of jet information. Furthermore, TD is a complete and systematic expansion of jet information utilizing subjets which is analogous to the expansion in perturbation theory, where in the expansion converges rather quickly. We would like to emphasize the following: TD concretely defines all the relevant observables within the framework and gives well-organized jet information. Furthermore, we show in the draft that they capture physical aspects of not only QCD jets but also W and top jets. The role that TD plays in our paper is to provide an answer to how one bridges the conventional approach of studying jets using a list of physics-motivated variables, to modern machine-learning approach of studying jets using neural network to exhaustively extract useful features. This implies that TD should be a very useful and reliable framework for jet physics studies, which on its own is a novel jet representation method. Nevertheless, we describe in detail, the TD method in the Appendix in order not to distract the readers from the main, physical point of our paper regarding flavor dependence of jet quenching in heavy ion collisions.

To answer the referee's questions,

In the original telescoping jets paper, only a single variable called volatility was used to capture certain aspects of jets. On the other hand, TD is a jet substructure basis which is designed to represent, both qualitatively and quantitatively all aspects of jets. As emphasized in the above paragraph, TD is a fixed-order subjet expansion which converges quickly, as implied by the moderate, but characteristic increase of the quark gluon discrimination performance as the order increases. In the case of top jet tagging, TD not only captures the three-prong, hard jet substructure but also other jet features. For example, the color-singlet isolation of W jets and the partonic origin of subjets, which become manifest at N = 2 and N = 3 orders, but are not seen at the N = 1 order.

Figure 18 shows the comparisons of different orders of TD w/ a CNN for the various classification problems such as Top vs QCD, W vs QCD and quark vs gluon jets. If we look at the right most panel, at a given top efficiency of 30% (which is a nominal working point in experiments), we find that the mistag rate ~ 1 in 200 for T1 and as we increase the orders, the rate becomes quite small ~ 1 in 450 which is greater than a 200% improvement. Thus, in going to higher orders of TD, the structure of the decaying top is better represented with encapsulating more of the interesting physics that separates it from a QCD background jet.

A neural network consisting of a Multi-Layer Perceptron with hidden layer(s) (which we describe in the paper) is conveniently used to combine the TD variables and is much simpler than the convolutional neural network that is employed for the jet image method (which we also describe in the paper). The training and testing of the simple neural network can be done in the order O(10s)-O(100s) on most modern laptops. One could have used other MLPs and their variants such as boosted decision tree (BDT) or Random Forests (RF) which are getting quite popular in the literature but that would shift the main focus of the paper to a study of different methods which is not what we intended. The use of multivariate methods available on the market is to primarily combine variables in a non-intuitive linear/non-linear combination (due to the use of non-linear activation functions and hidden layers) and make optimal use of the correlations among them that a traditional analysis using likelihood ratio methods is likely to ignore.

We thank the referee for the raising the question regarding the impact of a realistic underlying event/large background fluctuations. We are aware and intend to perform a further study to address the concern by comparing the q-g discrimination performance for a variety of underlying event descriptions and jet quenching models in a new paper. However, we would like to first clarify that, in our current JEWEL study, the pp and AA jets are represented as a collection of ideally background-subtracted jets. The pp jets are essentially generated by a version of PYTHIA-6 which is implemented in JEWEL simulations without pileup (but with a model of the pp underlying event). The AA jets in JEWEL include both with and without medium recoil contributions and the jets with recoils are background-subtracted using the GridSub method introduced by one of the authors in an earlier publication (in JHEP). However, JEWEL does not generate/simulate the entire heavy ion background, as we have emphasized in the draft, but it includes interactions of the propagating parton with the thermalized medium model implemented in JEWEL. With the application of the GridSub method, one can study an idealized situation of quenched jets which are comparable to unquenched vacuum jets. With the implementation of these procedures, JEWEL describes the jet suppression and a variety of jet substructure modifications (jet shape and splitting function, for example) quite well. In order to extract any meaningful physics statements about flavor dependent jet quenching, it is essential for our model to provide a satisfactory description of data and JEWEL indeed does that. Our envisioned goal for this paper is to present the first quantification of the baseline comparison of quenched quark vs quenched gluon jets and follow up with a study including the underlying event specifically utilizing new MC toolkits that have surfaced in the community since the time of our submission (such as JETSCAPE).

We have mentioned in the paper that the reason why we choose JEWEL for our study is precisely due to its qualitative description of jet quenching measurements at the LHC especially with the inclusions of recoils and the background subtraction methods.

It is indeed correct that pTD and particle multiplicity are IRC unsafe at the particle level. However, the jet constituents used in our work to evaluate all jet observables are essentially grouped into “calorimeter-like towers” with pixel size 0.08 x 0.08 to reflect experimental limitations. This regulates all the IRC singularities and makes our observables IRC safe. Nevertheless, this also implies that jet features that are inherently beyond the angular resolution of 0.08 x 0.08 would be lost in our analysis. This is one of the reasons why the q-g discrimination performances are largely similar amongst the three methods used in this work. Furthermore, these observables such as pTD, have indeed been measured in heavy ion collisions by the ALICE collaboration and JEWEL does a good job of reproducing the structure as seen in data without any additional tunings whatsoever. There have also been publications by both CMS and ATLAS with the use of multiplicity and pTD specifically in the use of q/g classifications.

We would like to clarify the comment from the referee about the use of soft drop observables. We used TD to illuminate the inner working of the soft drop algorithm by comparisons between two methods. Taking into account the limitations imposed by soft drop, and we believe that TD is more flexible and robust by having an additional handle in the order at which TD is calculated. We find that TD is even more related to a variation of soft drop called recursive soft drop, and one of us had many discussions with the recursive soft drop authors. We would also like to clarify that, soft drop is not just a grooming method, and TD can be used as a grooming method. Soft Drop has been used to define observables which are physically meaningful, for example, the groomed momentum sharing zg or the splitting function. On the other hand, TD defines exclusively the subjet regions and disregards radiation in the outside region, which is exactly a grooming procedure.

In fact, what soft drop does is similar to TD by identifying the two subjet branches while dropping soft, wide angle radiation. Since soft drop is based on the C/A de-clustering tree which is angular ordered, the size of the subjet branches has to be smaller than the angle between the branches. This is the reason why in the comparison of TD to soft drop we choose the subjet size to be proportional to the angle between the subjet axes. The prefactor of 0.3, which can surely be varied, offers an additional degree of freedom within the TD framework to scan the energy distribution surrounding the individual subjets. We do vary the prefactor in our studies and show the case with the prefactor 0.3 just as an example, and it is not meant to be a fixed number.

We would like to emphasize again that, the main point of our paper is to use quark and gluon jet substructure to probe heavy ion collisions, instead of simply performing q-g discrimination. It is then very natural to discuss the physical meaning of the TD variables which we use to study jet quenching. The same clarification applies also to the use of the Lund diagram technique, which is another jet substructure representation closely related to the C/A re-clustered tree and soft drop. Our use of Lund diagrams in this paper is to drive home to point of differences between q/g vacuum/quenched jets. Since the Lund diagram is necessarily crafted by using soft drop recursively on the leading prong, it records the leading branch and its kinematics is intimately related to TD, which represents jets using subjet. In the draft, we have highlighted this specific point and tried to make clear that the physics picture we learn from TD and physics variables are also reflected in the Lund diagram. We have modified the draft to better reflect the discussion to the reader.

We show how jet modifications can be probed in the TD framework by looking at the jet substructure observables order-by-order, and at each order we see new information. We show that the information zg carries in soft drop can be studied in the TD framework more generally, and we further look at the angular distribution of subjets, as well as subjet substructure like the subjet masses. Even more, we can construct higher-level observables using the basis observables and give an example of delta m^2 between groomed and ungroomed jet masses, and these observable shows striking signatures of jet quenching as modelled in JEWEL by highlighting the differences between quark and gluon jets. Such an observable could only be created by looking at the impact of q-g discrimination on TD at various orders. We have updated the draft to clearly describe the usefulness of these comparisons and that we urge different experiments to measure these observables to make definitive statements on quark and gluon energy loss.

With reference to the referee’s comments that it is not clear that TD systematically improves q/g discrimination leading to insights about jet quenching mechanism as implemented in MC simulations, we note the following. Figure 13 and 14 shows the comparisons of the ROC curves for the various methods used in the analysis and we find that the TD at 3rd order performs almost as good as the DCNN if not better. With the use of TD, we can extract the corresponding variables that facilitate this performance and these extractions are done in the fragmentation basis of the jet structure (as we discussed in the above paragraphs and in the paper). Since it is quite complicated to extract any physics from the convolution layers in a DCNN, TD provides an excellent option to study inherent differences without any biases.

One of the main reasons why we showed performance plots for discriminating between quenched and unquenched jets is to study the effect of quenching between q vs g jets as implemented in JEWEL. As we mentioned in the paper and shown in Figure 17, all the methods showcase better performance for differentiating quenched gluon jet from an unquenched gluon jet as opposed to quark jets, leading to our conclusion that gluon jets are further modified. We have edited the text to better reflect on a use case for this particular study.

It is a qualitative feature of JEWEL-simulated heavy ion collisions that the q-g discrimination performance decreases from pp to AA. We are actively invested in studying how other jet quenching models and calculations reflect in our methods. Such a study offers us to the opportunity to solve the inverse problem of mapping from jet modifications to the inner working of jet quenching using a consistent framework that we have introduced in this paper.

Sincerely,

### 2018-11-19 qg_ML.tex
first update of referee responses and small text updates to the draft

--- 2018-03-09 qg_ML.tex
@@
 		\item Jet mass: $m = \sqrt{(\sum_{i\in {\rm jet}} E_i)^{2} - (\sum_{i\in {\rm jet}} \vec{p_i})^2}$.
-	        \item First radial moment (girth): $\sum_{i \in {\rm jet}} p_{T_i} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\phi^2_{{\rm jet}, i}}$.
+	        \item First radial moment (girth): $\sum_{i \in {\rm jet}} p_{T_i} \Delta R_{{\rm jet}, i}/p^{\rm jet}_{T}$, where $\Delta R_{{\rm jet}, i}=\sqrt{\Delta \eta^2_{{\rm jet}, i}+\Delta\phi^2_{{\rm jet}, i}}$.
 		\item 0.5 radial moment:  $\sum_{i \in {\rm jet}} p_{T_i} (\Delta R_{{\rm jet}, i})^{0.5}/p^{\rm jet}_{T}$.
@@
 Figure~\ref{fig:comp_subjet_m} shows the distributions of $\lambda\equiv m/p_T$, which is the ratio between subjet mass and subjet $p_T$, of hard subjets (left panels) and soft subjets (right panels) for quark (blue dashed) and gluon (red dashed) jets.
-Here hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
+Here, hard or soft refers to subjet momenta with harder meaning the one with higher transverse momentum.
 We see hints that the $\lambda$ distributions of hard subjets exhibit flavor dependence, while those of soft subjets are similar for quark and gluon jets.
@@
 This suggests that the qualitative feature of the soft event activities in the \jw simulations, persists to affect not only the subjet topology but also the subjet substructure.
+The advantage of utilizing the subjet topological features arising from telescoping at the third order is that these observables can be studied for a variety of MC models and furthermore, can be measured in data at the LHC.
+In effect, these observables offers predictions for flavor dependent jet quenching effects and help constrain the jet quenching phenomenon.
 
@@
 The quenching mechanism in \textsc{Jewel} probed by telescoping deconstruction variables and soft-drop can be further illuminated using the Lund diagram \cite{Andersson1989}.
-The Lund diagram~\cite{Andersson1989} is a diagrammatic jet representation associated with the Cambridge/Aachen clustering tree structure.
+The Lund diagram~\cite{Andersson1989} is a diagrammatic jet representation associated with the Cambridge/Aachen de-clustering tree structure.
 It directly records the branching kinematics along the hard branches in the C/A clustering tree in terms of longitudinal momentum fraction $z$ of the soft branch and the angular separation $\Delta R$ between the two branches.
@@
 The variables are chosen such that the Lund diagram is uniform in the fixed strong coupling constant limit for jets in pp collisions due to the infrared structure of QCD and as such, they have been shown to be useful in distinguishing between QCD and boosted heavy particle jets \cite{Salam:2016yht}.
+
+In the case of heavy ion jets we are necessarily interested in the interplay between the soft and the hard scales in the jets and jet substructure techniques probed by the telescoping deconstruction and softdrop highlight these differences via observables and measurements.
+The Lund Diagram however, offers a visualized impact of jet quenching for a given jet in its harder prong.
+There have been recent studies in the literature~\cite{Dreyer:2018nbf} where Lund diagrams themselves can be utilized as a jet representation for jet classification problems by the use of state of the art neural networks.
+In this paper, we will utilize Lund diagrams to study the physics of jet quenching and substructure modification between quark and gluon jets as implemented in \textsc{Jewel}.
 
@@
 The above studies demonstrate that telescoping deconstruction provides a physical organization of jet information where qualitatively new features can be captured order-by-order in the subjet expansion.
-It systematically improves quark gluon discrimination, leading to insights about jet quenching mechanisms as implemented in Monte Carlo simulations.
+By studying the distributions of telescoping quantities at varying orders, we are able to study observables that offer a higher degree of constraint regarding jet quenching mechanisms and its flavor dependence.
 Going beyond to higher T$N$ orders or using subjet superstructure \cite{Gallicchio:2010sw} and subjet charge \cite{Krohn:2012fg} information are beyond the scope of this work and left for future studies.
@@
 Quenched jets end up with significantly more jet constituents which are distributed away from the jet axis, resulting in the saturation of the TD classification performance at the T1 order.
-This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and more quenched jets.
+This modification appears to affect gluon jets significantly more than quark jets in \jw since a gluon jet, with its larger particle multiplicity and wider parton shower, can interact with a larger number of scattering centers leading to broader and significantly more quenched jets.
+
+Such a study is however predicated on identifying collections of quark and gluon jets in both AA and pp collisions.
+Experiments that have a full hermetic coverage jet detectors can take advantage of kinematic and event selection constraints such as forward-rapidity vs mid-rapidity or photon/Z-jets vs dijets at varying \pt intervals in order to select these flavor specific collections.
+One could also use any of the aforementioned classification methods such as DCNNs or TD to select these jet collections and perform further differential analyses.
+Another intriguing possibility of such a model i.e that can pick out quenched jets vs unquenched jets, is to test it on high multiplicity pA or pp collisions.
+Measurements from the LHC~\cite{ALICE:2017jyt, Khachatryan:2016txc} have shown that these events are rather special in that they showcase AA like collective behavior but do not appear to have any significant jet quenching or parton energy loss signatures.
+This could be a test of QGP like behavior in these high multiplicity collisions which would in turn have wide reaching inferences regarding formation of the QGP in small systems.
 

### 2018-11-20 qg_ML.tex
one more run through text

@@
 On the other hand, we see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively, with the caveat that pixel multiplicity is unreliable in a heavy ion environment and the $p_T^D$ is an IRC unsafe observable.
+Even though these are IRC unsafe, we study them for two primary purposes .
+Firstly, with our GridSub we are effectively combining these into discrete cells which reduces the unsafe behavior.
+Secondly these observables have been historically used in measurements of identifying quark and gluon jets in pp collisions at the LHC~\cite{ATL-PHYS-PUB-2017-017, CMS-DP-2016-070, CMS-PAS-JME-13-002}.
 
@@
 Such a study is however predicated on identifying collections of quark and gluon jets in both AA and pp collisions.
-Experiments that have a full hermetic coverage jet detectors can take advantage of kinematic and event selection constraints such as forward-rapidity vs mid-rapidity or photon/Z-jets vs dijets at varying \pt intervals in order to select these flavor specific collections.
+Experiments that have a full hermetic coverage jet detectors can take advantage of kinematic and event selection constraints such as forward-rapidity vs mid-rapidity or photon/Z-jets vs dijets at varying \pt~intervals in order to select these flavor specific collections.
 One could also use any of the aforementioned classification methods such as DCNNs or TD to select these jet collections and perform further differential analyses.
-Another intriguing possibility of such a model i.e that can pick out quenched jets vs unquenched jets, is to test it on high multiplicity pA or pp collisions.
+Another intriguing possibility of such a model that can pick out quenched jets vs unquenched jets, is to test it on high multiplicity pA or pp collisions.
 Measurements from the LHC~\cite{ALICE:2017jyt, Khachatryan:2016txc} have shown that these events are rather special in that they showcase AA like collective behavior but do not appear to have any significant jet quenching or parton energy loss signatures.

### 2020-02-24 qg_ML.tex
added figure of embedding comparision and one paragraph

@@
 With the application of the grid subtraction (GridSub) technique, the collection of jets in \jw represent ideally subtracted jets in the experiment.
-This enables us to focus on and highlight the effects of jet quenching, as opposed to additional, nontrivial smearing effects.
+This enables us to focus on and highlight the effects of jet quenching, as opposed to additional, nontrivial smearing effects due to the fluctuating underlying event.
 
@@
 On the other hand, the performance drops very slightly (change in auc of $0.01$) with the background subtraction performed.
-This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction~\footnote{Underlying events and their subtractions in experiments can both influence and introduce jet substructure features.
-Such effects can only be studied with full heavy ion event simulations, which we leave for future studies.}.
+This suggests that some intrinsic differences between quark jets and gluon jets are possibly removed in the subtraction.
+
+\begin{figure}[h]
+	\centering
+        \includegraphics[width=0.8\textwidth]{Fig17}
+	\caption{ROC curves of quark v.s.
+gluon efficiency highlighting the effect of the embedded \textsc{Hydjet} background.
+The dashed curves represent the embedded event in comparison to the solid curves which show the vacuum events.
+The different colored curves correspond to various classification models studied in this paper.}
+	\label{fig:ROC_qvsg_wHydjet}
+\end{figure}
+
+The fluctuating underlying event and their subtraction in experiment can both influence and introduce features in jet substructure.
+We employ the embedding technique which is used in almost all experiments to estimate the effect of these fluctuations on a event by event basis.
+We take an unbiased heavy ion event produced by the \textsc{Hydjet}~\cite{Lokhtin:2008xi} MC for most central collisions ($0-5\%$) and embed the particles produced by \jw (both vacuum and in-medium) together.
+In this embedded event, we have included the effect of the uncorrelated background to jets.
+Since the \jw events with jet quenching are for a centrality window of $0-20\%$, this underlying event presents an over estimate of the background which we take as an upper bound for the effects due to the underlying event.
+In order to extract jets from these embedded events, we perform constituent subtraction~\cite{Berta:2014eza} which reduces the contribution of the background to the jet and its constituents.
+The effect of including the heavy ion underlying event is shown in Figure~\ref{fig:ROC_qvsg_wHydjet} in the dashed curves having an overall reduction in classification performance as shown by a reduced auc for both vacuum jets and quenched jets but we would like to note that classification is still experimentally viable albeit with reduced efficiency.
+Further studies of the exact nature of the background influence on jet substructure via detector simulations are reserved for future publications.
 
@@
 	   \centering
-	   \includegraphics[width=0.62\textwidth]{Fig17a}
-	   \includegraphics[width=0.62\textwidth]{Fig17b}
+	   \includegraphics[width=0.62\textwidth]{Fig18a}
+	   \includegraphics[width=0.62\textwidth]{Fig18b}
 	   \caption{ROC curves of pp-jet v.s.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{Fig18a}
-\includegraphics[width=.32\columnwidth]{Fig18b}
-\includegraphics[width=.32\columnwidth]{Fig18c}
-\includegraphics[width=.32\columnwidth]{Fig18d}
-\includegraphics[width=.32\columnwidth]{Fig18e}
-\includegraphics[width=.32\columnwidth]{Fig18f}
+\includegraphics[width=.32\columnwidth]{Fig19a}
+\includegraphics[width=.32\columnwidth]{Fig19b}
+\includegraphics[width=.32\columnwidth]{Fig19c}
+\includegraphics[width=.32\columnwidth]{Fig19d}
+\includegraphics[width=.32\columnwidth]{Fig19e}
+\includegraphics[width=.32\columnwidth]{Fig19f}
 \caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{Fig19a}
-\includegraphics[width=.32\columnwidth]{Fig19b}
-\includegraphics[width=.32\columnwidth]{Fig19c}
+\includegraphics[width=.32\columnwidth]{Fig20a}
+\includegraphics[width=.32\columnwidth]{Fig20b}
+\includegraphics[width=.32\columnwidth]{Fig20c}
 \caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging.
@@
 \centering
-\includegraphics[width=.32\columnwidth]{Fig20a}
-\includegraphics[width=.32\columnwidth]{Fig20b}
-\includegraphics[width=.32\columnwidth]{Fig20c}
-\includegraphics[width=.32\columnwidth]{Fig20d}
-\includegraphics[width=.32\columnwidth]{Fig20e}
-\includegraphics[width=.32\columnwidth]{Fig20f}
+\includegraphics[width=.32\columnwidth]{Fig21a}
+\includegraphics[width=.32\columnwidth]{Fig21b}
+\includegraphics[width=.32\columnwidth]{Fig21c}
+\includegraphics[width=.32\columnwidth]{Fig21d}
+\includegraphics[width=.32\columnwidth]{Fig21e}
+\includegraphics[width=.32\columnwidth]{Fig21f}
 \caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row).

### 2020-02-25 qg_ML.tex
updated figure17 with the embedding curve with DCNN and text

@@
 	\caption{ROC curves of quark v.s.
-gluon efficiency highlighting the effect of the embedded \textsc{Hydjet} background.
-The dashed curves represent the embedded event in comparison to the solid curves which show the vacuum events.
-The different colored curves correspond to various classification models studied in this paper.}
+gluon efficiency with the DCNN method highlighting the effect of the embedded \textsc{Hydjet} background shown in the red curves in comparison to the blue curves which are in vacuum at the generator level.
+The solid and dotted  curves correspond to PbPb quenched and pp events respectively.}
 	\label{fig:ROC_qvsg_wHydjet}
@@
 The fluctuating underlying event and their subtraction in experiment can both influence and introduce features in jet substructure.
-We employ the embedding technique which is used in almost all experiments to estimate the effect of these fluctuations on a event by event basis.
+We employ the embedding technique which is used in almost all experiments to estimate the effect of these fluctuations on an event by event basis.
 We take an unbiased heavy ion event produced by the \textsc{Hydjet}~\cite{Lokhtin:2008xi} MC for most central collisions ($0-5\%$) and embed the particles produced by \jw (both vacuum and in-medium) together.
@@
 In order to extract jets from these embedded events, we perform constituent subtraction~\cite{Berta:2014eza} which reduces the contribution of the background to the jet and its constituents.
-The effect of including the heavy ion underlying event is shown in Figure~\ref{fig:ROC_qvsg_wHydjet} in the dashed curves having an overall reduction in classification performance as shown by a reduced auc for both vacuum jets and quenched jets but we would like to note that classification is still experimentally viable albeit with reduced efficiency.
+The effect of including the heavy ion underlying event is shown in Figure~\ref{fig:ROC_qvsg_wHydjet} in the dashed curves having an overall reduction in DCNN classification performance as shown by an increased auc for both vacuum jets and quenched jets but we would like to note that classification is still experimentally viable albeit with reduced efficiency.
+We test the DCNN method for this embedding study since it is agnostic to jet substructure artifacts introduced during the embedding and background subtraction procedures.
 Further studies of the exact nature of the background influence on jet substructure via detector simulations are reserved for future publications.

### 2020-08-27 qg_ML.tex
updated draft with separate appendix for model details and added figures for networks

@@
 
-We combine the five jet substructure observables using a multivariate model implemented in Keras \cite{keras}, with two hidden dense layers of 10 nodes each.
-We use the rectified linear unit (ReLU) \cite{nair2010rectified} and sigmoid activation functions for the dense layers and the output layer, respectively.
-The model training utilizes the Adam optimizer \cite{adam} and the binary cross entropy loss function.
-We perform cross validation using Monte Carlo samples which are split into two random halves for training and testing.
+We combine the five jet substructure observables using a multivariate model implemented in Keras \cite{keras}. The details of the model description are given in Appendix~\ref{app:models}.
 
@@
 
-A deep convolutional neural network is implemented in Keras with the TensorFlow \cite{DBLP:journals/corr/AbadiBCCDDDGIIK16} backend.
-The neural network model consists of three convolution layers of sequentially reducing sizes, each with 8 filters followed by a max-pooling layer.
-The convolution layers use a hyperbolic tangent activation function, and the output layer uses a sigmoid activation function.
-The output layer is preceded by three layers, a dropout layer with a rejection score of 0.20, a dense layer with 20 nodes and an additional dropout layer with a reduced rejection score of 0.10.
-The additional dropout layer helps filtering less important features from previous layers and thus increases the speed and efficiency of the model training.
-The model is trained using the Adam optimizer with binary cross entropy loss function.
+A deep convolutional neural network is implemented in Keras~\cite{DBLP:journals/corr/AbadiBCCDDDGIIK16}.
+The neural network architecture and the corresponding activation functions are detailed in the appendix.
 
@@
 \section*{Appendix}
+
+\section{Description of the MLP and CNN network architectures}
+\label{app:models}
+The MLP utilized in this classification analysis utilizes two hidden dense layers of 10 nodes each as shown in Fig~\ref{fig:mlp}.
+We use the rectified linear unit (ReLU) \cite{nair2010rectified} and sigmoid activation functions for the dense layers and the output layer, respectively.
+The model training utilizes the Adam optimizer \cite{adam} and the binary cross entropy loss function.
+We perform cross validation using Monte Carlo samples which are split into two random halves for training and testing.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.6\columnwidth]{FigMLP.png}
+\caption{Schematic of the MLP utilized for physics observables where the two hidden layers are shown and particular nodes that are not linked are dropped (shown as an example) during training.}
+\label{fig:mlp}
+\end{figure}
+
+The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer.
+The network architecture is schematically shown in Fig~\ref{fig:dcnn}.
+A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2.
+The dropout rate was taken to be 0.1 for all layers.
+All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}.
+Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used.
+The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.8\columnwidth]{FigDCNN_wolabels.png}
+\caption{Schematic of the DCNN with maxpooling layers (not shown) located in-between each convolution layer. The output layer for classification is taken as a result of a dense layer (not shown).}
+\label{fig:dcnn}
+\end{figure}
+
 \section{Probing subjet energy flows with telescoping deconstruction}
@@
 The jet images are pre-processed and standardized according to the procedure in \Ref{Komiske:2016rsd}.
-The CNN architecture consists of three 48-filter convolutional layers with filter sizes of $8\times 8$, $4\times 4$, and $4\times 4$ followed by a 128-unit dense layer and a 2-unit softmaxed output layer.
-A $2\times 2$ maxpooling is performed after each convolutional layer with a stride length of 2.
-The dropout rate was taken to be 0.1 for all layers.
-All neural networks are implemented using the Python deep learning library Keras~\cite{keras} with the Theano backend~\cite{bergstra2010theano}.
-Rectified linear unit (ReLU) activation functions~\cite{nair2010rectified} and He-uniform model weight initialization~\cite{heuniform} are used.
-The networks are trained using the Adam algorithm~\cite{adam} with a learning rate of $10^{-3}$ and a batch size of 256 for 50 epochs with a patience parameter of 8, and the best model is selected based on validation set performance.
 
